---
title: "MSC QC and preprocessing"
date: '2022-07-11'
author: 'Jesse Rop'
output:
  html_document:
    toc: true
    theme: united
    code_folding: hide
    df_print: paged
---

<!-- Refactor changelog: removed obvious unused libraries (rmarkdown, parallel, SeuratData, SeuratDisk, plotly, Hmisc, RColorBrewer, ggvenn, MetBrewer, ggh4x, EnhancedVolcano, DropletUtils, ggforce, scran, SingleR, scDblFinder); added small helpers for repeated file loading and RDS saving; no duplicate knitr chunk labels were present. -->

```{r}
suppressPackageStartupMessages({
  library(Seurat)
  library(dplyr) 
  library(stringr)
  library(purrr)
  library(readr)
  library(tibble)
  library(ggplot2)
  library(tidyr)
  library(conflicted)
  library(BPCells)
  
  conflict_prefer("select", "dplyr")
  conflict_prefer("filter", "dplyr")
  conflict_prefer("rename", "dplyr")
  conflict_prefer("map", "purrr")
  conflict_prefer("count", "dplyr")
  conflicts_prefer(base::saveRDS)
  conflicts_prefer(base::as.data.frame)
  conflicts_prefer(base::intersect)
  conflicts_prefer(dplyr::desc)
  conflicts_prefer(dplyr::count)
})

```

```{r, setup}
knitr::opts_knit$set(root.dir = '/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/Talleh')
```
### Variable declarations

#### Source common functions, variables and tables
```{r}
##Boxplot labeling function
source("~/Talleh_code/project_plotting_fns.R")

# Helper inputs:
# - file_paths: character vector of files to name and keep in sample order.
# - strip_pattern: regex removed from each path to recover sample names.
load_named_paths <- function(file_paths, strip_pattern) {
  file_paths %>%
    map(~.x %>% set_names(nm = str_remove_all(., strip_pattern))) %>%
    unlist(recursive = FALSE) %>%
    .[sample_name]
}

# Helper inputs:
# - file_paths: named Cell Ranger or CellBender matrix paths.
# - strip_pattern: regex removed from each path to recover sample names.
# - output_subdir: directory written under data/processed/Pf/<sample>/bpcells/.
write_sample_bpcells <- function(file_paths, strip_pattern, output_subdir) {
  load_named_paths(file_paths, strip_pattern) %>%
    map(~open_matrix_10x_hdf5(path = .x)) %>%
    imap(~write_matrix_dir(mat = .x, dir = paste0("data/processed/Pf/", .y, "/bpcells/", output_subdir), overwrite = T))
}

# Helper inputs:
# - matrix_dirs: character vector of matrix directories to read.
# - strip_pattern: regex removed from each path to recover sample names.
read_sample_bpcells <- function(matrix_dirs, strip_pattern) {
  load_named_paths(matrix_dirs, strip_pattern) %>%
    map(~open_matrix_dir(dir = .x)) %>%
    imap(~CreateSeuratObject(counts = .x, project = paste0(.y, ".mrna")))
}

# Helper inputs:
# - file_paths: character vector of table files.
# - strip_pattern: regex removed from each path to recover sample names.
# - reader: import function such as read_csv or read_tsv.
load_named_tables <- function(file_paths, strip_pattern, reader) {
  load_named_paths(file_paths, strip_pattern) %>%
    map(reader)
}

```

## Select Sample Names

```{r}
# Select sample names
sample_name <- c("MSC49", "MSC68")

```

## Read in raw cellranger matrices 

### Write BPCells Matrices (Filtered)


```{r, message=F, warning=F, eval=T}
## Get filtered matrices file list
all_msc_bp_filt_seu_list_files <- Sys.glob("data/raw/Pf/*/outs/filtered_feature_bc_matrix.h5")

all_msc_bp_filt_seu_list_files
```

```{r, message=F, warning=F, eval=T}
# Write BPCells matrices for filtered feature matrices
all_msc_bp_filt_counts <- write_sample_bpcells(all_msc_bp_filt_seu_list_files, 'data/raw/Pf/|/outs.*h5', 'counts_mtx')
```

### Read BPCells Matrices (Filtered)

```{r, message=F, warning=F, eval=T}
# Read BPCells matrices for filtered data

all_msc_bp_filt <- read_sample_bpcells(list.files("data/processed/Pf", pattern = "^counts_mtx$", full.names = T, recursive = T, include.dirs = T), 'data/processed/Pf/|/bpcells.*')

```


```{r, eval=F}
## List raw cellranger h5 matrices 
list.files("data/raw/Pf/", pattern = "raw_feature_bc_matrix.h5$", full.names = T, recursive = T, include.dirs = T)
```

### Write BPCells Matrices (Raw)

```{r, message=F, warning=F, eval=T}
## Get raw matrices file list
all_msc_bp_raw_seu_list_files <- Sys.glob("data/raw/Pf/*/outs/raw_feature_bc_matrix.h5")

all_msc_bp_raw_seu_list_files
```

```{r, message=F, warning=F, eval=T}
# Write BPCells matrices for filtered feature matrices
all_msc_bp_filt_counts <- write_sample_bpcells(all_msc_bp_raw_seu_list_files, 'data/raw/Pf/|/outs.*h5', 'raw_counts_mtx')
```

### Read BPCells Matrices (Raw)

```{r, message=F, warning=F, eval=T}
# Read BPCells matrices for raw data
all_msc_bp_raw <- read_sample_bpcells(list.files("data/processed/", pattern = "raw_counts_mtx$", full.names = T, recursive = T, include.dirs = T), 'data/processed//Pf/|/bpcells.*')
```


## Gene Counting and Human Gene Removal Function

```{r}
# Function to count PF and human genes, add to metadata, and remove human genes
pf_hs_gene_sum_fn <- function(seurat_obj){
  
  # Identify human and pf genes
  pf_genes <- grep("^PF3D7", rownames(seurat_obj), value = TRUE)
  hs_genes <- grep("^ENSG", rownames(seurat_obj), value = TRUE)
  print(head(hs_genes))
  
  # Extract expression matrix and subset to those genes
  
  pf_matrix <- as(GetAssayData(seurat_obj, layer = "counts")[pf_genes, , drop = FALSE], Class = "dgCMatrix")  
  hs_matrix <- as(GetAssayData(seurat_obj, layer = "counts")[hs_genes, , drop = FALSE], Class = "dgCMatrix")  
  
  # sum the umis and genes across all genes per cell
  pf_umi_count <- colSums(pf_matrix)
  hs_umi_count <- colSums(hs_matrix)
  pf_gn_count <- colSums(pf_matrix > 0)
  hs_gn_count <- colSums(hs_matrix > 0)
  
  # Store in metadata
  seurat_obj[["pf_umi_count"]] <- pf_umi_count
  seurat_obj[["hs_umi_count"]] <- hs_umi_count
  seurat_obj[["pf_gn_count"]] <- pf_gn_count
  seurat_obj[["hs_gn_count"]] <- hs_gn_count
  
  # Keep only pf genes
  seurat_obj <- subset(seurat_obj, features = pf_genes)
  
  return(seurat_obj)
}
```

### Process Filtered and Raw Sets

#### Create a folder to store the seurat objects
```{r, eval=T}
##write out for normalization processing in farm

for (i in 1:length(sample_name)) {
  ## Create new directories
  system(paste0('mkdir -p data/processed/Pf/', sample_name[i], '/seu_obj'))
  # system(paste0('rm data/processed/Pf/', sample_name[i], '/seu_obj/*'))
}


```

```{r}
## Count the pf and human genes and add to metadata for filtered set and also remove human genes
all_msc_bp_filt_noHsGns <- map(all_msc_bp_filt, ~pf_hs_gene_sum_fn(seurat_obj = .x))
imap(all_msc_bp_filt_noHsGns, ~saveRDS(.x, paste0("data/processed/Pf/",.y,"/seu_obj/all_msc_bp_filt_noHsGns.RDS")))

```

```{r}
# Count the pf and human genes and add to metadata for raw set and also remove human genes
all_msc_bp_raw_noHsGns <- map(all_msc_bp_raw, ~pf_hs_gene_sum_fn(seurat_obj = .x))
imap(all_msc_bp_raw_noHsGns, ~saveRDS(.x, paste0("data/processed/Pf/",.y,"/seu_obj/all_msc_bp_raw_noHsGns.RDS")))

```

##### Save Metadata

```{r}
# Save metadata for filtered set
all_msc_bp_filt_noHsGns_mdata <- map(all_msc_bp_filt_noHsGns, ~.x@meta.data) %>% .[sample_name]
imap(all_msc_bp_filt_noHsGns_mdata, ~saveRDS(.x, paste0("data/processed/Pf/",.y,"/seu_obj/all_msc_bp_filt_noHsGns_mdata.RDS")))

```

```{r}
# Save metadata for raw set
all_msc_bp_raw_noHsGns_mdata <- map(all_msc_bp_raw_noHsGns, ~.x@meta.data) %>% .[sample_name]
# saveRDS(all_msc_bp_raw_noHsGns_mdata, "data/processed/Pf/m2_all/all_msc_bp_raw_noHsGns_mdata.RDS")
imap(all_msc_bp_raw_noHsGns_mdata, ~saveRDS(.x, paste0("data/processed/Pf/",.y,"/seu_obj/all_msc_bp_raw_noHsGns_mdata.RDS")))

```


## Retain only the cellbender called cells from the full raw matrix

```{r, message=F, warning=F, eval=T}
## Get cellbender mdata file list
cbender_pf100_mdata_list_files <- Sys.glob("data/raw/Pf/*/cbender/cb_mdata.csv")
cbender_pf100_mdata_list_files
```

```{r, message=F, warning=F, eval=T}
## read filtered barcodes
cbender_pf100_mdata <- load_named_tables(cbender_pf100_mdata_list_files, 'data/raw/Pf/|/cb.*', read_csv) %>%
  .[!is.na(.)]

```

### Compare cellbender with cellranger

```{r, message=F, warning=F, eval=T}
bcodes_dirs <- Sys.glob("data/raw/Pf/*/outs/filtered_feature_bc_matrix/barcodes.tsv.gz")
bcodes_dirs

```


```{r, message=F, warning=F, eval=T}
## read filtered barcodes
cranger_bcodes <- load_named_tables(bcodes_dirs, 'data/raw/Pf/|/outs/filtered.*', ~read_tsv(.x, col_names = "bcode"))
```


```{r, message=F, warning=F, eval=T}
cranger_bcodes <- map(cranger_bcodes, ~.x %>% mutate(is_cell_cr = "Cr"))

```


```{r}
## Create a column summarising the agreement of call methods for each cell - called 'mthds_union'
cr_cbender_mdata_uni <- map2(cbender_pf100_mdata, cranger_bcodes[names(cbender_pf100_mdata)], ~.x %>%
                               full_join(., .y, by = c("barcode"="bcode")) %>%
                               mutate(cr = case_when(is_cell_cr == "Cr" ~ "Cell"),
                                      cb = case_when(cell_probability > 0.5  ~ "Cell", cell_probability <= 0.5  ~ "nonCell"),
                                      is_cell_cb = case_when(cell_probability > 0.5  ~ "Cb", cell_probability <= 0.5  ~ "nonCell"),
                                      mthds_union = paste0(is_cell_cr, is_cell_cb)) %>%
                               mutate(across(starts_with("mthds_union"), ~case_when(
                                 str_detect(., "Cb|Cr") ~ str_remove_all(., "nonCell|NA"),
                                 !str_detect(., "Cb|Cr") ~ str_remove_all(., "NA"),
                                 TRUE ~ .
                               ))) 
)


```



```{r}
map(cr_cbender_mdata_uni, ~.x %>% count(is_cell_cr, is_cell_cb, mthds_union))
```


```{r}
## Convert mthds_union to factor
## Get all possible combinations of factors factoring in the order in which the variables were introduced
c_mthds <- c("Cr", "Cb")

c_combos <- map(1:length(c_mthds), ~ combn(c_mthds, .x, paste0, collapse = "_")) %>% unlist() %>% dput()
c_combos <- c(rev(c_combos), "nonCell")
c_combos
set.seed(12)

## get colors for the cell_call categories
c_combos_cols <- sample(c('#e6194b', '#4363d8', '#46f0f0', '#3cb44b', '#911eb4', '#f58231'), length(c_combos)) %>% set_names(c_combos)

c_combos_cols <- ifelse(names(c_combos_cols) == "nonCell", "lightgrey", c_combos_cols)  %>% set_names(c_combos)

```


```{r}
## Add umi count metadata and modify the cell call categories appropriately
cr_cbender_mdata_slct_uni <- map2(cr_cbender_mdata_uni, all_msc_bp_raw_noHsGns_mdata[names(cr_cbender_mdata_uni)], ~.x %>%
                                    mutate(across(contains("mthds_union"), ~factor(str_replace_all(., c("CrCb" = "Cr_Cb")), levels = c_combos))) %>%
                                    left_join(., .y %>% select(contains("pf_"), contains("hs_")) %>% rownames_to_column("barcode"), by = "barcode")
)

```


```{r}
map(cr_cbender_mdata_slct_uni, ~.x %>% count(is_cell_cr, is_cell_cb, mthds_union))

```


```{r}
map(cr_cbender_mdata_slct_uni, ~.x %>% count(mthds_union, is.na(pf_umi_count)))
```


```{r, fig.width=12, fig.height=5}
call_methods_overlap_barplot_fn(cr_cbender_mdata_slct_uni, "mthds_union", pf_umi_filt = 0, min_grp_size = 0)
```

```{r, fig.width=12, fig.height=5}
call_methods_overlap_barplot_fn(cr_cbender_mdata_slct_uni, "mthds_union", pf_umi_filt = 100, min_grp_size = 0)
```


```{r}
## Create the final cell call category to go forward with

cr_cbender_mdata_slct_uni_pf100 <- map2(cr_cbender_mdata_slct_uni, cranger_bcodes[names(cr_cbender_mdata_slct_uni)], ~.x %>%
                                          mutate(fnl_call = case_when(mthds_union %in%  c("Cr_Cb", "Cb") ~ "Cell",
                                                                      mthds_union == "Cr" ~ "other_calls",
                                                                      mthds_union == "nonCell" ~ "nonCell",
                                                                      T ~ "missing") ) %>%
                                          filter(pf_umi_count >= 100)
)


```

```{r}
map(cr_cbender_mdata_slct_uni_pf100, ~.x %>% count(is_cell_cr, is_cell_cb, mthds_union, fnl_call))

```

######################################START - Write souporcell barcodes################################

```{r}

resc_bcodes_hspf <- map(cr_cbender_mdata_slct_uni_pf100, ~{
  .x  %>%
    filter(fnl_call %in%  c("Cell")) %>% 
    pull(barcode) 
})  

```

```{r}
map(resc_bcodes_hspf, length) %>% bind_rows(., .id = "donor")
```


```{r, message=F, warning=F, eval=T}
# gogo()
```
Write out barcodes to rerun souporcell including all possible barcode that are potentially actual cells

```{r}

imap(resc_bcodes_hspf, ~{
  system(paste0('mkdir data/processed/Pf/', .y, '/soupc_resc_hspf'))
  
  .x %>% write(., gzfile(paste0('data/processed/Pf/', .y, '/soupc_resc_hspf/barcodes.tsv.gz')))
}) 
```

Run souporcell like below for the filtered barcodes of pf in mixed infections

```{r, eval=T}
# gogo()
# bsub < /lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/Talleh/scripts_nbooks/soupc_v25_m2_bsub.sh
# --bam "/lustre/scratch126/tol/teams/lawniczak/projects/malaria_single_cell/mali_field_runs/2022/data/cellranger_runs/Pf_Talleh/5736STDY*/outs/possorted_genome_bam.bam" \
# --bcodes "/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/Talleh/data/processed/Pf/MSC*/soupc_resc_hspf/barcodes.tsv.gz" \
# --soup_dir "soupc_resc_hspf" \
```


######################################END - Write souporcell barcodes################################

```{r, message=FALSE}
# Add cellbender metadata
all_bp_raw_w_cbendr_mdata <- list(all_msc_bp_raw_noHsGns, cr_cbender_mdata_slct_uni_pf100[names(all_msc_bp_raw_noHsGns)]) %>%
  pmap(~ {
    ..2 %>%
      column_to_rownames("barcode") %>%
      select(starts_with("cell_"), contains("background_fraction"), contains("droplet_efficiency"), mthds_union, fnl_call) %>%
      AddMetaData(..1, .)
  })

```



```{r}
all_bp_CbendrCr <- map(all_bp_raw_w_cbendr_mdata, ~subset(.x, subset = fnl_call == "Cell"))
```


```{r}
all_bp_CbendrCr 
```

Remove cells with less than 100 pf reads
```{r}
map(all_bp_CbendrCr, ~{table(.x@meta.data$pf_umi_count >= 100)})
```


```{r, eval=T}
cell_counts <- pmap(list(all_msc_bp_raw_noHsGns, all_msc_bp_filt_noHsGns, all_bp_CbendrCr), ~{
  data.frame("raw" = dim(..1)[2], "cr_filt" = dim(..2)[2], "cbendr" = dim(..3)[2])
  
}) %>%
  bind_rows(., .id = "donor")

```


```{r, eval=T}
cell_counts
```

```{r, eval=T}
## write out seurat bpcells object files for normalization processing in farm
### cellbender
imap(all_bp_CbendrCr, ~saveRDS(.x, paste0("data/processed/Pf/",.y,"/seu_obj/bp_CbendrCr.RDS")))
all_bp_CbendrCr

```


### write out seurat dense matrix since this may be faster and more transparent. bpcells is opaque and may be prone to some weird errors
```{r, eval=T}
## convert to dense
all_bp_CbendrCr_dense <- all_bp_CbendrCr

for (nm in names(all_bp_CbendrCr_dense)) {
  all_bp_CbendrCr_dense[[nm]][["RNA"]]$counts <- as(all_bp_CbendrCr_dense[[nm]][["RNA"]]$counts, Class = "dgCMatrix")
}

```

```{r, eval=T}
## save
imap(all_bp_CbendrCr_dense, ~saveRDS(.x, paste0("data/processed/Pf/",.y,"/seu_obj/dense_CbendrCr.RDS")))

```

