---
title: "MSC QC and preprocessing"
date: '2022-07-11'
author: 'Jesse Rop'
output:
  html_document:
    toc: true
    theme: united
    code_folding: hide
    df_print: paged
---

**Purpose:** This notebook performs preliminary stage annotation and QC for MSC samples. It loads packages, reads annotation outputs from SingleR and scmap, compares annotations, and writes stage labels for downstream pseudobulk genotyping. Keep code chunks unchanged.

```{r}
suppressPackageStartupMessages({
  library(Seurat)
  library(dplyr) 
  library(stringr)
  library(purrr)
  library(tibble)
  library(ggplot2)
  library(tidyr)
  library(conflicted)
  library(SingleCellExperiment)
  library(SingleR)
  library(ggsankey)
  library(scmap)
  library(scuttle)
  
  conflict_prefer("select", "dplyr")
  conflict_prefer("filter", "dplyr")
  conflict_prefer("rename", "dplyr")
  conflict_prefer("map", "purrr")
  conflict_prefer("count", "dplyr")
  conflicts_prefer(base::saveRDS)
  conflicts_prefer(SingleCellExperiment::colData)
  conflicts_prefer(base::as.data.frame)
  conflicts_prefer(base::intersect)
  conflicts_prefer(dplyr::desc)
  conflicts_prefer(dplyr::count)
})
```



# MSC stage annotation and gene expression QC

## Setup and read inputs

Purpose: load required libraries and set knitr options for reproducible rendering.

```{r, setup}
knitr::opts_knit$set(root.dir = '/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/Talleh')

```

#### Source common functions and variables
```{r}
source("~/Talleh_code/pf_common_vars.R")
source("~/Talleh_code/project_common_fns.R")

```

## Read in sample name details

```{r}
# Select sample names
sample_name <- c("MSC49", "MSC68")

```

## Read in seurat objects
```{r, message=F, warning=F, eval=T}
## Get mdata file list
all_seu_list_files <- Sys.glob("data/processed/Pf/*/seu_obj/dense_CbendrCr.RDS")
all_seu_list_files
```

```{r, message=F, warning=F, eval=T}
## read filtered barcodes
all_seu <- map(all_seu_list_files, ~.x %>% 
                 set_names(nm = (str_remove_all(.,'data/processed/Pf/|/seu_obj/dense.*RDS')))) %>% 
  unlist(., recursive = F) %>% 
  .[sample_name] %>%
  map(readRDS)


```

## Stage Annotation

```{r, message=FALSE, warning=FALSE}
# Create abbreviated color map for plotting (short names used in figures)
sp_col_new_abrv <- sp_col_new
names(sp_col_new_abrv) <- str_replace_all(names(sp_col_new), c("discrepantFM" = "dFM", "weak_score" = "WS", "Early " = "E", "Late " = "L", "ring" = "R", "trophozoite" = "T", "Female" = "F", "Male" = "M", " \\(LE\\)" = "LE", "Gametocyte \\(developing\\)" = "Gd"))

sp_col_new_abrv
```


## SingleR and scmap preliminary annotation
Purpose: run reads of SingleR and scmap outputs, generate preliminary stage labels for cells, and produce diagnostic plots.


```{r}
# bsub < ~Talleh_code/submsn_seu_norm_CbEdCr_singleR.sh

```

Read in cell annotation singleR processed in farm
```{r, message=F, warning=F}
##Read singleR annotations
all_pref_pred <- list.files("data/processed/Pf", pattern = "dense.*_(CbendrCr)_pred_sR.RDS$", full.names = T, recursive = T, include.dirs = T) %>%
  set_names(nm = (str_remove_all(.,'data/processed/Pf/|/seu_obj/|bpc_seu|bp|dense.*'))) %>% 
  .[sample_name] %>%
  .[!is.na(.)] %>%
  map(readRDS)

```


```{r}
# Show label counts for the first two SingleR prediction objects
map(all_pref_pred[1:2], ~table(.x$labels))
```


```{r, eval=T}
# Plot delta distribution for the first SingleR result (disabled)
map(all_pref_pred[1], ~plotDeltaDistribution(.x))

```


### Evaluate SingleR manually

Purpose: inspect SingleR scoring and resolve ambiguous male/female calls.
```{r, message=FALSE}
## Get predictions for stages
all_pref_pred_df <- map(all_pref_pred, ~.x[, c("scores", "labels", "pruned.labels", "reference")] %>% 
                          as.data.frame() %>% 
                          rename_with(~str_replace_all(., c("\\.\\."="_", "\\."="_", "_$"="")), contains("score")) %>%
                          rownames_to_column("bcode"))
```


```{r, message=FALSE, warning=FALSE}
all_pref_pred_df <- map2(all_pref_pred_df, all_pref_pred, ~{
  df <- .x %>% 
    mutate(sr_score_ref = paste0("scores_", names(.y@listData$orig.results)[reference], "_scores"),
           sr_labl_ref = paste0("scores_", names(.y@listData$orig.results)[reference], "_labels")) 
  
  
  df <- df %>% 
    mutate(actl_score = purrr::map2_dbl(row_number(), sr_score_ref, ~df[.x, .y]),
           actl_labl = purrr::map2_chr(row_number(), sr_labl_ref, ~df[.x, .y]),
           MaleFem_consistency = case_when(if_any(ends_with("_labels"), ~str_detect(., "Female")) & if_any(ends_with("_labels"), ~str_detect(., "Male")) ~ "MaleFem",
                                           T ~ labels))
  
  
  return(df)
  
  
})

```

```{r, message=FALSE, warning=FALSE}

map(all_pref_pred_df, ~.x %>% count(labels, actl_labl, MaleFem_consistency) %>% filter(MaleFem_consistency == "MaleFem"))
```


```{r, message=FALSE, warning=FALSE}
# Derive final SingleR stage labels and coarse groupings
# This transformation processes the predictions from SingleR to create a final dataframe with stage labels.
all_pref_pred_df_fnl <- map(all_pref_pred_df, ~.x %>% 
                              select("bcode", "labels", "pruned.labels", "reference", "actl_labl", "actl_score", "MaleFem_consistency") %>% 
                              mutate(fine_stg = case_when(is.na(pruned.labels) ~ "weak_score",
                                                          MaleFem_consistency == "MaleFem" ~ "discrepantFM",
                                                          T ~ pruned.labels),
                                     stage_afm = case_when(str_detect(fine_stg, "ring|troph|schizont|developing") ~ "Asexual",
                                                           str_detect(fine_stg, "Female") ~ "Female",
                                                           str_detect(fine_stg, "Male") ~ "Male",
                                                           T ~ fine_stg),
                                     stage_ag = ifelse(stage_afm %in% c("Female", "Male"), "Gametocyte", stage_afm )) %>%
                              rename_with(~paste0("sR_", .), .cols = -c(bcode)))

```


```{r, message=FALSE, warning=FALSE}
# Create abbreviated color map for plotting (short names used in figures)
sp_col_new_abrv <- sp_col_new
names(sp_col_new_abrv) <- str_replace_all(names(sp_col_new), c("discrepantFM" = "dFM", "weak_score" = "WS", "Early " = "E", "Late " = "L", "ring" = "R", "trophozoite" = "T", "Female" = "F", "Male" = "M", " \\(LE\\)" = "LE", "Gametocyte \\(developing\\)" = "Gd"))

sp_col_new_abrv
```



##################################################  END - SingleR use markers to id cells ##################################################

## Scmap preliminary annotation

Purpose: process scmap multi-reference outputs and harmonize preliminary stage labels.


```{r}
## !!! NOTE - scmap ran here
## Assigning stages

#  W (number of top assigned labels that need to be identical)
W <- 3

# COS_SIM (cosine similarity threshhold needed for all top 3 cells to pass for stage labelled to be assigned)
COS_SIM <- 0.4

## Read in parameters for the first (MCA) reference/atlas object
SCE_REF <- "annotation_references/v3labmscs.ref.sce.stgasx.RDS"
STG_LABL <- "stageHL_ref"
STG_ANOT <- "stages_mca"

## Read in parameters for the second (Lasse) reference/atlas object
SCE_REF2 <- "annotation_references/lasse_natural_infcns_sce.RDS"
STG_LABL2 <- "pseudotime.stages"
STG_ANOT2 <- "stages_lasse"

stg_anot_labls <-c(STG_ANOT, STG_ANOT2)


```

```{r}

# Read in the reference datasets and perform preprocessing
sce_ref_list <- map2(list(SCE_REF, SCE_REF2), list(STG_LABL, STG_LABL2),  ~{
  ##Read MCA V3 + field blood stage reference dataset
  sce_ref <- readRDS(.x)
  
  ## Log transform counts using SCE formula
  sce_ref <- logNormCounts(sce_ref, assay.type = "counts", transform = "log")
  
  ## Add gene names to relevant slot. Needed by scmap
  rowData(sce_ref)$feature_symbol <- rownames(sce_ref)
  
  ## Add the column to use as stage labels
  sce_ref@colData[,'ref_stage'] <- sce_ref@colData[,.y]
  
  
  ##scmap feature selection
  sce_ref <- selectFeatures(sce_ref, suppress_plot = T)
  
  ##Cell indexing
  set.seed(155)
  sce_ref <- indexCell(sce_ref)
  
})

```

```{r}
bpc_seu_prelim_anotd <- map(all_seu, ~{
  ## Read in seurat object
  all_msc <- .x
  
  ### Preprocessing MSC query datasets
  
  ## Convert seurat object into single cell experiment filling in the slots appropriately
  msc_sce <- SingleCellExperiment(assays = list(counts = as(object = all_msc[["RNA"]]$counts, Class = "dgCMatrix")), colData = all_msc@meta.data)
  
  # logcounts(msc_sce) <- log2((SingleCellExperiment::counts(msc_sce) + 1))
  msc_sce <- logNormCounts(msc_sce, assay.type = "counts", transform = "log")
  
  rowData(msc_sce)$feature_symbol <- rownames(msc_sce)
  
  msc_sce <- msc_sce[!duplicated(rownames(msc_sce)), ]
  
  ### Stage label assignment (projection)
  # transferring labels from  reference datasets
  
  ## Run scmap
  
  msc_jc <-  scmapCell(msc_sce, list(
    metadata(sce_ref_list[[1]])$scmap_cell_index,
    metadata(sce_ref_list[[2]])$scmap_cell_index
  ) %>% set_names(stg_anot_labls)
  )
  
  ##Projection
  # Assigning stage labels only when the top 3 most closest reference cells to a query cell all have a similar stage label and a cosine similarity greater than 0.4
  msc_lbls <- scmapCell2Cluster(
    msc_jc, 
    list(
      SummarizedExperiment::colData(sce_ref_list[[1]])$ref_stage,
      SummarizedExperiment::colData(sce_ref_list[[2]])$ref_stage
    ), 
    w = as.numeric(W), threshold = as.numeric(COS_SIM))
  
  
  # Obtaining the reference cell that is closest to each query cell in neighbourhood space regardless of cosine similarity. For those where the top 10 most closest cells constitute several stages, we also obtain the second most similar stage.
  ## Get the top V3 cell and top 3 similarity scores that is most similar to each MSC14 cell 
  
  sc_topv3_cells_sims <- pmap(list(stg_anot_labls, sce_ref_list), ~{
    stg_lab <- ..1
    ref_obj <- ..2
    msc_jc[[stg_lab]]$similarities %>% 
      .[1:3,] %>% 
      t() %>% 
      data.frame() %>% 
      rename_all(., ~str_replace(., 'X', 'sim_lab')) %>% 
      rownames_to_column('bcode') %>%
      left_join(., msc_jc[[stg_lab]]$cells %>% 
                  .[1:3,] %>% t() %>% 
                  data.frame() %>% 
                  rename_all(., ~str_replace(., 'X', 'msc_ref_cells')) %>% 
                  rownames_to_column('bcode'),
                by = 'bcode') %>% 
      # mutate(across(starts_with('msc_ref_cells'), ~colnames(msc_combi.sce)[.]))
      mutate(across(starts_with('msc_ref_cells'), ~colnames(ref_obj)[.]))
  })
  
  
  # Adding annotations to MSC query datasets
  
  ## Adding annotations to query cells
  stg_anots <- map(sce_ref_list, ~.x@colData[,c('ref_stage'), drop=F] %>% data.frame())
  
  ## scmap merge the top labels and similarity scores
  top_stg_sim <- map2(sc_topv3_cells_sims, stg_anots, ~.x %>% 
                        left_join(.y %>% rownames_to_column('msc_ref_cells1') %>% rename('stage_lab1' = 'ref_stage'), by = 'msc_ref_cells1') %>%
                        left_join(.y %>% rownames_to_column('msc_ref_cells2') %>% rename('stage_lab2' = 'ref_stage'), by = 'msc_ref_cells2') %>%
                        left_join(.y %>% rownames_to_column('msc_ref_cells3') %>% rename('stage_lab3' = 'ref_stage'), by = 'msc_ref_cells3') %>% 
                        column_to_rownames('bcode')
  )
  
  
  # Put the annotations together and write out
  msc_anots <- pmap(list(stg_anot_labls, top_stg_sim), ~{
    list(colnames(msc_jc[[..1]]$cells),
         msc_lbls$scmap_cluster_labs[,..1],
         msc_lbls$combined_labs) %>%
      set_names(c("cell_bc", paste0(..1, "_lab"), "combined_labs")) %>%
      data.frame()  %>% 
      full_join(., ..2 %>% rownames_to_column("cell_bc"), by = "cell_bc")
  }) %>% 
    set_names(stg_anot_labls) %>%
    bind_rows(., .id = "ref")
  
  return(msc_anots)
  
})

# saveRDS(msc_anots, paste0(O_FILE,"/msc_anots.RDS"))
# saveRDS(msc_anots, O_FILE)


```



```{r}
# Apply scmap reference-specific label refinement rules based on similarity scores
multi_ref_stg_tbl1 <- map(bpc_seu_prelim_anotd, ~{
  obj <- .x
  map(c("stages_mca", "stages_lasse"), ~{
    obj %>% 
      filter(ref == .x ) %>%
      mutate(!!rlang::sym(paste0(.x, "_lf")) := case_when(
        !!rlang::sym(paste0(.x, "_lab")) == "unassigned" & if_any(starts_with("sim_"), ~.x > 0.3) & if_all(starts_with("stage_"), ~str_detect(.x, "ring|rophozoite|eveloping|Ring|chizont")) ~ stage_lab1,
        !!rlang::sym(paste0(.x, "_lab")) == "unassigned" & if_any(starts_with("sim_"), ~.x > 0.4) & if_all(starts_with("stage_"), ~str_detect(.x, "Male| male|\\(male|eveloping")) ~ stage_lab1,
        !!rlang::sym(paste0(.x, "_lab")) == "unassigned" & if_any(starts_with("sim_"), ~.x > 0.4) & if_all(starts_with("stage_"), ~str_detect(.x, "Female|female|eveloping")) ~ stage_lab1,
        TRUE ~ !!rlang::sym(paste0(.x, "_lab"))
      )) 
  })
})

```


```{r}
map(multi_ref_stg_tbl1, ~.x[[1]] %>% count(stages_mca_lf, combined_labs))

```



```{r}
# Harmonize stage labels across scmap references and apply consensus logic
multi_ref_stg_tbl2 <- map(multi_ref_stg_tbl1, ~{
  tbl <- .x
  
  map(tbl, ~.x %>% select(any_of(c('cell_bc', 'stages_mca_lf', 'stages_lasse_lf', 'combined_labs')), contains("sim_"), contains("stage_"))) %>% 
    purrr::reduce(., ~full_join(.x, .y, by = c("cell_bc", "combined_labs"), suffix = c("_mca", "_lasse"))) %>%
    mutate(
      stage_final = case_when(
        combined_labs == "unassigned" & if_any(c('stages_mca_lf', 'stages_lasse_lf'), ~str_detect(.x, "ring|rophozoite|eveloping|Ring|chizont")) ~ stages_lasse_lf, 
        combined_labs == "unassigned" & if_any(c('stages_mca_lf', 'stages_lasse_lf'), ~str_detect(.x, "Male| male|\\(male|eveloping")) ~ stages_mca_lf,  
        combined_labs == "unassigned" & if_any(c('stages_mca_lf', 'stages_lasse_lf'), ~str_detect(.x, "Female|female|eveloping")) ~ stages_mca_lf, 
        T ~ combined_labs),
      stage_final2 = case_when(
        stage_final == "unassigned" & if_any(c('stages_mca_lf', 'stages_lasse_lf'), ~str_detect(.x, "ring|rophozoite|eveloping|Ring|chizont")) ~ stages_mca_lf, 
        T ~ stage_final)
    )
})

```

```{r}
# Count cells with unassigned stage labels after harmonization
map(multi_ref_stg_tbl2, ~.x %>% filter(stage_final2 == "unassigned") %>% dim() %>% .[1]) %>% data.frame() %>% t()

```

```{r}
# Format final stage labels: standardize names, remove abbreviated placeholders, and derive coarse stage groupings
multi_ref_stg_tbl2_slct <- map(multi_ref_stg_tbl2, ~ .x %>%
                                 column_to_rownames("cell_bc") %>%
                                 select(combined_labs, stage_final, stage_final2, sim_lab1_mca) %>% 
                                 mutate(across(c(stage_final2), ~
                                                 str_replace_all(., c("Trophozoite" = " trophozoite", 
                                                                      "Ring" = " ring",
                                                                      "Schizont" = " schizont", 
                                                                      "early" = "Early",
                                                                      "late" = "Late",
                                                                      "DevelopingGametocyte" = "Gametocyte (developing)",
                                                                      "aleGametocyte" = "ale",
                                                                      "gametocyte" = "Gametocyte",
                                                                      "LE" = "(LE)",
                                                                      "unassigned" = "Unassigned",
                                                                      "Gametocyte \\(male\\)" = "Male",
                                                                      "Gametocyte \\(female\\)" = "Female")))
                                 ) %>% 
                                 mutate(stage_fld = str_replace_all(stage_final2, c("Mid trophozoite" = "Early trophozoite", ".*schizont" = "Schizont")),
                                        stage_afm = case_when(str_detect(stage_fld, "ring|rophozoite|chizont|eveloping") ~ "Asexual",
                                                              str_detect(stage_fld, "Female") ~ "Female",
                                                              str_detect(stage_fld, "Male") ~ "Male",
                                                              T ~ stage_fld),
                                        stage_ag = ifelse(stage_afm %in% c("Female", "Male"), "Gametocyte", stage_afm ))
)

```

################################################## END - Scmap preliminary annoation ##############################################

## Compare scmap and SingleR

Purpose: compare stage labels from scmap and SingleR, summarize agreement, and create sankey/heatmap diagnostics.

```{r, message=FALSE, warning=FALSE}

all_pref_pred_df_fnl_cbender <- all_pref_pred_df_fnl

```

```{r, message=FALSE, warning=FALSE}

singleR_scmap_df_snky <- map2(all_pref_pred_df_fnl_cbender, multi_ref_stg_tbl2_slct[names(all_pref_pred_df_fnl_cbender)], ~.x %>% 
                                # rename("stage_final2" = sR_fine_stg) %>%
                                select(sR_fine_stg, sR_stage_afm, bcode) %>%
                                left_join(., .y %>% 
                                            rownames_to_column("bcode") %>% 
                                            select(stage_fld, stage_afm, bcode), 
                                          by = "bcode")  %>%  
                                mutate(across(c(stage_fld, sR_fine_stg), ~str_replace_all(., c("discrepantFM" = "dFM", "weak_score" = "WS", "Early " = "E", "Late " = "L", "ring" = "R", "trophozoite" = "T", "Schizont" = "S", "Female" = "F", "Male" = "M", " \\(LE\\)" = "LE", "Gametocyte \\(developing\\)" = "Gd")))) 
)


```


```{r, fig.width=4, fig.height=4}
# Define sankey plot helper used to compare label assignments between methods
snk_pt_fn <- function(decod_tbl, strns_2_plt = c('Strain_2nd', 'n_strains'), labls = paste0("K=", opt_clusters_nms[[3]][1:2]), bp_face = c("bold", "plain"), ttl = "plot", strain_colours = strain_cols_n) {
  decod_tbl %>% 
    mutate(across(contains("train"), ~str_replace_all(., c("Doublet" = "Dbt", "Negative" = "Neg")))) %>%
    make_long(all_of(strns_2_plt)) %>% 
    add_count(x,node, "n") %>%
    ggplot(., aes(x = x, next_x = next_x,node = node,next_node = next_node,fill = factor(node),label = paste0(node,": ",n))) +
    geom_sankey() +
    theme_void()+
    scale_fill_manual(values = strain_colours) +
    scale_x_discrete(labels=labls)+
    geom_sankey_label(size = 3.6, color = 1, fill = "white") +
    labs(title = ttl) +
    theme_sankey(base_size = 9) +
    theme(legend.position = "none",
          axis.title.x = element_blank(),
          axis.text.x = element_text(size = 15, face = bp_face))
  
}


```

```{r, message=FALSE, warning=FALSE}
# Generate sankey plots comparing SingleR fine stage to scmap field stage
imap(singleR_scmap_df_snky, ~snk_pt_fn(., strns_2_plt = c("sR_fine_stg", "stage_fld"), labls = c("SingleR", "scmap"), bp_face = c("bold", "bold"), ttl = .y, strain_colours = sp_col_new_abrv))

```


```{r, message=FALSE, warning=FALSE}
# Generate sankey plots comparing SingleR coarse stage to scmap coarse stage
imap(singleR_scmap_df_snky, ~snk_pt_fn(., strns_2_plt = c("sR_stage_afm", "stage_afm"), labls = c("SingleR", "scmap"), bp_face = c("bold", "bold"), ttl = .y, strain_colours = sp_col_new_abrv))

```


```{r}
# Summarize agreement between SingleR and scmap as percentage table for heatmap plotting
df_counts <-singleR_scmap_df_snky %>%
  bind_rows(.id = "donor") %>%
  filter(!str_detect(donor, "MSC12")) %>%
  filter(sR_stage_afm %in% c("Asexual", "Female", "Male")) %>%
  filter(stage_afm %in% c("Asexual", "Female", "Male")) %>%
  group_by(donor, sR_stage_afm, stage_afm) %>%
  summarise(Count = n(), .groups="drop") %>% 
  group_by(donor) %>%
  mutate(Pct = Count / sum(Count)) %>%
  mutate(sR_don_stg = str_remove_all(paste0(donor, sR_stage_afm), "SC|sexual|emale|ale|cbendr"), 
         scmap_don_stg = str_remove_all(paste0(donor, stage_afm), "SC|sexual|emale|ale|cbendr")) %>% 
  pivot_wider(id_cols = scmap_don_stg, names_from = c(sR_don_stg), values_from = Pct, values_fill = 0)

```

```{r, fig.width=12, fig.height=10}
# Plot faceted heatmaps
df_counts  %>%
  column_to_rownames("scmap_don_stg") %>%
  t() %>%
  ComplexHeatmap::Heatmap(., cluster_rows = F, cluster_columns = F, show_column_names = T, show_row_names = T)
```


################################################## END - Compare scmap with singleR ##############

## Write out stages for stage-specific pseudobulk genotyping

Purpose: save harmonized stage labels to disk for downstream pseudobulk genotyping and integration with souporcell.
```{r, eval = T}
all_seu_prelim_stg_norm_cb_mdata <- map(multi_ref_stg_tbl2_slct, ~.x %>% rownames_to_column("bcode") %>% select(bcode, stage_fld, stage_afm, stage_ag)) 

species <- "Pf"
sv_dir <- "pbulk_gtypes_preQC_resc"
system(paste0("mkdir -p data/processed/", species, "/" , sv_dir, "/m2_all"))

saveRDS(all_seu_prelim_stg_norm_cb_mdata, paste0("data/processed/", species, "/m2_all/all_seu_prelim_stg_norm_cb_mdata.RDS"))
# saveRDS(all_seu_prelim_stg_norm_cb_mdata, paste0("data/processed/", species, "/" , sv_dir, "/all_seu_prelim_stg_norm_cb_mdata.RDS"))

```



```{r, eval = T}
map(all_seu_prelim_stg_norm_cb_mdata, ~.x %>% count(stage_ag, stage_afm, stage_fld)) 
```

### Add annotations to Seurat object
Purpose: add preliminary stage metadata back to Seurat objects so downstream normalization and analyses retain stage labels.

```{r}
## Adding stage annotations to query cells
all_seu_prelim_stg <- list(all_seu, all_pref_pred_df_fnl[str_remove(names(all_seu), "3$")], multi_ref_stg_tbl2_slct[str_remove(names(all_seu), "3$")]) %>%
  pmap(~ ..2 %>%
         column_to_rownames("bcode") %>%
         AddMetaData(..1, .) %>%
         AddMetaData(., ..3) 
  )

```


```{r}
all_seu_prelim_stg %>% map(., ~.x@meta.data %>% count(stage_final2, combined_labs, sR_fine_stg) )

```

```{r}
all_seu_prelim_stg %>% map(., ~.x@meta.data %>% count(stage_final2, sR_fine_stg))


```

```{r}
all_seu_prelim_stg %>% map(., ~.x@meta.data %>% 
                             count(stage_final2, combined_labs, sR_fine_stg) %>% 
                             filter(str_detect(sR_fine_stg, "LE"))
)

```


```{r, eval=T}
## write out seurat bpcells object files for normalization processing in farm
imap(all_seu_prelim_stg, ~saveRDS(.x, paste0("data/processed/Pf/",.y, "/seu_obj/bp_prelim_stg.RDS")))

```
