---
title: "msc_strain_decon_new"
date: '2023-08-10'
author: 'Jesse Rop'
output:
  html_document:
    toc: true
    theme: united
    code_folding: hide
    df_print: paged
---


```{r, message=FALSE}
suppressPackageStartupMessages({
  library(Seurat)
  library(dplyr) 
  library(stringr)
  library(purrr)
  library(readr)
  library(tibble)
  library(ggplot2)
  library(tidyr)
  library(cowplot)
  library(pheatmap)
  library(conflicted)
  library(ComplexHeatmap)
  library(ggvenn)
  library(ggsankey)
  
  conflict_prefer("select", "dplyr")
  conflict_prefer("count", "dplyr")
  conflict_prefer("filter", "dplyr")
  conflict_prefer("rename", "dplyr")
  conflict_prefer("saveRDS", "base")
  conflicts_prefer(base::as.data.frame)
})

```


# MSC parasite strain deconvolution

## Variable declarations


```{r setup}
knitr::opts_knit$set(root.dir = '/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/Talleh/')
```

### Common functions

#### Source common functions, variables and tables
```{r}
source("~/Talleh_code/pf_common_vars.R")
source("~/Talleh_code/project_common_fns.R")

```

## Specify variables and sample name details

```{r}
# Select sample names
sample_name_nms <- c("MSC49", "MSC68")

```

```{r}
# Variables
sample_set = 'Mali2'
algn_nm = 'minmap'
species = 'Pf'

soup_dir = "soupc_resc_hspf"
sv_dir = "pbulk_gtypes_preQC_resc_hspf"


strain_levels_univrs <- c(paste0("SC", 1:15), "Doublet", "Negative")

## Alignment methods
algn_nms = c('minmap', 'hsat')

## Combine donor ids with alignment methods since each donor is ran once per alignment method
sample_name_nms_al <- rep(sample_name_nms, length(algn_nms))
algn_nms_al <- rep(algn_nms, each = length(sample_name_nms))

sm_aln_nms_al <- paste(rep(sample_name_nms, length(algn_nms)),
                       rep(algn_nms, each = length(sample_name_nms)),
                       sep = '_')

## number of optimum clusters to investigate per sample
opt_clusters_nms <- list(
  "MSC49" = c(4:8),
  "MSC68" = c(2:6)
) %>% set_names(sample_name_nms)


opt_narrow_clusters_nms <-  list(
  "MSC49" = c(4:6),
  "MSC68" = c(3:4)
) %>% set_names(sample_name_nms)

opt_narrow_clusters_nms_aln <- rep(opt_narrow_clusters_nms, length(algn_nms)) %>% set_names(sm_aln_nms_al)

optimum_k = c("MSC49" =5, 
              "MSC68" =3) %>% set_names(sample_name_nms)

optimum_k_aln = rep(optimum_k, length(algn_nms)) %>% set_names(sm_aln_nms_al)


## number of optimum clusters (K) iterations for each donor
opt_clusters_len <- map(opt_clusters_nms, length)

## number of optimum clusters to investigate per sample factoring the number of alignments
opt_clusters_nms_aln <- rep(opt_clusters_nms, length(algn_nms)) %>% set_names(sm_aln_nms_al)
opt_clusters_len_aln <- map(opt_clusters_nms_aln, length)


## Donors factoring the number of alignments
sample_name <- map2(sample_name_nms, opt_clusters_len[sample_name_nms], ~
                      rep(.x, each = .y))

sample_name <- unlist(flatten(sample_name))

opt_clusters <- unlist(flatten(opt_clusters_nms[sample_name_nms]))

## sorted donor_opt_cluster names
sorted_names <- paste0(sample_name, "_", opt_clusters)

## sorted factoring alignment
sorted_names_al <- paste0(sample_name,
                          "_",
                          rep(algn_nms, each = length(sample_name)),
                          "_",
                          opt_clusters)

iter = opt_clusters_len

downsample_cells = "yes"
```



## Log-likelihood elbow plots for optimal k selection

```{r, message=FALSE, warning=FALSE, eval=T}
## Read in log likelihoods for estimation of number of clusters from elbow plots
clusters_log_likelihoods_lst = map2(sample_name_nms_al, rep(algn_nms, each= length(sample_name_nms)), ~read_table(paste0('data/processed/',species,'/', .x,'/', soup_dir, '/',.y,'/clusters_log_likelihoods.txt'), col_names = c('No_of_clusters', 'td1','td2','td3', 'td4','log_likelihood')) %>% select(-(starts_with("td"))) %>% mutate(across(No_of_clusters, ~str_remove_all(., ".*out_dir_k|/clusters.err:best")))) %>% set_names(paste(sample_name_nms_al, rep(algn_nms, each= length(sample_name_nms)), sep ='_'))
```


```{r, eval=T} 
map(clusters_log_likelihoods_lst, ~.x %>%
      mutate(across(c('No_of_clusters'), ~factor(as.numeric(.)))) %>%
      ggplot(., aes(x = No_of_clusters, y = log_likelihood)) +
      geom_point() +
      geom_line(aes(group = 1)) +
      # facet_grid(rows = .~id)+
      theme(axis.text=element_text(size=10),
            axis.title=element_text(size=16,face="bold"),
            legend.text = element_text(size=15),
            title = element_text(size=16,face="bold"),
            strip.text.x = element_text(size = 18, face = "bold")))
```


```{r, fig.width= 20, fig.height=4, eval=T} 
clusters_log_likelihoods_lst %>% 
  bind_rows(., .id = "sample_algnr") %>% 
  # separate_wider_regex(cols = "sample_algnr", c(sample = "MSC[0-9]*|MSC[0-9]*_[A-Z]*|MSC[0-9]*[A-Za-z]*|MSC[0-9]*[A-Za-z]*_[A-Z]*", "_",algnr = "[a-z]*$")) %>%
  separate_wider_regex(cols = "sample_algnr", patterns = c(sample = ".*", "_", algnr  = "minmap$|hsat$")) %>%
  mutate(across(c('sample'), ~factor(., levels = sample_name_nms))) %>%
  mutate(across(c('No_of_clusters'), ~factor(as.numeric(.)))) %>%
  ggplot(., aes(x = No_of_clusters, y = log_likelihood)) +
  geom_point() +
  geom_line(aes(group = 1)) + 
  facet_grid(algnr ~ sample, scales = "free_y")
```

## Souporcell data loading and strain assignment

Load souporcell `clusters.tsv` files for all candidate k-values across all sample-aligner combinations.

```{r, eval=T}
soupc_all_al <- list.files(paste0("data/processed/", species), pattern = "clusters.tsv", full.names = T, recursive = T, include.dirs = T)%>% set_names(nm = (str_remove_all(.,paste0('data/processed/', species, '/|', soup_dir,'/|k|/clusters.tsv'))) %>% str_replace_all(., '/', '_'))  %>% .[sorted_names_al]%>%
  map_df(read_delim, delim = "\t", col_types = 'cccddd', .id = "sample_k")


```

Extract barcode, status, and assignment columns from `clusters.tsv` and convert numeric cluster indices to labelled strain identifiers (SC1, SC2, …, Doublet, Negative).

```{r, eval=T}

soupo_kselect_al <- soupc_all_al %>%
  dplyr::select(sample_k, barcode, status, assignment) %>% 
  dplyr::mutate("Strain" = case_when(
    status == "doublet"    ~ "Doublet",
    status == "unassigned" ~ "Negative",
    TRUE                   ~ paste0("SC", as.integer(assignment) + 1L)
  ))
```




### Compare minimap2 vs HISAT2 strain assignments

Pivot the per-barcode strain labels wide by aligner, producing one column per aligner for each barcode, grouped by sample and k-value.

```{r, eval=T}
soupo_kselect_al_cor <- soupo_kselect_al %>%
  separate_wider_regex(cols = "sample_k", patterns = c(sample = "M.*", "_", aligner  = "minmap|hsat", "_", k = "\\d+")) %>%
  unite(col = "sample_k", c('sample','k'), remove = FALSE)  %>%
  select("sample_k","aligner","barcode","Strain") %>% 
  dplyr::group_split(sample_k) %>%
  purrr::set_names(purrr::map_chr(., ~.x$sample_k[1])) %>%
  .[sorted_names]

hsat_mnmp_comp = map(soupo_kselect_al_cor, ~pivot_wider(.x, names_from = aligner, values_from = Strain) %>%
                       dplyr::select(barcode, all_of(algn_nms)))

```

### Heatmap function

Cross-tabulation heatmap showing per-barcode agreement between two sets of strain labels (rows vs columns), used here to compare minimap2 and HISAT2 assignments at each k-value.

```{r , fig.width = 10, fig.height = 5, warning=FALSE}
phmap_lab_transfr <- function(dset_lab, col_labs, row_labs, row_labs_c, col_labs_c, strn_order, ttl){
  anno_row = dset_lab %>% dplyr::count({{ col_labs }}, {{ row_labs }}) %>%
    pivot_wider(values_from = n, names_from = {{ col_labs }}) %>%
    column_to_rownames(names(.)[1]) %>%
    mutate({{ row_labs }} := row_labs_c) %>%
    dplyr::select({{ row_labs }})
  
  anno_col = dset_lab %>% dplyr::count({{ col_labs }}, {{ row_labs }}) %>%
    pivot_wider(values_from = n, names_from = {{ row_labs }}) %>%
    column_to_rownames(names(.)[1]) %>%
    mutate({{ col_labs }} := col_labs_c) %>%
    dplyr::select({{ col_labs }})
  
  hm_mtx <- dset_lab %>% dplyr::count({{ col_labs }}, {{ row_labs }}) %>%
    arrange({{ col_labs }}, {{ row_labs }}) %>%
    mutate(across(c({{ col_labs }}, {{ row_labs }}), ~droplevels(factor(., levels = strain_levels_univrs)))) %>%
    spread({{ col_labs }},n,fill = 0) %>% ## spread preserves factor-level column ordering; pivot_wider does not
    # pivot_wider(values_from = n, names_from = {{ col_labs }}) %>%
    # replace(is.na(.), 0) %>%
    column_to_rownames(names(.)[1]) #%>%
  # .[,strn_order] %>%
  
  # print(head(hm_mtx))
  
  hm_mtx %>%
    ComplexHeatmap::pheatmap(display_numbers = T,treeheight_row=0, treeheight_col=0, annotation_row = anno_row, annotation_col = anno_col, legend = F, annotation_legend = T, cluster_rows =F, cluster_cols = F, fontsize = 20, annotation_names_row=T, annotation_names_col = T,
                             # color = 'white',
                             number_format = "%.0f", main = ttl#,
                             # annotation_colors = list('minmap' = "green", 'hsat' = "lightblue")
    )
  
}

```




```{r , fig.width = 10, fig.height = 5, warning=FALSE}

imap(hsat_mnmp_comp, possibly(~phmap_lab_transfr(.x, col_labs= `minmap`, row_labs=`hsat`, row_labs_c ='hsat', col_labs_c= 'minmap', ttl = .y)), 'k=1 so no groups to compare which is necessary for the heatmap')

```

```{r , fig.width = 10, fig.height = 5, warning=FALSE}
imap(hsat_mnmp_comp, possibly(~phmap_lab_transfr(.x, col_labs= `minmap`, row_labs=`hsat`, row_labs_c ='minmap', col_labs_c= 'hsat', ttl = .y)), 'k=1 so no groups to compare which is necessary for the heatmap')

```



## Within-aligner strain harmonization across k-values

Split barcode-level strain assignments into a named list (one element per sample-aligner-k combination) ready for harmonization.

```{r}
soupo_kselect_list <- soupo_kselect_al %>%
  dplyr::group_split(sample_k) %>%
  purrr::set_names(purrr::map_chr(., ~.x$sample_k[1])) %>%
  map(. %>% column_to_rownames(., "barcode") %>% dplyr::select(Strain))

soupo_kselect_list <- soupo_kselect_list[as.character(sorted_names_al)]


##Get strain levels for samples with more than one strain and mutate Strain column to reflect this
strain_levels <- map(soupo_kselect_list, ~.x %>% dplyr::filter(str_detect(Strain, '^SC')) %>% pull(Strain) %>% unique() %>% sort() %>% c(.,"Doublet","Negative"))

soupo_kselect_list <- map2(soupo_kselect_list,strain_levels, ~.x %>% mutate(across(Strain, ~factor(., levels = .y)))) 
```



```{r, message=FALSE}
## Merge all k-value strain columns into a single wide-format tibble per sample-aligner, joining on barcode.

soupo_kmerge <- map(sm_aln_nms_al, ~{
  soupo_kselect_list[str_detect(names(soupo_kselect_list), paste0(.x, "_"))] %>% 
    imap(., ~{
      nms = .y
      .x %>% rename_with(~paste0(.,'_', nms), everything()) %>% rownames_to_column('bcode')
    }) %>% 
    purrr::reduce(., left_join, by = 'bcode')
}) %>%
  set_names(sm_aln_nms_al)

soupo_seu_kmerge <- soupo_kmerge
```



### Helper functions

#### `strain_harmonize_fn`: assign a consensus label from two k-value strain columns

For each cell, finds the most frequent pairing between `strn1` and `strn2` and writes the consensus label to `out_strn`. Doublet/Negative cells are handled explicitly; unresolved cells receive a new SC label.

```{r}

strain_harmonize_fn <- function(strn_tbl = tbl1, strn1 = "Strain_2nd", strn2 = "Strain_final", out_strn = "n_strains") {
  
  strn_tbl %>% group_by(!!rlang::sym(strn1)) %>% 
    add_count(!!rlang::sym(strn2), name = "stf_n") %>%
    mutate(across(stf_n, ~case_when(str_detect(!!rlang::sym(strn2), "SC") & str_detect(!!rlang::sym(strn1), "Doublet|Negative") ~ 0,
                                    str_detect(!!rlang::sym(strn1), "SC") & str_detect(!!rlang::sym(strn2), "Doublet|Negative") ~ 0,
                                    str_detect(!!rlang::sym(strn1), "Doublet|Negative") & str_detect(!!rlang::sym(strn2), "Doublet|Negative") ~ 0,
                                    T ~ .)),
           top_st = max(stf_n),
           nw_st = case_when(!!rlang::sym(strn2) == "Doublet" | !!rlang::sym(strn2) == "Negative" ~ !!rlang::sym(strn2),
                             top_st == stf_n & top_st != 0 ~ !!rlang::sym(strn1))) %>% 
    group_by(top_st, nw_st) %>%
    mutate(lab_uniq = row_number(),
           across(lab_uniq, ~case_when(top_st == 0 | nw_st == "Doublet" | nw_st == "Negative" ~ NA, T ~ .))) %>%
    group_by(!!rlang::sym(strn2)) %>%
    mutate(top_st2 = case_when(top_st == stf_n & top_st != 0 & lab_uniq == 1 ~ top_st, T ~ 0), 
           top_top_st = max(top_st2),
           ty_brakr = case_when(!!rlang::sym(strn2) == "Doublet" | !!rlang::sym(strn2) == "Negative" ~ !!rlang::sym(strn2),
                                top_top_st == stf_n & top_top_st != 0 ~ !!rlang::sym(strn1))) %>% 
    arrange(!!rlang::sym(strn2))  %>%
    fill(ty_brakr, .direction = "downup") %>%
    group_by(is.na(ty_brakr))%>%
    mutate(nro_count = data.table::rleid(!!rlang::sym(strn2))) %>% ## run-length ID within is.na(ty_brakr) groups — gives each unresolved strn2 value a unique integer for new SC label assignment
    ungroup()%>%
    mutate(sc_numbr = as.numeric(str_remove_all(!!rlang::sym(strn1), "Doublet|Negative|SC")))%>%
    mutate(!! out_strn:= case_when(is.na(ty_brakr) ~ paste0("SC",(max(sc_numbr, na.rm = T)+nro_count)), T ~ ty_brakr)) %>%
    # select(c("bcode", strn1, out_strn)) %>% 
    select(-c('stf_n', 'top_st',   'ty_brakr', 'is.na(ty_brakr)', 'nro_count')) #%>%
  #column_to_rownames('bcode')
  
}
```


#### `snk_pt_fn`: Sankey plot of strain label transitions across k-values

```{r}

snk_pt_fn <- function(decod_tbl, strns_2_plt = c('Strain_2nd', 'n_strains'), labls = paste0("K=", opt_clusters_nms[[3]][1:2]), bp_face = c("bold", "plain"), ttl = "plot", strain_colours = strain_cols_n) {
  decod_tbl %>% 
    mutate(across(contains("train"), ~str_replace_all(., c("Doublet" = "Dbt", "Negative" = "Neg")))) %>%
    make_long(all_of(strns_2_plt)) %>% 
    add_count(x,node, "n") %>%
    ggplot(., aes(x = x, next_x = next_x,node = node,next_node = next_node,fill = factor(node),label = paste0(node,": ",n))) +
    geom_sankey() +
    theme_void()+
    scale_fill_manual(values = strain_colours) +
    scale_x_discrete(labels=labls)+
    geom_sankey_label(size = 3.6, color = 1, fill = "white") +
    labs(title = ttl) +
    theme_sankey(base_size = 9) +
    theme(legend.position = "none",
          axis.title.x = element_blank(),
          axis.text.x = element_text(size = 15, face = bp_face))
  
}

```


### Harmonize strain labels within each aligner

Build column-name vectors for each sample-aligner-k combination, then chain pairwise `strain_harmonize_fn` calls across consecutive k-values (lowest k → next k → … → highest k), producing one harmonized label per cell per aligner.

```{r}
don_k_names_aln <- imap(opt_clusters_nms_aln, ~paste0('Strain_',.y, '_', .x))
don_hmonizd_k_names_aln <- imap(opt_clusters_nms_aln, ~paste0('Strain_hmo_',str_remove(.y, "SC"), '_', .x))

## Initial two-step harmonization (k_low → k_mid → k_high) used to derive the full iterative function below
strn_harmnzd_tbls <- list(soupo_seu_kmerge, don_k_names_aln, don_hmonizd_k_names_aln) %>%
  pmap(~{
    
    strain_harmonize_fn(strn_tbl = ..1, strn1 = ..2[1], strn2 = ..2[2], out_strn = ..3[2]) %>%
      strain_harmonize_fn(strn_tbl = ., strn1 = ..3[2], strn2 = ..2[3], out_strn = ..3[3])
    
  })

list(strn_harmnzd_tbls, don_k_names_aln, opt_clusters_nms_aln, don_hmonizd_k_names_aln) %>%
  pmap(~{  
    
    snk_pt_fn(..1, strns_2_plt = c(..2[1], ..4[2:3]), labls = paste0("K=", ..3[1:3]), bp_face = c("bold", "plain", "plain"))
    
  })

```



#### `strn_hm_tbl_fn`: generalised iterative strain harmonization across any number of k-values

Left-folds `strain_harmonize_fn` over all k-values: step 1 uses the raw k1 label as `strn1`; subsequent steps use the harmonized output of the previous step, so the final column reflects the consensus across all k-values.

```{r}

strn_hm_tbl_fn <- function(soupo_seu_kmerge_f, don_k_names_aln_f, don_hmonizd_k_names_aln_f) {
  list(soupo_seu_kmerge_f, don_k_names_aln_f, don_hmonizd_k_names_aln_f) %>%
    pmap(~{
      purrr::reduce(seq_len(length(..2) - 1), function(tbl, i) {
        strain_harmonize_fn(
          strn_tbl = tbl,
          strn1    = if (i == 1) ..2[1] else ..3[i],
          strn2    = ..2[i + 1],
          out_strn = ..3[i + 1]
        )
      }, .init = ..1)
    })
}


```



```{r}

strn_harmnzd_tbls <- strn_hm_tbl_fn(soupo_seu_kmerge_f = soupo_seu_kmerge, don_k_names_aln_f =don_k_names_aln, don_hmonizd_k_names_aln_f = don_hmonizd_k_names_aln)
```

### Sankey plots: within-aligner strain label transitions

```{r, fig.width=6, fig.height=4, eval=T}
snk_plts <- list(strn_harmnzd_tbls, don_k_names_aln, opt_clusters_nms_aln, don_hmonizd_k_names_aln) %>%
  pmap(~{  
    
    snk_pt_fn(..1, strns_2_plt = c(..2[1], ..4[2:length(..4)]), labls = paste0("K=", ..3[1:5]), bp_face = c("plain"))
    
  })

```

```{r, eval=T}
snk_plts[str_detect(names(snk_plts), "minmap")] %>% imap(., ~.x + labs(title = .y))
```

```{r, eval=T}
snk_plts[str_detect(names(snk_plts), "hsat")] %>% imap(., ~.x + labs(title = .y))
```


```{r, fig.height=6, fig.width=24, eval=T}
## Save for visualization of sankey plots for the manuscript
saveRDS(snk_plts, "data/processed/Pf/m2_all/strain_calling/snk_plts.RDS")

```




```{r}
## Alias the lowest-k raw strain column to its harmonized name (so all Strain_hmo_* columns are present), then convert all harmonized columns to ordered factors.

strn_harmnzd_tbls <- list(strn_harmnzd_tbls, map(don_k_names_aln, ~.x[1]), map(don_hmonizd_k_names_aln, ~.x[1])) %>%
  pmap(~{
    ..1 %>% 
      mutate(!! ..3:= !!rlang::sym(..2)) %>%
      mutate(across(contains("Strain_hmo"), ~droplevels(factor(.,levels = strain_levels_univrs))))
    
  })

```


```{r, fig.height=6, fig.width=24, eval=T}
## Save for visualization of sankey plots for the manuscript
saveRDS(strn_harmnzd_tbls, "data/processed/Pf/m2_all/strain_calling/sstrn_harmnzd_tbls.RDS")
saveRDS(don_k_names_aln, "data/processed/Pf/m2_all/strain_calling/don_k_names_aln.RDS")
saveRDS(opt_clusters_nms_aln, "data/processed/Pf/m2_all/strain_calling/opt_clusters_nms_aln.RDS")
saveRDS(don_hmonizd_k_names_aln, "data/processed/Pf/m2_all/strain_calling/don_hmonizd_k_names_aln.RDS")
```

## Consensus strain assignment and QC

Build per-barcode metadata combining strain labels across all k-values and both aligners. `Strain_qc` is the label at the chosen optimal k; `Strain_max_qc` uses the highest-k label. `dbt_neg_all` flags whether all k-value calls are doublet/negative.

```{r, message=FALSE}
## NOTE: set Strain_qc_cutoff > 20 when exploring k-values; use 0 once the optimal k is fixed, to retain all cells.
Strain_qc_cutoff=0

strn_c_mdata <- list(strn_harmnzd_tbls, 
                     names(strn_harmnzd_tbls), 
                     opt_clusters_nms_aln[names(strn_harmnzd_tbls)] %>% map(.,max), 
                     optimum_k_aln, 
                     imap(opt_narrow_clusters_nms_aln, ~paste0(.y, '_', .x))) %>%
  pmap(~{
    ..1 %>%
      unite(col = 'Strn_c', starts_with("Strain_hmo_M"), sep='_', remove = F) %>% 
      add_count(Strn_c) %>% 
      mutate(
        'Strain_max_qc' = case_when( 
          !str_detect(Strn_c, 'Negative|Doublet') & n>Strain_qc_cutoff ~ !!rlang::sym(paste0('Strain_hmo_', str_remove(..2, 'SC'), '_',..3)), 
          T ~ 'PoorQC'),
        'Strain_qc' = case_when(
          !str_detect(!!rlang::sym(paste0('Strain_hmo_', str_remove(..2, 'SC'), '_',..4)), 'Negative|Doublet') & n>Strain_qc_cutoff ~ !!rlang::sym(paste0('Strain_hmo_', str_remove(..2, 'SC'), '_',..4)), 
          T ~ 'PoorQC'),
        'Strain_DN' = !!rlang::sym(paste0('Strain_hmo_', str_remove(..2, 'SC'), '_',..4)),
        'dbt_neg_all' = case_when(
          if_all(c(paste0('Strain_', ..5)), ~ .== "Doublet") ~ "DoubletAll",
          if_all(c(paste0('Strain_', ..5)), ~ .== "Negative") ~ "NegativeAll",
          if_all(c(paste0('Strain_', ..5)), ~ .== "Doublet" | .== "Negative") ~ "DbtNegMix", 
          T ~ 'LikelySinglet')
      )
  })



```


```{r}
## Cell counts per cross-k consensus (Strn_c), QC strain (Strain_qc), and aligner-specific optimal-k label
list(strn_c_mdata, sm_aln_nms_al, optimum_k_aln)  %>%
  purrr::pmap(.,~{
    ..1 %>% dplyr::count(Strn_c, Strain_qc, !!rlang::sym(paste0("Strain_hmo_", str_remove(..2, "SC"), "_", ..3))) %>% arrange(desc(n))
  })
```



```{r}
## Inspect barcodes called Doublet or Negative across all k-value columns — these are candidates for removal
list(strn_c_mdata, sm_aln_nms_al, optimum_k_aln)  %>%
  purrr::pmap(.,~{
    ..1 %>% filter(if_all(starts_with('Strain_hmo_M'), ~ .== "Doublet" | .== "Negative")) %>% dplyr::count(Strn_c, dbt_neg_all, !!rlang::sym(paste0("Strain_hmo_", str_remove(..2, "SC"), "_", ..3)))
  })
```




```{r}
## Summary of final strain assignments: counts by Strain_qc alone, then by full consensus combination
list(strn_c_mdata) %>%
  purrr::pmap(., ~{ ..1 %>% dplyr::count(Strain_qc) %>% arrange(desc(n)) })

list(strn_c_mdata, sm_aln_nms_al, optimum_k_aln) %>%
  purrr::pmap(., ~{ ..1 %>% dplyr::count(Strn_c, Strain_qc, !!rlang::sym(paste0("Strain_hmo_", str_remove(..2, "SC"), "_", ..3))) %>% arrange(desc(n)) })
```


```{r}
## Count cells per strain label at the optimal k (Strain_DN includes Doublet and Negative)
map(strn_c_mdata, . %>% dplyr::count(Strain_DN))

```

```{r}
strn_c_mdata <- map(strn_c_mdata,~mutate(., 'strain' = Strain_DN))


```


```{r}
## Confirm cell counts per strain and total barcode count per sample
map(strn_c_mdata, . %>%
      dplyr::count(strain)
)

map(strn_c_mdata, dim)

```


```{r}
##Save strain + new metadata as RDS
map(strn_c_mdata, 
    . %>% 
      dplyr::select(-c(starts_with("Strain_MSC"))) %>%
      dplyr::select("bcode", "Strn_c", starts_with("Strain"), "strain", "dbt_neg_all")) %>%
  saveRDS(paste0("data/processed/", species, "/" , sv_dir, "/strn_c_new_mdata_m22_no_stage.RDS"))


```

### PCA calculation
```{r}
soupo_PC_mtx <- soupc_all_al %>%
  dplyr::select(-c(status, assignment, log_prob_singleton, log_prob_doublet)) %>% 
  dplyr::group_split(sample_k) %>%
  purrr::set_names(purrr::map_chr(., ~.x$sample_k[1])) %>%
  map(. %>% column_to_rownames(., "barcode") %>%
        dplyr::select(where(~!all(is.na(.)))) %>%
        dplyr::select(-sample_k) %>%
        mutate(across(where(is.numeric), ~ (. / rowSums(across(where(is.numeric)))))) %>%
        as.matrix() %>%
        t())

soupo_PC_mtx <- soupo_PC_mtx[as.character(sorted_names_al)]
```



```{r}
##Calculate PCA
soupo_PCA <- soupo_PC_mtx %>% 
  map(. %>% prcomp(.))
```



```{r}
## make directory for strain summary output
system(paste0("mkdir -p data/processed/",species, "/" , sv_dir))

```

```{r}
## Save PCA
saveRDS(soupo_PCA, paste0("data/processed/", species, "/" , sv_dir,"/soupo_PCA_m22.RDS"))
```

# !! THE BELOW IS OPTIONAL BUT GIVES CONFIDENCE THAT HISAT AND MINIMAP ARE LARGELY IN AGREEMENT

### Validate within-aligner harmonization

Check that each original strain label maps to exactly one harmonized label. All `k_test` values should be `TRUE`.

```{r}
strn_harmnzd_assesmnt <- list(strn_harmnzd_tbls, don_k_names_aln, don_hmonizd_k_names_aln) %>%
  pmap(~{
    tbl <- ..1
    nms <- ..2
    nms_hmo <- ..3
    
    map2(nms, nms_hmo,  ~{
      tbl %>% 
        mutate("strn_o" =!!rlang::sym(.x) , "strn_n" = !!rlang::sym(.y)) %>% 
        dplyr::count(strn_o, strn_n) %>% 
        arrange(strn_o)
    }) %>% 
      set_names(nms) %>% 
      bind_rows(., .id = "strn_k") 
    
  })

strn_harmnzd_assesmnt
```

```{r}
## Full detail: one row per (original strain, harmonized strain) pair per k
bind_rows(strn_harmnzd_assesmnt) #%>% view
```

```{r}
## Summary: k_test == TRUE means the number of unique harmonized labels equals k (one-to-one mapping confirmed)
hmonizd_k_test_summary <- bind_rows(strn_harmnzd_assesmnt) %>% group_by(strn_k) %>% mutate(dup = duplicated(strn_o) | duplicated(strn_n)) %>% count(dup) %>% mutate(k=n-2) %>% mutate(k_test = as.numeric(str_remove(strn_k, "Str.*_")) == k)
hmonizd_k_test_summary

```

```{r}
## Confirm all k_test are TRUE; any FALSE indicates a harmonization collision
table(hmonizd_k_test_summary$k_test)
```


### Harmonize minimap2 and HISAT2 labels

Change HISAT2 labels to match those of minimap2, then merge harmonized strain labels for further cross-aligner harmonization.
```{r , fig.width = 10, fig.height = 5, warning=FALSE}
mnmp_hsat_comp_tbl <- list(strn_harmnzd_tbls[str_detect(names(strn_harmnzd_tbls), "minmap")], strn_harmnzd_tbls[str_detect(names(strn_harmnzd_tbls), "hsat")]) %>% 
  pmap(., ~..1 %>% 
         left_join(., ..2, by = "bcode") %>%
         select(bcode, starts_with("Strain_hmo")) 
  )
mnmp_hsat_comp_tbl[[1]]
```

```{r}
don_k_names_minmap <- imap(opt_clusters_nms_aln[str_detect(names(opt_clusters_nms_aln), "minmap")], ~paste0('Strain_hmo_', str_remove(.y, "SC"), '_', .x)) %>% unlist(., recursive = F)
don_k_names_hsat <- imap(opt_clusters_nms_aln[str_detect(names(opt_clusters_nms_aln), "hsat")], ~paste0('Strain_hmo_', str_remove(.y, "SC"), '_', .x)) %>% unlist(., recursive = F)


don_hmonizd_minmap_base <- imap(opt_clusters_nms_aln[str_detect(names(opt_clusters_nms_aln), "hsat")], ~paste0('Strain_hmo_', str_replace(str_remove(.y, "SC"), "hsat","hsatHMO"), '_', .x)) %>% unlist(., recursive = F)

mnmp_hsat_comp_tbl_repd <-  map2(mnmp_hsat_comp_tbl, opt_clusters_nms_aln[str_detect(names(strn_harmnzd_tbls), "minmap")],  ~rep(list(.x),  length(.y))) %>% unlist(., recursive=F)
mnmp_hsat_comp_tbl_repd <- mnmp_hsat_comp_tbl_repd %>% set_names(map2(names(mnmp_hsat_comp_tbl_repd), unlist(opt_clusters_nms_aln[str_detect(names(strn_harmnzd_tbls), "minmap")]), ~str_replace_all(.x, "minmap.*", paste0("K", .y))))


strn_harmnzd_tbls_mn_hs <- list(mnmp_hsat_comp_tbl_repd, don_k_names_minmap, don_k_names_hsat, don_hmonizd_minmap_base) %>%
  pmap(~{
    
    strain_harmonize_fn(strn_tbl = ..1, strn1 = ..2, strn2 = ..3, out_strn = ..4) 
    
  })


strn_harmnzd_tbls_mn_hs[[1]]
```


```{r , fig.width = 10, fig.height = 5, warning=FALSE}
## Keep only the strain columns needed for cross-aligner comparison (minimap2, HISAT2, HISAT2-harmonized)
strn_harmnzd_tbls_mn_hs2 <- pmap(list(strn_harmnzd_tbls_mn_hs, don_k_names_minmap, don_k_names_hsat, don_hmonizd_minmap_base), ~..1 %>% select(bcode, ..2, ..3, ..4))

```



#### Validate cross-aligner harmonization: minimap2 vs HISAT2-harmonized agreement

```{r , fig.width = 10, fig.height = 5, warning=FALSE}
aln_comp_tbl <- map(strn_harmnzd_tbls_mn_hs2, ~.x %>%
                      select(bcode, contains("_minmap_"), contains("hsatHMO")) %>%
                      pivot_longer(cols = starts_with("Strain_hmo"),
                                   names_to = c(".value", "K"),
                                   names_pattern = ".*[0-9].*_(.*)_([0-9]+$)")
) 

imap(aln_comp_tbl, possibly(~phmap_lab_transfr(.x, col_labs= `minmap`, row_labs=`hsatHMO`, row_labs_c ='hsatHMO', col_labs_c= 'minmap', ttl = .y)), 'k=1 so no groups to compare which is necessary for the heatmap')

```



### Sankey plots: cross-aligner strain label transitions (minimap2 → hsatHMO → hsat)

```{r, fig.width=6, fig.height=4, eval=T}
snk_plts_hmo <- list(strn_harmnzd_tbls_mn_hs2, don_k_names_minmap,  don_k_names_hsat, str_remove_all(don_k_names_hsat, "S.*_"), don_hmonizd_minmap_base) %>%
  pmap(~{  
    
    snk_pt_fn(..1, strns_2_plt = c(..2, ..5, ..3), labls = c("minimap", "hisatHMO", "hisat"), bp_face = c("plain"))
    
  })

```


```{r, eval=T}
snk_plts_hmo <- snk_plts_hmo %>% imap(., ~.x + labs(title = .y))
```


```{r, fig.width=6, fig.height=4, eval=T}
snk_plts_hmo[str_detect(names(snk_plts_hmo), "MSC14|MSC49|MSC55|MSC60")]
```

#### Validate HISAT2 harmonization: original vs harmonized labels should be 100% concordant

HISAT2 (`hsat`) and HISAT2-harmonized (`hsatHMO`) strain labels should match perfectly for every barcode. Any discordance indicates an error in the HISAT2 → minimap2 label mapping step.
```{r , fig.width = 10, fig.height = 5, warning=FALSE}
aln_comp_tbl <- map(strn_harmnzd_tbls_mn_hs2, ~.x %>% 
                      select(bcode, contains("_hsat_"), contains("hsatHMO")) %>%
                      pivot_longer(cols = starts_with("Strain_hmo"),
                                   names_to = c(".value", "K"),
                                   names_pattern = ".*[0-9].*_(.*)_([0-9]+$)")
) 

imap(aln_comp_tbl, possibly(~phmap_lab_transfr(.x, col_labs= `hsat`, row_labs=`hsatHMO`, row_labs_c ='hsatHMO', col_labs_c= 'hsat', ttl = .y)), 'k=1 so no groups to compare which is necessary for the heatmap')

```

