---
title: "MSC QC and preprocessing"
date: '2022-07-11'
author: 'Jesse Rop'
output:
  html_document:
    toc: true
    theme: united
    code_folding: hide
    df_print: paged
---

```{r setup_libs, include=FALSE, message=FALSE, warning=FALSE}
# Load core packages (quiet)
suppressPackageStartupMessages({
  library(Seurat)
  library(sctransform)
  library(glmGamPoi)
  library(patchwork)
  library(ggrepel)
  library(ggpubr)
  library(dplyr)
  library(stringr)
  library(purrr)
  library(readr)
  library(tibble)
  library(ggplot2)
  library(tidyr)
  library(cowplot)
  library(conflicted)
  library(SingleCellExperiment)
  library(parallel)
  library(ggforce)
  
  conflict_prefer("select", "dplyr")
  conflict_prefer("filter", "dplyr")
  conflict_prefer("rename", "dplyr")
  conflict_prefer("map", "purrr")
  conflict_prefer("count", "dplyr")
  conflicts_prefer(base::saveRDS)
  conflicts_prefer(SingleCellExperiment::colData)
  conflicts_prefer(base::as.data.frame)
  conflicts_prefer(base::intersect)
  conflicts_prefer(dplyr::desc)
  conflicts_prefer(dplyr::count)
})


```

```{r setup_libs, include=FALSE, message=FALSE, warning=FALSE}
## Helper functions to reduce redundancy and improve reproducibility
# get_umap_mdata: attach UMAP embeddings to meta.data for a list of Seurat objects
get_umap_mdata <- function(seu_list) {
  map(seu_list, ~{
    .x@meta.data %>%
      rownames_to_column("bcode") %>%
      left_join(., .x@reductions$umap@cell.embeddings %>% as.data.frame() %>% rownames_to_column("bcode"), by = "bcode")
  })
}

## read_seurat_objs: read matching files and return list for requested samples
read_seurat_objs <- function(glob_pattern, samples) {
  files <- Sys.glob(glob_pattern)
  files %>%
    set_names(nm = (str_remove_all(., 'data/processed/Pf/|/seu_obj.*'))) %>%
    unlist(recursive = FALSE) %>%
    { .[samples] } %>%
    { .[!is.na(.)] } %>%
    map(readRDS)
}

## normalize_to_RNA: set `RNA` assay and call `NormalizeData()` for each object
normalize_to_RNA <- function(seu_list) {
  map(seu_list, ~{
    DefaultAssay(.x) <- "RNA"
    NormalizeData(.x)
  })
}
```

# MSC stage annotation and gene expression QC 

## setup, read input files and exploratory visualizations and general functions

```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = '/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/Talleh')

```

### Variable declarations


#### Source common functions, variables and tables
```{r}
## Load project-specific helper scripts and shared variables
## Boxplot labeling and project helper functions
source("~/Talleh_code/pf_common_vars.R")
source("~/Talleh_code/project_common_fns.R")

```

<!-- Helpers and shared variables are defined above; the following loads project-specific globals. -->

## Read in sample name details

```{r}
# Select sample names
sample_name <- c("MSC49", "MSC68")

```


```{r, eval=T}
# Read annotated Seurat objects (prelim stage) for the selected samples
bp_prelim_stg <- list.files("data/processed/Pf", pattern = "bp_prelim_stg.RDS$", full.names = T, recursive = T, include.dirs = T) %>% 
  set_names(nm = (str_remove_all(.,'data/processed/Pf/|/seu_obj/|bp_prelim_stg.*.RDS'))) %>% 
  .[sample_name] %>%
  map(readRDS)
```


```{r, eval=T}
# Run SCT preprocessing wrapper on each object (saves results in list)
all_seu_prelim_stg_sct <- map(bp_prelim_stg, ~sct_run_fn(Seurat_object = .x))
all_seu_prelim_stg_sct
```



```{r, warning=FALSE, message=FALSE}
# Quick PCA plots coloured by stage for SCT objects (visual check)
imap(all_seu_prelim_stg_sct, ~DimPlot(.x, dims = c(1,2), group.by = "stage_fld", pt.size = 0.1, reduction = "pca") + scale_color_manual(values = sp_col_new) + labs(title = .y))
imap(all_seu_prelim_stg_sct, ~DimPlot(.x, dims = c(1,2), group.by = "stage_fld", pt.size = 0.1) + scale_color_manual(values = sp_col_new) + labs(title = .y))
```

```{r, warning=FALSE, message=FALSE}

## Extract UMAP embeddings into metadata frame for plotting/export
all_seu_prelim_stg_sct_umap_mdata <- get_umap_mdata(all_seu_prelim_stg_sct)

```


```{r, message=FALSE, warning=FALSE, eval=T}
# Save the poorQC cells barcodes for plotting
saveRDS(all_seu_prelim_stg_sct_umap_mdata, "data/processed/Pf/m2_all/all_seu_prelim_stg_sct_umap_mdata.RDS")

```

```{r, message=F, warning=F, eval=T}
# Read normalized prelim-stage Seurat objects (if present) for the selected samples
all_seu_prelim_stg_norm <- read_seurat_objs("data/processed/Pf/*/seu_obj/bp_prelim_stg_norm.RDS", sample_name)
```

```{r, warning=FALSE, message=FALSE}
# Diagnostic PCA/UMAP plots for normalized objects
imap(all_seu_prelim_stg_norm, ~DimPlot(.x, dims = c(2,3), group.by = "stage_fld", pt.size = 0.1, reduction = "pca") + scale_color_manual(values = sp_col_new) + labs(title = .y))
imap(all_seu_prelim_stg_norm, ~DimPlot(.x, dims = c(1,2), group.by = "stage_fld", pt.size = 0.1) + scale_color_manual(values = sp_col_new) + labs(title = .y))
```

```{r, warning=FALSE, message=FALSE}

## Extract UMAP embeddings for normalized objects
all_seu_prelim_stg_norm_umap_mdata <- get_umap_mdata(all_seu_prelim_stg_norm)
```


```{r, message=FALSE, warning=FALSE, eval=T}
# Save the poorQC cells barcodes for plotting
saveRDS(all_seu_prelim_stg_norm_umap_mdata, "data/processed/Pf/m2_all/all_seu_prelim_stg_norm_umap_mdata.RDS")

```


```{r, warning=FALSE, message=FALSE, fig.width=12, fig.height=3.5}
# If normalized versions are not available, derive them from SCT objects
all_seu_prelim_stg_norm <- normalize_to_RNA(all_seu_prelim_stg_sct)
```

```{r, warning=FALSE, message=FALSE, fig.width=12, fig.height=3.5}
# Example marker gene feature plots (UMAP)
imap(all_seu_prelim_stg_norm, ~FeaturePlot(.x, reduction = "umap",  label = T, features = c("PF3D7-0113600", "PF3D7-1102500", "PF3D7-1222600"), order = T, ncol = 3) + labs(title = .y))
```

```{r}
all_seu_prelim_stg_norm[1] %>% map(., ~.x@meta.data %>% 
                                     count(stage_final2, combined_labs, stage_fld) %>% 
                                     filter(str_detect(stage_fld, "LE"))
)

```


```{r, warning=FALSE, message=FALSE, fig.width=4, fig.height=4.5}
## in cellbender but not in edrops
plts_stg <- imap(all_seu_prelim_stg_norm, possibly(~{
  cell_stg_plot <- DimPlot(.x, group.by = "stage_fld", cols = sp_col_new, pt.size = 0.2) + 
    theme_void() + 
    theme(legend.position = "none", panel.border = element_rect(color = "black", fill = NA, size = 1))  +
    labs(title = .y)
  
  cell_stg_plot$data <- cell_stg_plot$data %>% rownames_to_column('bcode') %>% left_join(., .x@meta.data[,'seurat_clusters', drop = F] %>% rownames_to_column('bcode'), by='bcode') %>% column_to_rownames('bcode')
  
  cell_stg_plot<- LabelClusters(cell_stg_plot, id = "seurat_clusters", size = 6, repel = T)
  return(cell_stg_plot)
  
}, "no cells"))

```

```{r, warning=FALSE, message=FALSE, fig.height= 3.5, fig.width= 10, eval=T}
plot_grid(plotlist = plts_stg, ncol = 4)
```



## 1. MITOCHONDRIAL PERCENT VISUALIZATION AND CELL REMOVAL

Get mt_filt based on automatic removal of quantiles
```{r}
all_seu_prelim_stg_norm <- all_seu_prelim_stg_norm%>% 
  map(., ~{
    
    mit_fts <- rownames(.x)[str_detect(rownames(.x), "MIT")]
    api_fts <- rownames(.x)[str_detect(rownames(.x), "API")]
    
    # Compute mitochondrial and apicoplast percent features and add to metadata
    .x %>% PercentageFeatureSet(., features = mit_fts, col.name = 'percent.mt', assay = "RNA") %>%
      PercentageFeatureSet(., features = api_fts, col.name = 'percent.api', assay = "RNA") 
  })

```

```{r}
## Determine mt% cutoff per sample (99.5th percentile) to flag high-mt cells
mt_filt <- unlist(map(all_seu_prelim_stg_norm, ~quantile(.x@meta.data[, "percent.mt"], 0.995, na.rm = T)))
```


```{r, results='asis'}
std_theme <- theme(axis.text=element_text(size=18), 
                   axis.title=element_text(size=18,face="bold"), 
                   legend.text = element_text(size=22), 
                   legend.title = element_text(size=22,face="bold"))


```

```{r, results='asis'}
## Plot UMI count vs percent.mt with per-sample mt cutoff line
pmap(list(all_seu_prelim_stg_norm, mt_filt, names(all_seu_prelim_stg_norm)), ~{
  FeatureScatter(..1, feature1 = "pf_umi_count", feature2 = "percent.mt", group.by = "stage_fld", pt.size = 0.1)+ 
    scale_color_manual(values = sp_col_new) +
    geom_hline(yintercept = ..2)  +
    labs(title = ..3) +
    std_theme
  
})

```


```{r, eval=T}

after_umi100_filt <- map(all_seu_prelim_stg_norm, dim)%>% as.data.frame() %>% .[2,]
after_umi100_filt
```



```{r}
map2(all_seu_prelim_stg_norm, mt_filt, ~.@meta.data %>% dplyr::count(percent.mt>.y ))
```

```{r}
## Barcodes removed
mt_filt_rm_bcodes <- map2(all_seu_prelim_stg_norm, mt_filt, ~colnames(subset(.x, percent.mt < .y, invert = T ) ))

```

```{r, eval=T}
## write out object with mitochondrial percent count for comparison with older iteration of other same datasets
imap(all_seu_prelim_stg_norm, ~saveRDS(.x, paste0("data/processed/Pf/",.y, "/seu_obj/seu_prelim_stg_norm_for_comp_b4_pct_mito_filt.RDS")))

```


```{r, eval=T}
## Optionally remove cells exceeding the per-sample mitochondrial cutoff (mt_filt)
## This step drops cells with percent.mt >= the sample-specific threshold
all_seu_prelim_stg_norm <- all_seu_prelim_stg_norm %>%
  map2(., mt_filt, ~.x  %>%
         subset(., percent.mt < .y )
  )

```

```{r, eval=T}
## write out object with mitochondrial percent count for comparison with older iteration of other same datasets
imap(all_seu_prelim_stg_norm, ~saveRDS(.x, paste0("data/processed/Pf/",.y, "/seu_obj/seu_prelim_stg_norm_for_comp_after_pct_mito_filt.RDS")))

```

```{r, eval=T}

## Report cell counts remaining after mt filtering
after_mt_filt <- map(all_seu_prelim_stg_norm, dim)%>% as.data.frame() %>% .[2,]
after_mt_filt
```


```{r, warning=FALSE, message=FALSE}
imap(all_seu_prelim_stg_norm, ~DimPlot(.x, dims = c(1,3), group.by = "stage_fld", pt.size = 0.1)+ scale_color_manual(values = sp_col_new) + labs(title = .y))
imap(all_seu_prelim_stg_norm, ~DimPlot(.x, dims = c(1,2), group.by = "stage_fld", pt.size = 0.1)+ scale_color_manual(values = sp_col_new) + labs(title = .y))
```


```{r, warning=FALSE, message=FALSE}
imap(all_seu_prelim_stg_norm, ~VlnPlot(.x, features = "nFeature_RNA", group.by = "stage_fld", pt.size = 0.1) + scale_color_manual(values = sp_col_new) + scale_y_log10() +  labs(title = .y))
```


### 2.1 GENE UMI COUNT VISUALIZATIONS

```{r, fig.height = 3.5, fig.width = 12}

all_seu_prelim_stg_norm_mdata_mrg <- map(all_seu_prelim_stg_norm, ~.@meta.data) %>% 
  bind_rows(., .id = 'donor') 

```

```{r, fig.height = 3.5, fig.width = 12}
all_seu_prelim_stg_norm_mdata_mrg %>%
  ggplot(aes(x = log10(pf_gn_count), y = log10(pf_umi_count), colour = mthds_union)) +
  geom_point(size = 0.01) +
  geom_abline() +
  facet_wrap(donor ~., nrow = 1) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  guides(color = guide_legend(title = 'Call methods', override.aes = list(size = 4), ncol = 1))

```



```{r, warning=FALSE, message=FALSE, fig.width=4, fig.height=4.5}
## umi
plts_nc <- imap(all_seu_prelim_stg_norm, possibly(~FeaturePlot(.x, label = T, features = "pf_umi_count") +  theme_void() + theme(legend.position = "none") + labs(title = .y), "no cells"))

```

```{r, warning=FALSE, message=FALSE, fig.height= 3.5, fig.width= 10, eval=T}
plot_grid(plotlist = plts_nc, ncol = 4)
```


```{r, warning=FALSE, message=FALSE, fig.width=4, fig.height=4.5}
plts_l100 <- imap(all_seu_prelim_stg_norm, possibly(~DimPlot(.x, label = T, cells.highlight =  colnames(.x[,.x@meta.data$pf_gn_count <100]), sizes.highlight = 0.05) + 
                                                      theme_void() + theme(legend.position = "none", panel.border = element_rect(color = "black", fill = NA, size = 1)) +
                                                      labs(title = .y), "no cells"))

```

```{r, warning=FALSE, message=FALSE, fig.height= 3.5, fig.width= 10, eval=T}
plot_grid(plotlist = plts_l100, ncol = 4)
```


## 2. CELLBENDER METRICS CELL VISUALIZATIONS



### 2.1 CELL PROBABILITY VISUALIZATIONS


```{r, warning=FALSE, message=FALSE, fig.width=4, fig.height=4.5}
## in cellbender but not in edrops
## Highlight cells with low CellBender probability (< 0.9)
plts_seu_cp9 <- imap(all_seu_prelim_stg_norm, possibly(~DimPlot(.x, label = T, cells.highlight =  colnames(.x[ , .x@meta.data$cell_probability < 0.9]), sizes.highlight = 0.05) + theme_void() + theme(legend.position = "none", panel.border = element_rect(color = "black", fill = NA, size = 1))  + labs(title = .y), "no cells"))

```

```{r, warning=FALSE, message=FALSE, fig.height= 3.5, fig.width= 10, eval=T}
plot_grid(plotlist = plts_seu_cp9, ncol = 4)

```


### 2.1 BACKGROUND FRACTION VISUALIZATION


## Assess cellbender estimated metrics cell probability and background fraction
```{r, fig.height = 3.5, fig.width = 12}
## Visualise estimated background fraction from CellBender by method and donor
all_seu_prelim_stg_norm_mdata_mrg %>%
  ggplot(aes(x = mthds_union, y = background_fraction)) +
  geom_boxplot() +
  facet_wrap(donor ~., ncol = 4) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


```{r, fig.height = 3.5, fig.width = 15}
## Scatter of gene count vs background fraction (check extreme cells)
all_seu_prelim_stg_norm_mdata_mrg %>%
  ggplot(aes(x = pf_gn_count, y = background_fraction, colour = mthds_union)) +
  geom_point(size = 0.01) +
  geom_hline(yintercept = 0.5) +
  geom_vline(xintercept = 0.5) +
  geom_abline() +
  facet_wrap(donor ~., ncol = 4, scales = "free") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  guides(color = guide_legend(title = 'Call methods', override.aes = list(size = 4), ncol = 1))

```

```{r, warning=FALSE, message=FALSE, fig.width=4, fig.height=4.5}
## background fraction cellbender
all_seu_prelim_stg_norm_mdata_mrg %>% count(donor, background_fraction> 0.5 )

```


```{r, warning=FALSE, message=FALSE, fig.width=4, fig.height=4.5}
## 
## Highlight cells with high estimated background fraction (> 0.5)
plts_seu_bg5 <- imap(all_seu_prelim_stg_norm, possibly(~DimPlot(.x, label = T, cells.highlight =  colnames(.x[, .x@meta.data$background_fraction > 0.5 ]), sizes.highlight = 0.05) + theme_void() + theme(legend.position = "none", panel.border = element_rect(color = "black", fill = NA, size = 1))  +
                                                         labs(title = .y), "no cells"))

```

```{r, warning=FALSE, message=FALSE, fig.height= 3.5, fig.width= 15, eval=T}
plot_grid(plotlist = plts_seu_bg5, ncol = 4)

```



```{r, warning=FALSE, message=FALSE}
#ap2g
imap(all_seu_prelim_stg_norm, ~FeaturePlot(.x, dims = c(1,2), features = "PF3D7-1222600", order = T, slot = "data")+ labs(title = .y))
```


```{r, warning=FALSE, message=FALSE, fig.width=4, fig.height=4.5}
## background fraction cellbender
plts_cbg <- imap(all_seu_prelim_stg_norm, possibly(~FeaturePlot(.x, label = T, features = "background_fraction") +  theme_void() + theme(legend.position = "none") + labs(title = .y), "no cells"))

```

```{r, warning=FALSE, message=FALSE, fig.height= 3.5, fig.width= 10, eval=T}
plot_grid(plotlist = plts_cbg, ncol = 4)
```


### 4 UMI AND GENE COUNTS OUTLIERS VISUALIZATIONS AND REMOVAL

#### 4.1 GENE UMI COUNT VISUALIZATIONS

```{r, fig.height = 3.5, fig.width = 12}
all_seu_prelim_stg_norm_mdata_mrg %>%
  ggplot(aes(x = log10(pf_gn_count), y = log10(pf_umi_count), colour = mthds_union)) +
  geom_point(size = 0.01) +
  geom_abline() +
  facet_wrap(donor ~., nrow = 1) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  guides(color = guide_legend(title = 'Call methods', override.aes = list(size = 4), ncol = 1))

```



```{r, warning=FALSE, message=FALSE, fig.width=4, fig.height=4.5}
## umi
plts_nc <- imap(all_seu_prelim_stg_norm, possibly(~FeaturePlot(.x, label = T, features = "pf_umi_count") +  theme_void() + theme(legend.position = "none") + labs(title = .y), "no cells"))

```

```{r, warning=FALSE, message=FALSE, fig.height= 3.5, fig.width= 10, eval=T}
plot_grid(plotlist = plts_nc, ncol = 4)
```


```{r, warning=FALSE, message=FALSE, fig.width=4, fig.height=4.5}
plts_l100 <- imap(all_seu_prelim_stg_norm, possibly(~DimPlot(.x, label = T, cells.highlight =  colnames(.x[,.x@meta.data$pf_gn_count <100]), sizes.highlight = 0.05) + 
                                                      theme_void() + theme(legend.position = "none", panel.border = element_rect(color = "black", fill = NA, size = 1)) +
                                                      labs(title = .y), "no cells"))

```

```{r, warning=FALSE, message=FALSE, fig.height= 3.5, fig.width= 10, eval=T}
plot_grid(plotlist = plts_l100, ncol = 4)
```

#### 4.2 Remove outliers
########################### START Remove outliers ##############################

```{r}
# Compute stage- and cluster-level quantile cutoffs for outlier detection
# (used to add metadata cutoffs per cell)
# gn_uc <- c(1900, 2600, 3300, 3800, 3200)
# gn_lc <- c(250, 250, 250, 300, 250)
all_seu_prelim_stg_norm <- pmap(list(all_seu_prelim_stg_norm, rep('nFeature_RNA', length(sample_name)), rep('nCount_RNA', length(sample_name))), ~{
  ..1@meta.data %>%
    rownames_to_column('bcode') %>%
    group_by(stage_fld) %>%
    mutate("gn_lci_cl_stg" = quantile(!!rlang::sym(..2), 0.01, na.rm = T),
           "gn_uci_cl_stg" = quantile(!!rlang::sym(..2), 0.99, na.rm = T),
           "rna_lci_cl_stg" = quantile(!!rlang::sym(..3),  0.01, na.rm = T),
           "rna_uci_cl_stg" = quantile(!!rlang::sym(..3), 0.99, na.rm = T)
    ) %>%
    ungroup()%>%
    group_by(seurat_clusters) %>%
    mutate("gn_lci_cl_seu" = quantile(!!rlang::sym(..2), 0.01, na.rm = T),
           "gn_uci_cl_seu" = quantile(!!rlang::sym(..2), 0.99, na.rm = T),
           "rna_lci_cl_seu" = quantile(!!rlang::sym(..3),  0.01, na.rm = T),
           "rna_uci_cl_seu" = quantile(!!rlang::sym(..3), 0.99, na.rm = T)
    ) %>%
    ungroup()%>%
    select(bcode,contains("ci_cl")) %>%
    column_to_rownames('bcode')%>%
    AddMetaData(..1, .)
})
```


```{r, warning=FALSE, message=FALSE, fig.width=4, fig.height=4.5}
## in cellbender but not in edrops
plts_bg <- imap(all_seu_prelim_stg_norm, possibly(~{
  DimPlot(.x, label = T, 
          cells.highlight =  .x@meta.data %>% rownames_to_column('bcode') %>% filter(pf_gn_count<gn_lci_cl_seu | nCount_RNA < rna_lci_cl_seu | nCount_RNA > rna_uci_cl_seu | pf_gn_count > gn_uci_cl_seu)  %>% pull(bcode), 
          sizes.highlight = 0.05) + 
    theme_void() + theme(legend.position = "none", panel.border = element_rect(color = "black", fill = NA, size = 1))  +
    labs(title = .y)}, "no cells"))

```

```{r, warning=FALSE, message=FALSE, fig.height= 3.5, fig.width= 10, eval=T}
plot_grid(plotlist = plts_bg, ncol = 4)
```
#### Violin plots


```{r, fig.height = 8, fig.width = 12}

fct_theme <- theme(text=element_text(size=14),
                   axis.text=element_text(size=20), 
                   axis.title=element_text(size=20,face="bold"), 
                   legend.text = element_text(size=20), 
                   legend.title = element_text(size=20,face="bold"),
                   strip.text =element_text(size=22,face="plain", colour = 'black'))

```


```{r, fig.height = 4.5, fig.width = 12}

## Prepare combined metadata table with donor IDs for violin plots
md_don_cl <- map(all_seu_prelim_stg_norm, ~.@meta.data #%>% 
                 # mutate(across(c(sim_lab1, sim_lab2), as.numeric))
) %>% 
  bind_rows(., .id = 'donor') %>%
  mutate(across(donor, ~factor(., levels = as.character(sample_name))))

```

```{r, fig.height = 8, fig.width = 12}
## Violin plotting helper: QC visualisation per cluster/stage
my21_qc_vln_fn <- function(clust_var = "StagePrelimC", clust_cols, ci_sufx = "_stg", row_numbers = 2) {
  pmap(list(c('pf_umi_count', 'pf_gn_count'), c('Transcript', 'Gene'), paste0(c("rna_lci_cl","gn_lci_cl"), ci_sufx), paste0(c("rna_uci_cl","gn_uci_cl"), ci_sufx)), ~{
    
    md_don_cl %>% 
      ggplot(aes(x = !!rlang::sym(clust_var), y = !!rlang::sym(..1)/1000, color = !!rlang::sym(clust_var))) +
      scale_color_manual(values = clust_cols) +
      theme_classic() +
      geom_violin(drop = F) +
      ggforce::geom_sina(size = 0.1,  alpha = 0.5)+
      facet_wrap(donor ~., nrow = row_numbers, scales = "free_x")+
      labs(y = "Gene count (X 1000)")+
      geom_segment(aes(x=as.numeric(!!rlang::sym(clust_var))-0.3, xend=as.numeric(!!rlang::sym(clust_var))+0.3,
                       y = !!rlang::sym(..3)/1000, yend= !!rlang::sym(..3)/1000 ), inherit.aes=F, color="orange", size=0.5)+
      geom_segment(aes(x=as.numeric(!!rlang::sym(clust_var))-0.3, xend=as.numeric(!!rlang::sym(clust_var))+0.3,
                       y = !!rlang::sym(..4)/1000, yend= !!rlang::sym(..4)/1000 ), inherit.aes=F, color="red", size=0.5) +
      labs(x= 'Cluster', y= paste0(..2, ' count (X 1000)')) +
      fct_theme+
      theme(legend.position = "bottom",
            axis.text.x = element_text(angle =40, hjust = 1, size = 10)) +
      guides(color = guide_legend(title = 'Cluster',override.aes = list(size = 4), nrow = 1))
  })
}

```


```{r, fig.height = 5, fig.width = 15}
my21_qc_vln_fn(clust_var = "stage_fld", clust_cols = sp_col_new, ci_sufx = "_stg", row_numbers = 1 ) 

```



```{r}

## filter based on gene/transcript count
all_seu_filtOlMtCb_rm <- map(all_seu_prelim_stg_norm, ~subset(.x, nFeature_RNA < gn_lci_cl_seu | nCount_RNA < rna_lci_cl_seu | nCount_RNA > rna_uci_cl_seu | nFeature_RNA > gn_uci_cl_seu))

map(all_seu_filtOlMtCb_rm, ~.x@meta.data %>% count(stage_fld))

```


```{r, fig.height = 4, fig.width = 15}
my21_qc_vln_fn(clust_var = "seurat_clusters", clust_cols = cluster_cols, ci_sufx = "_seu", row_numbers = 1 ) 

```


```{r, warning=FALSE, message=FALSE, fig.width=4, fig.height=4.5}
plts_outlrs <- imap(all_seu_prelim_stg_norm, possibly(~{
  DimPlot(.x, label = T, 
          cells.highlight =  .x@meta.data %>% rownames_to_column('bcode') %>% filter(nFeature_RNA<gn_lci_cl_seu | nCount_RNA < rna_lci_cl_seu | nCount_RNA > rna_uci_cl_seu | nFeature_RNA > gn_uci_cl_seu)  %>% pull(bcode), 
          sizes.highlight = 0.05) + 
    theme_void() + theme(legend.position = "none", panel.border = element_rect(color = "black", fill = NA, size = 1))  +
    labs(title = .y)}, "no cells"))

```

```{r, warning=FALSE, message=FALSE, fig.height= 3.5, fig.width= 15, eval=T}
plot_grid(plotlist = plts_outlrs, ncol = 4)
```



Number of cells per strain_gt to remove based on REVISED CUT-OFFS to assist souporcell to perform clustering correctly
```{r}

pmap(list(all_seu_prelim_stg_norm), ~..1@meta.data %>% dplyr::count(nFeature_RNA<gn_lci_cl_seu ,nCount_RNA < rna_lci_cl_seu, nCount_RNA > rna_uci_cl_seu, nFeature_RNA > gn_uci_cl_seu ))
```


```{r, warning=FALSE, message=FALSE}
#REX1
imap(all_seu_prelim_stg_norm, ~FeaturePlot(.x, dims = c(1,2), features = "PF3D7-0935900", reduction = "umap")+ labs(title = .y))

```


```{r, warning=FALSE, message=FALSE}
#ap2g
imap(all_seu_prelim_stg_norm, ~FeaturePlot(.x, dims = c(1,2), features = "PF3D7-1222600", order = T, slot = "data")+ labs(title = .y))
```
```{r, warning=FALSE, message=FALSE}
#ap2g
imap(all_seu_prelim_stg_norm, ~FeaturePlot(.x, dims = c(2,3), features = "PF3D7-1222600", order = T, slot = "data")+ labs(title = .y))
imap(all_seu_prelim_stg_norm, ~FeaturePlot(.x, dims = c(1,3), features = "PF3D7-1222600", order = T, slot = "data")+ labs(title = .y))
```

Remove those with too few genes transcripts

## 3.  REMOVALS

###########################START Remove cells######################################
Remove those with too few genes transcripts

```{r}
after_mt_filt <- map(all_seu_prelim_stg_norm, dim)%>% as.data.frame() %>% .[2,]
after_mt_filt
```
## 3.1  OUTLIER REMOVALS

```{r}
## Identify barcodes failing the gene/UMI thresholds (for reporting)
low_count_rm_bcodes <- map(all_seu_prelim_stg_norm, ~colnames(subset(.x, nFeature_RNA > gn_lci_cl_seu & nCount_RNA > rna_lci_cl_seu & nCount_RNA < rna_uci_cl_seu & nFeature_RNA < gn_uci_cl_seu, invert = T)))

```

```{r}

## Apply per-cluster/stage gene and UMI cutoffs to filter out outliers
all_seu_filtOlMtCb <- map(all_seu_prelim_stg_norm, ~subset(.x, nFeature_RNA > gn_lci_cl_seu & nCount_RNA > rna_lci_cl_seu & nCount_RNA < rna_uci_cl_seu & nFeature_RNA < gn_uci_cl_seu))

aftr_low_count_rm <- map(all_seu_filtOlMtCb, dim)%>% as.data.frame() %>% .[2,]
aftr_low_count_rm

```

save to check lost cells from original talleh dataset after outlier removal

```{r, eval=T}
imap(all_seu_filtOlMtCb, ~saveRDS(.x, paste0("data/processed/Pf/", .y,"/seu_obj/all_seu_filtOlMtCb_after_outlier_rm_b4_cbenderqc.RDS")))

```

## 3.1  CELL PROBABILITY REMOVALS

```{r}
map(all_seu_filtOlMtCb, ~.x@meta.data %>% count(cell_probability >= 0.9))
```


```{r}
imap(all_seu_filtOlMtCb, ~.x@meta.data %>% count((cell_probability >= 0.9), stage_fld))

```

```{r}
## Barcodes removed
cell_prob9_rm_bcodes <- map(all_seu_filtOlMtCb, possibly(~colnames(subset(.x, cell_probability >= 0.9, invert = T))))

```

```{r}
all_seu_filtOlMtCb <- map(all_seu_filtOlMtCb, ~subset(.x, cell_probability >= 0.9))


aftr_cell_prob9_rm <- map(all_seu_filtOlMtCb, dim)%>% as.data.frame() %>% .[2,]
aftr_cell_prob9_rm
```


## 3.1  BACKGROUND FRACTION REMOVALS


```{r, warning=FALSE, message=FALSE, fig.width=4, fig.height=4.5}
map(all_seu_filtOlMtCb, ~.x@meta.data %>% count(background_fraction > 0.5  ))

```

```{r}
map(all_seu_filtOlMtCb, ~.x@meta.data %>% count(background_fraction <= 0.5))

```


```{r, fig.height = 3.5, fig.width = 12}
all_seu_filtOlMtCb_mdata_mrg <- map(all_seu_filtOlMtCb, ~.@meta.data) %>% 
  bind_rows(., .id = 'donor') 


```

```{r, warning=FALSE, message=FALSE, fig.width=4, fig.height=4.5}
#
plts_seu_bg5 <- imap(all_seu_filtOlMtCb, possibly(~{
  DimPlot(.x, label = T, 
          cells.highlight =  colnames(
            .x[, !(.x@meta.data$background_fraction <= 0.5 )]
          ), 
          sizes.highlight = 0.05) +   
    theme_void() + 
    theme(legend.position = "none", panel.border = element_rect(color = "black", fill = NA, size = 1))  +
    labs(title = .y)
}, "no cells"))

```

```{r, warning=FALSE, message=FALSE, fig.height= 3.5, fig.width= 10, eval=T}
plot_grid(plotlist = plts_seu_bg5[unlist(map(plts_seu_bg5, ~!is.character(.x)))], ncol = 4)

```


```{r}
## Barcodes removed
cb_bg_frac_rm_bcodes <- map(all_seu_filtOlMtCb, 
                            possibly(~colnames(.x[, !(.x@meta.data$background_fraction <= 0.5 )]
                            )))

```


```{r}
all_seu_filtOlMtCb <- map(all_seu_filtOlMtCb, 
                          ~.x[, (.x@meta.data$background_fraction <= 0.5 )]
)

aftr_cb_bg_frac_rm <- map(all_seu_filtOlMtCb, dim) %>% as.data.frame() %>% .[2,]
aftr_cb_bg_frac_rm
```

###########################END Remove cells######################################


```{r, warning=FALSE, message=FALSE}
#ap2g
imap(all_seu_filtOlMtCb, ~FeaturePlot(.x, dims = c(1,2), features = "PF3D7-1222600", order = T, slot = "counts")+ labs(title = .y))

```


```{r, warning=FALSE, message=FALSE}

imap(all_seu_filtOlMtCb, ~DimPlot(.x, dims = c(1,2), group.by = "stage_fld", pt.size = 0.1)+ scale_color_manual(values = sp_col_new) + labs(title = .y))

```

```{r, warning=FALSE, message=FALSE}

imap(all_seu_filtOlMtCb, ~DimPlot(.x, dims = c(1,2), group.by = "seurat_clusters", pt.size = 0.1, label = T)+ labs(title = .y))

```

```{r, eval=T}
## write out seurat bpcells object files for normalization processing in farm
imap(all_seu_filtOlMtCb, ~saveRDS(.x, paste0("data/processed/Pf/", .y,"/seu_obj/all_seu_filtOlMtCb.RDS")))

```


```{r}

## Compile cell counts after each filtering step for summary and plotting
filt_flow <- list(after_umi100_filt, after_mt_filt, aftr_low_count_rm,  aftr_cell_prob9_rm, aftr_cb_bg_frac_rm) %>% set_names("100umi_filt", "mt_filt", "low_count",  "cell_prob9", "cb_bg_frac") %>% bind_rows(., .id = "filt_metric") %>% mutate(across(filt_metric, ~factor(., levels = c("100umi_filt", "mt_filt", "low_count",  "cell_prob9", "cb_bg_frac"))))

```

```{r}
## Number of cell population removed by respective filter
filt_diff <- (filt_flow[-nrow(filt_flow), ] - filt_flow[-1, ]) 

filt_diff$filt_metric <- paste0(filt_flow[-1, ]$filt_metric, "_rm")
filt_diff_t <- filt_diff  %>% column_to_rownames("filt_metric") %>% t() %>% as.data.frame() 
filt_diff_t

## Proportion of cell population removed by respective filter
filt_diff_prop <- round(filt_diff/filt_flow[-nrow(filt_flow), ] * 100, 2)
filt_diff_prop$filt_metric <- paste0(filt_flow[-1, ]$filt_metric, "_rm_prop")
filt_diff_prop_t <- filt_diff_prop %>% column_to_rownames("filt_metric") %>% t() %>% as.data.frame() 

## Merge number of cells and those that are removed
filt_flow_diff <- bind_rows(
  filt_flow %>% 
    mutate(row_id = row_number(), source = 2),
  filt_diff %>% 
    mutate(row_id = row_number(), source = 2)
) #%>%
# arrange(row_number() + (row_number() > nrow(df)))

filt_flow_diff_t <- filt_flow_diff %>% arrange(row_id, source) %>% 
  select(-row_id, -source) %>% column_to_rownames("filt_metric") %>% t() %>% as.data.frame() 

filt_flow_diff_t
```


```{r}
## Save filter-flow summary table for record-keeping
# saveRDS(filt_flow_diff_t, "data/processed/Pf/m2_all/filt_flow_diff_t.RDS")
saveRDS(filt_flow_diff_t, "data/processed/Pf/m2_all/filt_flow_diff_t.RDS") 

```

```{r}

filt_flow %>% 
  pivot_longer(cols = contains("MSC"), names_to = "donor", values_to = "cell_counts") %>%  
  ggplot(aes(x = filt_metric, y = cell_counts, colour = donor, group = donor)) + geom_point() + geom_line()

```


```{r}
## Aggregate lists of removed barcodes by filter for downstream plotting/saving
poorcellQC_bcodes_rm <- list(mt_filt_rm_bcodes, low_count_rm_bcodes, cell_prob9_rm_bcodes, cb_bg_frac_rm_bcodes) %>% 
  set_names(c("mt_filt_rm_bcodes", "low_count_rm_bcodes", "cell_prob9_rm_bcodes", "cb_bg_frac_rm_bcodes"))

```


```{r, message=FALSE, warning=FALSE, eval=T}
# Save the poorQC cells barcodes for plotting
saveRDS(poorcellQC_bcodes_rm, "data/processed/Pf/m2_all/lowQC_rm_plotting/poorcellQC_bcodes_rm.RDS")
```

!!!NOTE - If the below breaks the code then just remove and rerun. I added it but didn't rerun
```{r, warning=FALSE, message=FALSE}

# UMAP + metadata for filtered objects
all_seu_filtOlMtCb_umap_mdata <- get_umap_mdata(all_seu_filtOlMtCb)
```


```{r, message=FALSE, warning=FALSE, eval=T}
# Save the poorQC cells barcodes for plotting
saveRDS(all_seu_filtOlMtCb_umap_mdata, "data/processed/Pf/m2_all/lowQC_rm_plotting/all_seu_filtOlMtCb_umap_mdata.RDS")

```

