---
title: "MSC QC and preprocessing"
date: '2022-07-11'
author: 'Jesse Rop'
output:
  html_document:
    toc: true
    theme: united
    code_folding: hide
    df_print: paged
---

```{r}
suppressPackageStartupMessages({
  library(Seurat)
  library(sctransform)
  library(glmGamPoi)    # used by sctransform internals (kept)
  library(ggrepel)
  library(dplyr) 
  library(stringr)
  library(purrr)
  library(tibble)
  library(ggplot2)
  library(tidyr)
  library(cowplot)
  library(plotly)
  library(Hmisc)
  library(rmarkdown)
  library(conflicted)
  library(SingleCellExperiment)
  library(scDblFinder)
  library(findPC)


  # NOTE: packages commented below are possibly unused in this Rmd.
  # Verify across sourced files/project before permanently removing them.
  # Packages: patchwork, ggpubr, readr, SeuratData, SeuratDisk,
  #           RColorBrewer, ggvenn, MetBrewer, ggh4x, BPCells,
  #           parallel, EnhancedVolcano, DropletUtils, ggforce,
  #           scran, SingleR, ggsankey, findPC

  conflict_prefer("select", "dplyr")
  conflict_prefer("filter", "dplyr")
  conflict_prefer("rename", "dplyr")
  conflict_prefer("map", "purrr")
  conflict_prefer("count", "dplyr")
  conflicts_prefer(base::saveRDS)
  conflicts_prefer(SingleCellExperiment::colData)
  conflicts_prefer(base::as.data.frame)
  conflicts_prefer(base::intersect)
  conflicts_prefer(dplyr::desc)
  conflicts_prefer(dplyr::count)
})
```



# MSC stage annotation and gene expression QC 

## setup, read input files and exploratory visualizations and general functions

```{r, setup}
knitr::opts_knit$set(root.dir = '/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/Talleh')

```

### Variable declarations


#### Source common functions, variables and tables
```{r}
##Boxplot labeling function
source("~/Talleh_code/pf_common_vars.R")
source("~/Talleh_code/project_common_fns.R")

```

```{r}
## Helper functions (small, well-commented)

# extract_umap_mdata_single
# - Input: a Seurat object with a `reductions$umap` embedding
# - Output: a tibble joining `meta.data` with the UMAP coordinates and a `bcode` column
extract_umap_mdata_single <- function(obj){
  obj@meta.data %>%
    rownames_to_column("bcode") %>%
    left_join(., obj@reductions$umap@cell.embeddings %>% data.frame() %>% rownames_to_column("bcode"), by = "bcode")
}

# extract_umap_mdata
# - Input: a named list of Seurat objects
# - Output: a list of tibbles produced by `extract_umap_mdata_single`
extract_umap_mdata <- function(obj_list){
  purrr::map(obj_list, extract_umap_mdata_single)
}

```


## Read in sample name details

```{r}
# Select sample names
sample_name <- c("MSC49", "MSC68")

```

## DOUBLET REMOVAL

```{r, message=F, warning=F, eval=T}
## Get mdata file list
all_seu_filtOlMtCb_list_files <- Sys.glob("data/processed/Pf/*/seu_obj/all_seu_filtOlMtCb.RDS")

all_seu_filtOlMtCb_list_files
```

```{r, message=F, warning=F, eval=T}
## read filtered barcodes
all_seu_filtOlMtCb <- map(all_seu_filtOlMtCb_list_files, ~.x %>% 
                                set_names(nm = (str_remove_all(.,'data/processed/Pf/|/seu_obj.*')))) %>% 
  unlist(., recursive = F) %>% 
  .[sample_name] %>%
  .[!is.na(.)] %>%
  map(readRDS) 

```

```{r, eval=T}
all_seu_filtOlMtCb_sct <- map(all_seu_filtOlMtCb, ~sct_run_fn(Seurat_object = .x))
all_seu_filtOlMtCb_sct
```

```{r, warning=FALSE, message=FALSE}
imap(all_seu_filtOlMtCb_sct, ~DimPlot(.x, dims = c(2,3), group.by = "stage_fld", pt.size = 0.1, reduction = "pca")+ scale_color_manual(values = sp_col_new) + labs(title = .y))
imap(all_seu_filtOlMtCb_sct, ~DimPlot(.x, dims = c(1,2), group.by = "stage_fld", pt.size = 0.1)+ scale_color_manual(values = sp_col_new) + labs(title = .y))
```

```{r, warning=FALSE, message=FALSE}

all_seu_filtOlMtCb_sct_umap_mdata <- extract_umap_mdata(all_seu_filtOlMtCb_sct)
```


```{r, message=FALSE, warning=FALSE, eval=T}
# Save UMAP metadata (used for plotting and QC)
saveRDS(all_seu_filtOlMtCb_sct_umap_mdata, "data/processed/Pf/m2_all/all_seu_filtOlMtCb_sct_umap_mdata.RDS")

```

```{r, warning=FALSE, message=FALSE, fig.width=12, fig.height=3.5}
## Choose sct or standard normalization
# all_seu_filtOlMtCb_slct <- all_seu_filtOlMtCb_norm
all_seu_filtOlMtCb_slct <- all_seu_filtOlMtCb_sct

for (i in 1:length(all_seu_filtOlMtCb_slct)) {
  DefaultAssay(all_seu_filtOlMtCb_slct[[i]]) <- "RNA"
  all_seu_filtOlMtCb_slct[[i]] <- NormalizeData(all_seu_filtOlMtCb_slct[[i]])
  all_seu_filtOlMtCb_slct[[i]] <- ScaleData(all_seu_filtOlMtCb_slct[[i]], features = rownames(all_seu_filtOlMtCb_slct[[i]]))
}
```

#### Incorporate and visualize strain - souporcell (SC)
We did souporcell assuming a range of 3 strains (6,7 and 8)
```{r, eval=T}
algn_nm <- "hsat"
# algn_nm <- "minmap"

species <- "Pf"
sv_dir <- "pbulk_gtypes_preQC_resc_hspf"
# sv_dir <- "pbulk_gtypes_preQC_resc_hspf3"

sample_name_nms <- sample_name
##Read in strain information
soupo_kselect_list_pp <-  readRDS(paste0("data/processed/", species, "/" , sv_dir, "/strn_c_new_mdata_m22_no_stage.RDS"))  %>% .[paste(sample_name_nms, algn_nm, sep = "_")]

##Creating Binary variable for Doublet/Singlet for doublet finder in the samples with more than 2 strains
soupo_kselect_list_pp <- soupo_kselect_list_pp %>%
  map(. %>%
        mutate(strain_gt = case_when(dbt_neg_all == "DoubletAll" ~ 'Doublet',
                                     dbt_neg_all == "LikelySinglet" ~ 'Singlet', 
                                     dbt_neg_all == "DbtNegMix" ~ 'Singlet',
                                     dbt_neg_all == "NegativeAll" ~ 'Negative'),
               GT = case_when(dbt_neg_all == "DoubletAll" ~ 'Doublet', 
                              T ~  'Singlet')))

```

```{r}
## NOTE !! - We already think that donor M49 has five strains and M68 has 3 strains so we use that
optimum_k <- c(5, 3)
soupo_kselect_list_pp <- pmap(list(soupo_kselect_list_pp, names(soupo_kselect_list_pp), optimum_k), ~..1 %>% mutate(Strain = !!rlang::sym(paste0("Strain_hmo_", str_remove(str_extract(..2, "M.*[0-9]+"), "SC"), "_", algn_nm,"_", ..3))))

map(soupo_kselect_list_pp, ~.x %>% count(Strain, Strain_DN))
```


```{r, message=FALSE, warning=FALSE, eval=T}
## Save for comparison with m49 old
saveRDS(soupo_kselect_list_pp, "data/processed/Pf/m2_all/soupo_kselect_list_pp.RDS")
```


```{r, eval=T}

all_seu_filtOlMtCb_slct <- all_seu_filtOlMtCb_slct %>%
  map2(., soupo_kselect_list_pp, ~AddMetaData(.x, .y %>% column_to_rownames("bcode") %>%select(strain_gt, GT, Strain_DN, Strain)))
```

```{r, message=FALSE, warning=FALSE, eval=T}
## Save for comparison with m49 old
saveRDS(map(all_seu_filtOlMtCb_slct, ~.x@meta.data), "data/processed/Pf/m2_all/all_seu_filtOlMtCb_slct_w_strain_mdata.RDS")
```

```{r}
# imap(all_seu_filtOlMtCb_slct, possibly(~DimPlot(.x, group.by = "Strain_DN", reduction = "umap") + labs(title = .y) + scale_color_manual(values = strain_cols_n)))
```

#####################################START - sdblfinder #################################################################
## scDblFinder

!!! IMPORTANT: this block is farm-run code and required by downstream automation — do not delete
```{r, eval=T}
## scDblFinder Doublet removal

scdblf_fn <- function(dblt_filt, HVG = 700){
  
  ### Seurat to SCE
  DefaultAssay(dblt_filt) <- "RNA"
  dblt_sce <- SingleCellExperiment(assays = list(counts = as(object = dblt_filt[["RNA"]]$counts, Class = "dgCMatrix")), colData = dblt_filt@meta.data)
  
  ### random - if cluster is only one use random
  dblt_rand <- scDblFinder(dblt_sce, returnType = "table", nfeatures = HVG)
  
  ### cluster - if cluster is only one use random
  dblt_clst <- if(length(unique(dblt_sce$seurat_clusters)) > 1) { scDblFinder(dblt_sce, clusters = "seurat_clusters", returnType = "table", nfeatures = HVG) } else { dblt_rand }
  
  ## Calculate doublets using scDblFinder for soupx adjusted counts - soup correction seems to improve doublet detection based on Elias MCA analysis
  
  
  ## Put the different iterations together
  scf_outputs <- pmap(list(list(dblt_clst, dblt_rand), 
                    list("_norm", "_rand")), ~ {
        tbl <- data.frame(..1[,c("score", "class", "type")]) %>% 
          filter(type == "real") %>%
          select(score, class) %>%
          mutate(across(class, str_to_sentence)) 
        
        names(tbl) <- paste0("scf_", names(tbl), ..2)
        tbl <- tbl %>% 
          rownames_to_column("bcode")
         }) %>%
        purrr::reduce(., .f = full_join, by = "bcode")
  
  return(scf_outputs)
  
}

```

```{r, message=F, warning=F, eval=T}
## Run scdblfinder within the notebook - quite slow
scf_outputs <- map(all_seu_filtOlMtCb_slct, ~scdblf_fn(dblt_filt = .x, HVG = 700))
scf_outputs
```

Assess agreement between the calls
```{r, message=FALSE}
map(scf_outputs, ~.x %>% select(contains("class")) %>% group_by_all() %>% count())

```

Chsqr of the different scdblfinder flavours tests
```{r, message=FALSE}
pw_scf_comp <- scf_outputs[[1]] %>% select(contains("class")) %>% colnames() %>% combn(., 2, simplify = F)

pw_scf_comp_nms <- map(pw_scf_comp, ~{paste0(.x[[1]], '_VS_', .x[[2]])})

a <- map(scf_outputs, ~{
  tbl = .x
  map(pw_scf_comp, ~{chisq.test(tbl[,.x[[1]]], tbl[,.x[[2]]])[c('statistic', 'p.value')]} %>% base::as.data.frame() %>%  rownames_to_column('test') ) %>%  
    set_names(pw_scf_comp_nms) %>% 
    bind_rows(., .id = "comps")
}) %>% 
  bind_rows(., .id = "donor")

```



#####################################END - sdblfinder #################################################################


```{r}
all_seu_filtOlMtCb_w_dblt <- list(scf_outputs, all_seu_filtOlMtCb_slct) %>%
  pmap(~ ..1 %>%
         column_to_rownames('bcode') %>%
         AddMetaData(..2, .))


```


```{r}
# rm(all_seu_filtOlMtCb_slct)
```

```{r, message=FALSE}
## count old scdbl classifications
map(all_seu_filtOlMtCb_w_dblt, ~.x@meta.data %>% count(scf_class_norm, scf_class_rand))
```


```{r}
all_seu_filtOlMtCb_w_dblt
```

Add the doublet proportion per cluster for each of the doublet methods assessed

```{r, warning=FALSE, message=FALSE, fig.width=14, fig.height= 4}
# dbt_methods <- c(paste0(c("scf_class"), c("_norm", "_rand")), paste0(c("scf_class"), c("_supx_norm", "_supx_rand")), "scrub_doublet", "strain_gt")
dbt_methods <- c(paste0(c("scf_class"), c("_norm", "_rand")))


all_seu_filtOlMtCb_w_dblt <- map(all_seu_filtOlMtCb_w_dblt, ~{
  obj = .x
  map(dbt_methods, ~{
    obj@meta.data %>%
      rownames_to_column('bcode') %>%
      dplyr::add_count(seurat_clusters, !!rlang::sym(.x))  %>%
      group_by(seurat_clusters) %>%
      mutate(freq = round(n / n(), 3), 
             !!paste0(.x, "_dbt_nums") := case_when(!!rlang::sym(.x) == "Doublet" ~ n,
                                                    !!rlang::sym(.x) == "Singlet" ~ n()-n),
             !!paste0(.x, "_dbt_prop") := case_when(!!rlang::sym(.x) == "Doublet" ~ freq,
                                                    !!rlang::sym(.x) == "Singlet" ~ 1-freq),
             !!paste0(.x, "_dbt_ovrl_prop") := case_when(!!rlang::sym(.x) == "Doublet" ~ round(n / ncol(obj), 3),
                                                         !!rlang::sym(.x) == "Singlet" ~ round((n()-n) / ncol(obj), 3)),
             !!paste0(.x, "_dbt_allnums") := case_when(!!rlang::sym(.x) == "Doublet" ~ paste(n, n(), ncol(obj), sep = "_"),
                                                       !!rlang::sym(.x) == "Singlet" ~ paste((n()-n), n(), ncol(obj), sep = "_"))
      ) %>%
      ungroup() %>% 
      select('bcode', paste0(.x, "_dbt_prop"), paste0(.x, "_dbt_ovrl_prop"), paste0(.x, "_dbt_nums"), paste0(.x, "_dbt_allnums")) 
    
  }) %>%
    purrr::reduce(., .f = full_join, by = "bcode") %>% 
    column_to_rownames('bcode') %>%
    AddMetaData(obj, .)
})

```



```{r, fig.height = 3.5, fig.width = 15}
dbt_methods_gtyp <- dbt_methods

scf_fn <- function(obj = all_seu_filtOlMtCb_w_dblt[[1]], mks = dbt_methods_gtyp, mk_nms = dbt_methods_gtyp, nrws = ceiling(length(dbt_methods_gtyp)/5), redct = "umap", don_nm) {
  list(mks, mk_nms) %>%
    pmap(.,~{
      DimPlot(obj, dims = c(1,2), label = T, pt.size = 0.01, cells.highlight = obj@meta.data %>% filter(!!rlang::sym(..1) == "Doublet") %>% row.names(), sizes.highlight = 0.1, reduction = redct) + 
        labs(title = paste(don_nm, ..2)) +
        theme(legend.position = 'none')
    }) %>% 
    plot_grid(plotlist = ., nrow = nrws) 
}

```

```{r, fig.height = 4, fig.width = 15}
imap(all_seu_filtOlMtCb_w_dblt, ~scf_fn(obj = .x, nrws = ceiling(length(dbt_methods_gtyp)/5), redct = "pca", don_nm = .y))
```


```{r, fig.height = 4, fig.width = 15}
imap(all_seu_filtOlMtCb_w_dblt, ~scf_fn(obj = .x, nrws = ceiling(length(dbt_methods_gtyp)/5), redct = "umap", don_nm = .y))
```


##### Check doublet clusters
```{r}
map(all_seu_filtOlMtCb_w_dblt, ~.x@meta.data %>% filter(scf_class_norm_dbt_prop > 0.1) %>% count(seurat_clusters, scf_class_norm_dbt_prop,scf_class_rand_dbt_prop,  scf_class_norm_dbt_nums))

```

```{r, warning=FALSE, message=FALSE, fig.width=4, fig.height=4.5}
## in cellbender but not in edrops
dbt_props <- c("scf_class_norm_dbt_prop", "scf_class_rand_dbt_prop")
```


```{r, warning=FALSE, message=FALSE, fig.width=4, fig.height=4.5}
#
plts_dbt_scf <- map(dbt_props, ~{
  mthd <- .x
  plt = imap(all_seu_filtOlMtCb_w_dblt, possibly(~DimPlot(.x, label = T, cells.highlight =  colnames(.x[,.x@meta.data[,mthd] >= 0.5]), sizes.highlight = 0.05) + 
                                                   theme_void() + theme(legend.position = "none", panel.border = element_rect(color = "black", fill = NA, size = 1))  +
                                                   labs(title = .y), "no cells"))
  return(plt)
})

```

```{r, warning=FALSE, message=FALSE, fig.height= 3.5, fig.width= 10, eval=T}
map(plts_dbt_scf, ~plot_grid(plotlist = .x, ncol = 4))
```


```{r, warning=FALSE, message=FALSE, fig.width=4, fig.height=4.5}
#
plts_dbt_scf_rm <- imap(all_seu_filtOlMtCb_w_dblt, 
                        possibly(~DimPlot(.x, label = T, cells.highlight =  colnames(.x[,.x@meta.data[,"scf_class_norm"] == "Doublet"]), sizes.highlight = 0.05) + 
                                   theme_void() + 
                                   theme(legend.position = "none", panel.border = element_rect(color = "black", fill = NA, size = 1))  +  labs(title = .y), "no cells"))


```

```{r, warning=FALSE, message=FALSE, fig.height= 3.5, fig.width= 10, eval=T}
plot_grid(plotlist = plts_dbt_scf_rm, ncol = 4)
```

```{r, warning=FALSE, message=FALSE, fig.width=4, fig.height=4.5}
#
plts_dbt_scf_p_rm <- imap(all_seu_filtOlMtCb_w_dblt, 
                          possibly(~DimPlot(.x, label = T, 
                                            cells.highlight =  colnames(.x[,(.x@meta.data[,"scf_class_norm_dbt_prop"] >= 0.5 | 
                                                                               .x@meta.data[,"scf_class_rand_dbt_prop"] >= 0.5) #| 
                                                                           #.x@meta.data[,"scrub_doublet_dbt_prop"] >= 0.2]
                                            ]), 
                                            sizes.highlight = 0.05) + 
                                     theme_void() + theme(legend.position = "none", panel.border = element_rect(color = "black", fill = NA, size = 1))  +
                                     labs(title = .y), "no cells"))


```

```{r, warning=FALSE, message=FALSE, fig.height= 3.5, fig.width= 10, eval=T}
plot_grid(plotlist = plts_dbt_scf_p_rm, ncol = 4)
```




##### Remove remaining doublets
```{r}
map(all_seu_filtOlMtCb_w_dblt, ~.@meta.data %>% dplyr::count(scf_class_norm))
```
##### Check doublet clusters
```{r}
map(all_seu_filtOlMtCb_w_dblt, ~.x@meta.data %>% count(scf_class_norm_dbt_prop > 0.1))

```


```{r}
map(all_seu_filtOlMtCb_w_dblt, ~.x@meta.data %>% filter(!((scf_class_norm_dbt_prop > 0.5 & scf_class_rand_dbt_prop > 0.5) )) %>% rownames_to_column("bcode") %>% mutate(across(c(scf_class_norm_dbt_prop,scf_class_rand_dbt_prop), ~round(., 2))) %>% count(seurat_clusters, scf_class_norm_dbt_prop,scf_class_rand_dbt_prop))

```

```{r}
map(all_seu_filtOlMtCb_w_dblt, ~.x@meta.data %>% filter(scf_class_norm == "Singlet") %>% count(scf_class_norm_dbt_prop < 0.20 & scf_class_rand_dbt_prop < 0.20))

```

```{r}
map(all_seu_filtOlMtCb_w_dblt, ~.x@meta.data %>% filter(scf_class_norm == "Singlet" & scf_class_norm_dbt_prop < 0.20 & scf_class_rand_dbt_prop < 0.20))
```


```{r}
map(all_seu_filtOlMtCb_w_dblt, ~.x@meta.data %>% filter(scf_class_norm == "Singlet") %>% count(scf_class_norm_dbt_prop, scf_class_rand_dbt_prop) %>% filter(scf_class_norm_dbt_prop>0.5 ))
```


```{r}
map(all_seu_filtOlMtCb_w_dblt, ~.x@meta.data %>%  count(scf_class_norm_dbt_prop > 0.20 & scf_class_rand_dbt_prop > 0.20))
```


```{r, warning=FALSE, message=FALSE, fig.width=4, fig.height=4.5}
#
plts_dbt_rm <- imap(all_seu_filtOlMtCb_w_dblt, possibly(~DimPlot(.x, label = T, cells.highlight =  colnames(.x[,(.x@meta.data[,"scf_class_norm_dbt_prop"] >= 0.2 & .x@meta.data[,"scf_class_rand_dbt_prop"] >= 0.2) | .x@meta.data[,"scf_class_norm"] != "Singlet"]), sizes.highlight = 0.05) + 
                                                          theme_void() + theme(legend.position = "none", panel.border = element_rect(color = "black", fill = NA, size = 1))  +
                                                          labs(title = .y), "no cells"))


```

```{r, warning=FALSE, message=FALSE, fig.height= 3.5, fig.width= 10, eval=T}
plot_grid(plotlist = plts_dbt_rm, ncol = 4)
```
##### Remove remaining doublets



```{r}
##number of unfilt
raw_n <- map(all_seu_filtOlMtCb_w_dblt, dim) %>% as.data.frame()
```

```{r}
## Barcodes removed
scf_dblts_rm_bcodes <- map(all_seu_filtOlMtCb_w_dblt, possibly(~colnames(subset(., subset = scf_class_norm == "Singlet" | (scf_class_norm_dbt_prop < 0.20 | scf_class_rand_dbt_prop < 0.20), invert = T))))
```

```{r}
##Seurat object
##Remove stage doublets

all_filt_cb_no_ScfStgDblts <- all_seu_filtOlMtCb_w_dblt %>%
  map(. %>% subset(., subset = scf_class_norm == "Singlet")) %>%
  map(. %>% subset(., subset = scf_class_norm_dbt_prop < 0.20 | scf_class_rand_dbt_prop < 0.20)) 


scf_dblts_rm <- map(all_filt_cb_no_ScfStgDblts, dim) %>% as.data.frame() %>%.[2,]
scf_dblts_rm
```

```{r, message=FALSE, warning=FALSE, eval=T}
## Save for comparison with m49 old
saveRDS(map(all_filt_cb_no_ScfStgDblts, ~.x@meta.data), "data/processed/Pf/m2_all/all_filt_cb_no_ScfStgDblts_mdata.RDS")
```


```{r}
imap(all_filt_cb_no_ScfStgDblts, possibly(~DimPlot(.x, cells.highlight = colnames(.x[,.x@meta.data$scf_class_rand == "Doublet"]), reduction = "umap") + labs(title = .y)))
```


```{r}
map(all_filt_cb_no_ScfStgDblts, ~.x@meta.data %>% count(scf_class_norm))
```


```{r}
map(all_filt_cb_no_ScfStgDblts, ~.x@meta.data %>% count( scf_class_norm_dbt_prop, scf_class_rand_dbt_prop) )

```


```{r, warning=FALSE, message=FALSE}
#ap2g
imap(all_filt_cb_no_ScfStgDblts, ~FeaturePlot(.x, dims = c(1,2), features = "PF3D7-1222600", order = T, slot = "data")+ labs(title = .y))
```

```{r}
imap(all_filt_cb_no_ScfStgDblts, possibly(~DimPlot(.x, group.by = "stage_fld") + labs(title = .y) + scale_color_manual(values = sp_col_new))) 
imap(all_filt_cb_no_ScfStgDblts, possibly(~DimPlot(.x, group.by = "stage_fld", dims = c(1,3)) + labs(title = .y) + scale_color_manual(values = sp_col_new))) 

```

```{r, warning=FALSE, message=FALSE, fig.width=12, fig.height=3.5}
#SURF1.2 gexp02
imap(all_filt_cb_no_ScfStgDblts, ~FeaturePlot(.x, reduction = "umap",  label = T, features = c("PF3D7-0113600", "PF3D7-1102500", "PF3D7-1222600"), order = T, ncol = 3) + labs(title = .y))
```

```{r}
## !! NOTE - Skip scrublet for now since based on histograms and plot metrics it seems more cells than necessary are being removed
## Maybe try solo or some other doublet finding method
all_filt_cb_no_StgDblts <- all_filt_cb_no_ScfStgDblts

scrub_dblts <- map(all_filt_cb_no_StgDblts, dim)%>% as.data.frame() %>% .[2,] #%>% t()
scrub_dblts 
```



```{r}
doublets_bcodes_rm <- list(scf_dblts_rm_bcodes) %>% set_names(c("scf_dblts_rm_bcodes"))
```


```{r, message=FALSE, warning=FALSE, eval=T}
# Save the poorQC cells barcodes for plotting
saveRDS(doublets_bcodes_rm, "data/processed/Pf/m2_all/lowQC_rm_plotting/doublets_bcodes_rm.RDS")
```


################ START strain removal readded ################ 

```{r, warning=FALSE, message=FALSE, fig.width=4, fig.height=4.5}
#
plts_dbt_strn <- imap(all_filt_cb_no_StgDblts, 
                      possibly(~DimPlot(.x, label = T, 
                                        cells.highlight =  colnames(.x[,.x@meta.data[,"strain_gt"] == "Doublet"]), 
                                        sizes.highlight = 0.05) + 
                                 theme_void() + 
                                 theme(legend.position = "none", 
                                       panel.border = element_rect(color = "black", fill = NA, size = 1))  +  
                                 labs(title = .y), "no cells"))


```

```{r, warning=FALSE, message=FALSE, fig.height= 3, fig.width= 8, eval=T}
plot_grid(plotlist = plts_dbt_strn, ncol = 4)
```


```{r}
map(all_filt_cb_no_StgDblts, ~.x@meta.data  %>% 
      count(seurat_clusters, strain_gt == "Doublet") %>% 
      group_by(seurat_clusters) %>%
      mutate(freq = round(n / sum(n), 3)) %>% 
      filter(`strain_gt == "Doublet"` ==T & freq > 0.25))

```


```{r}
## Barcodes removed - 
# str_dbt_rm_bcodes <- map(all_filt_cb_no_StgDblts, possibly(~colnames(subset(.x, strain_gt != "Doublet", invert = T))))
str_dbt_rm_bcodes <- map(all_filt_cb_no_StgDblts, possibly(~colnames(subset(.x, Strain != "Doublet", invert = T))))
```


```{r}
## ¡¡¡¡¡¡ NOTE - STRAIN doublet removal Needs to move to another tab - Removing all doublets called by k=5 since this has already been determined
all_cb_filt_dbt <- map(all_filt_cb_no_StgDblts, ~subset(.x, Strain != "Doublet"))

str_dbt_rm <- map(all_cb_filt_dbt, dim) %>% as.data.frame() %>% .[2,] 
str_dbt_rm
```

```{r, message=FALSE, warning=FALSE, eval=T}
## Save for comparison with m49 old
saveRDS(map(all_cb_filt_dbt, ~.x@meta.data), "data/processed/Pf/m2_all/all_cb_filt_dbt_mdata.RDS")
```


```{r}
## Barcodes removed - 
strn_neg_rm_bcodes <- map(all_cb_filt_dbt, possibly(~colnames(subset(.x, Strain != "Negative", invert = T))))
```


```{r}
## ¡¡¡¡¡¡ NOTE - STRAIN negative removal Needs to move to another tab - Removing all doublets called by k=5 since this has already been determined
all_cb_filt_dbt_neg <- map(all_cb_filt_dbt, ~subset(.x, Strain != "Negative"))

strn_neg_rm <- map(all_cb_filt_dbt_neg, dim) %>% as.data.frame() %>% .[2,] 
strn_neg_rm
```

```{r, message=FALSE, warning=FALSE, eval=T}
## Save for comparison with m49 old
saveRDS(map(all_cb_filt_dbt_neg, ~.x@meta.data), "data/processed/Pf/m2_all/all_cb_filt_dbt_neg_mdata.RDS")
```


```{r}
doublets_bcodes_rm <- list(scf_dblts_rm_bcodes, str_dbt_rm_bcodes, strn_neg_rm_bcodes) %>% set_names(c("scf_dblts_rm_bcodes", "str_dbt_rm_bcodes", "strn_neg_rm_bcodes"))
```


```{r, message=FALSE, warning=FALSE, eval=T}
# Save the poorQC cells barcodes for plotting
saveRDS(doublets_bcodes_rm, "data/processed/Pf/m2_all/lowQC_rm_plotting/doublets_bcodes_rm.RDS")
```

```{r}
# gogo()
## scrublet inputs
```

```{r}
# rm(all_filt_cb_no_StgDblts)
```
################ END strain removal readded ################ 


Check to remove any left stubborn clusters with high background fraction and poor stage annotation

```{r, warning=FALSE, message=FALSE, fig.width=4, fig.height=4.5}
## in cellbender but not in edrops
plts_stg_no_dbt <- imap(all_cb_filt_dbt_neg, possibly(~{
  cell_stg_plot <- DimPlot(.x, group.by = "stage_fld", cols = sp_col_new, pt.size = 0.2) + 
    theme_void() + 
    theme(legend.position = "none", panel.border = element_rect(color = "black", fill = NA, size = 1))  +
    labs(title = .y)
  
  cell_stg_plot$data <- cell_stg_plot$data %>% rownames_to_column('bcode') %>% left_join(., .x@meta.data[,'seurat_clusters', drop = F] %>% rownames_to_column('bcode'), by='bcode') %>% column_to_rownames('bcode')
  
  cell_stg_plot<- LabelClusters(cell_stg_plot, id = "seurat_clusters", size = 6, repel = T)
  return(cell_stg_plot)
  
}, "no cells"))

```

```{r, warning=FALSE, message=FALSE, fig.height= 3.5, fig.width= 10, eval=T}
plot_grid(plotlist = plts_stg_no_dbt, ncol = 4)

```



```{r, warning=FALSE, message=FALSE, fig.width=4, fig.height=4.5}
## in cellbender but not in edrops
plts_sc_dbts <- imap(all_cb_filt_dbt_neg, possibly(~{
  DimPlot(.x, label = T, group.by = "seurat_clusters", pt.size = 0.1) + 
    theme_void() + theme(legend.position = "none", panel.border = element_rect(color = "black", fill = NA, size = 1))  +
    labs(title = .y)}, "no cells"))

```

```{r, warning=FALSE, message=FALSE, fig.height= 3.5, fig.width= 10, eval=T}
plot_grid(plotlist = plts_sc_dbts, ncol = 4)
```

```{r}

filt_flow_dbt <- list(scf_dblts_rm, str_dbt_rm) %>% set_names("scdblfinder", "strain") %>% bind_rows(., .id = "filt_metric") %>% mutate(across(filt_metric, ~factor(., levels = c("scdblfinder", "strain"))))

```

```{r, eval=F}
## Number of cell population removed by respective filter
# filt_diff <- (filt_flow_dbt[-nrow(filt_flow_dbt), ] - filt_flow_dbt[-1, ]) 
filt_diff <- filt_flow_dbt


filt_diff$filt_metric <- paste0(filt_flow_dbt[-1, ]$filt_metric, "_rm")
filt_diff_t <- filt_diff  %>% column_to_rownames("filt_metric") %>% t() %>% as.data.frame() 
filt_diff_t

## Proportion of cell population removed by respective filter
filt_diff_prop <- round(filt_diff/filt_flow_dbt[-nrow(filt_flow_dbt), ] * 100, 2)
filt_diff_prop$filt_metric <- paste0(filt_flow_dbt[-1, ]$filt_metric, "_rm_prop")
filt_diff_prop_t <- filt_diff_prop %>% column_to_rownames("filt_metric") %>% t() %>% as.data.frame() 

## Merge number of cells and those that are removed
filt_flow_dbt_diff <- bind_rows(
  filt_flow_dbt %>% 
    mutate(row_id = row_number(), source = 2),
  filt_diff %>% 
    mutate(row_id = row_number(), source = 2)
) #%>%
# arrange(row_number() + (row_number() > nrow(df)))

filt_flow_dbt_diff_t <- filt_flow_dbt_diff %>% arrange(row_id, source) %>% 
  select(-row_id, -source) %>% column_to_rownames("filt_metric") %>% t() %>% as.data.frame() 

filt_flow_dbt_diff_t
```


```{r, eval=F}
saveRDS(filt_flow_dbt_diff_t, "data/processed/Pf/m2_all/filt_flow_dbt_diff_t.RDS") 

```

```{r}

filt_flow_dbt 

```

```{r, eval=T}
all_cb_no_dbt <- all_cb_filt_dbt_neg

```

```{r, eval=T}
## write out seurat bpcells object files for normalization processing in farm
### Cbender filt
imap(all_cb_no_dbt, ~saveRDS(.x, paste0("data/processed/Pf/",.y,"/seu_obj/all_cb_no_dbt.RDS")))

```



```{r}
all_cb_no_dbt[["MSC49"]]@meta.data %>% count(scf_class_norm, scf_class_rand)

```

```{r, eval=T}
## Run SCT
all_cb_no_dbt_sct <- map(all_cb_no_dbt, ~sct_run_fn(Seurat_object = .x))
all_cb_no_dbt_sct
```



```{r, warning=FALSE, message=FALSE}
# Example: AP2-G feature plot uses this UMAP metadata (AP2-G = PF3D7-1222600)
all_cb_no_dbt_sct_umap_mdata <- extract_umap_mdata(all_cb_no_dbt_sct)

```


```{r, message=FALSE, warning=FALSE, eval=T}
# Save UMAP metadata (clean) for plotting and downstream checks
saveRDS(all_cb_no_dbt_sct_umap_mdata, "data/processed/Pf/m2_all/lowQC_rm_plotting/all_cb_no_dbt_sct_umap_mdata.RDS")

```

```{r, warning=FALSE, message=FALSE}
# go()
```

```{r, warning=FALSE, message=FALSE}
imap(all_cb_no_dbt_sct, ~DimPlot(.x, dims = c(1,2), group.by = "stage_fld", pt.size = 0.1, reduction = "pca")+ scale_color_manual(values = sp_col_new) + labs(title = .y))
imap(all_cb_no_dbt_sct, ~DimPlot(.x, dims = c(1,2), group.by = "stage_fld", pt.size = 0.1)+ scale_color_manual(values = sp_col_new) + labs(title = .y))
```

```{r, warning=FALSE, message=FALSE}
imap(all_cb_no_dbt_sct, ~DimPlot(.x, dims = c(1,2), group.by = "stage_fld", pt.size = 0.1, reduction = "pca")+ scale_color_manual(values = sp_col_new) + labs(title = .y))
imap(all_cb_no_dbt_sct, ~DimPlot(.x, dims = c(1,2), group.by = "stage_fld", pt.size = 0.1)+ scale_color_manual(values = sp_col_new) + labs(title = .y))
```



```{r, eval=T}
## write out seurat bpcells object files for normalization processing in farm
### Cbender filt
imap(all_cb_no_dbt_sct, ~saveRDS(.x, paste0("data/processed/Pf/",.y,"/seu_obj/all_cb_no_dbt_sct.RDS")))

```