---
title: "VCF Variant QC — cluster genotypes"
date: '`r Sys.Date()`'
author: 'Jesse Rop'
output:
  html_document:
    toc: true
    theme: united
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/Talleh")
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

## Overview

Reads a souporcell cluster-genotype VCF, annotates variants against the Pf7 Mali
reference panel, and retains a clean set of informative SNPs for downstream strain
deconvolution.

**Input FORMAT fields (souporcell):** `GT:AO:RO:T:E:GO:GN`
(`DP`/`QR`/`QA` FORMAT fields are defined in the VCF header but absent from variant rows;
per-sample read counts are taken from `RO` and `AO`.)

To knit from the command line:
```{bash knit-cmd, eval=FALSE}
module load HGI/softpack/groups/team222/Pf_scRNAseq/34
# Edit the source() path in the params chunk to select the desired config, then:
Rscript -e "rmarkdown::render('Talleh_code/vcf_variant_qc.Rmd')"
```

---

## Libraries

```{r libs}
# Add softpack R library path so packages resolve whether or not the module
# was loaded before this R session started
.libPaths(c("/opt/view/rlib/R/library", .libPaths()))

suppressPackageStartupMessages({
  library(SeqArray)
  library(Matrix)
  library(data.table)
  library(dplyr)   # loaded after data.table so dplyr verbs take precedence
  library(stringr)
  library(purrr)
  library(tidyr)
  library(ggplot2)
  library(tibble)
})
```

## Parameters

```{r params}
# Change this path to switch iterations — each config file sets all run-specific variables.
source("/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/Talleh/Talleh_code/configs/run_pbulk_postQC.R")
```

```{r load-strn-df}
# Load QC'd strain labels early so vartrix aggregation can be restricted to these barcodes.
strn_raw <- readRDS(label_rds)
strn_df  <- if (is.null(label_element)) strn_raw else strn_raw[[label_element]]

cat("Strain label barcodes loaded:", nrow(strn_df), "\n")
```

---

## 1. VCF → GDS

```{r vcf2gds}
if (!file.exists(gds_path)) {
  seqVCF2GDS(vcf_path, gds_path, storage.option = "ZIP_RA")
}

gds <- seqOpen(gds_path, allow.duplicate = TRUE)
```

```{r dry-run-filter}
if (DRY_RUN) {
  n_total <- length(seqGetData(gds, "variant.id"))
  # Sample randomly so dry-run variants are spread across all chromosomes,
  # not just the first chromosome (VCF is position-sorted).
  set.seed(42L)
  sel <- sort(sample(n_total, min(N_DRY, n_total)))
  seqSetFilter(gds, variant.sel = sel)
  message("DRY RUN: processing ", length(sel), " of ", n_total, " variants (random sample)")
}
```

---

## 2. Extract variant matrix

```{r extract-matrix}
filt_raw <- seqGetData(gds, "annotation/filter")
filt_vec <- if (is.matrix(filt_raw)) {
  apply(filt_raw, 1, function(x) paste(colnames(filt_raw)[x], collapse = ";"))
} else {
  filt_raw
}

sample_ids <- seqGetData(gds, "sample.id") |>
  str_replace_all(c(
    "PoorQC"     = "P",
    "Low_n"      = "Lo",
    "Female"     = "F",
    "Male"       = "M",
    "Asexual"    = "A",
    "Gametocyte" = "G"
  ))

# dosage and per-sample RO/AO are (n_samples × n_variants) matrices;
# after rbind + t() each row is one variant with all sample columns appended
raw_mat <- rbind(
  seqGetData(gds, "allele"),
  seqGetData(gds, "annotation/qual"),
  seqGetData(gds, "chromosome"),
  seqGetData(gds, "position"),
  seqGetData(gds, "annotation/info/DP"),
  filt_vec,
  seqGetData(gds, "$dosage"),
  seqGetData(gds, "annotation/format/RO"),
  seqGetData(gds, "annotation/format/AO")
) |>
  t() |>
  as.data.frame() |>
  setNames(c(
    "alleles", "quality", "chrom", "pos", "depth", "vcf_filter",
    sample_ids,
    paste0(sample_ids, "_RefN"),
    paste0(sample_ids, "_AltN")
  ))

variants <- as_tibble(raw_mat) |>
  mutate(
    quality = as.numeric(quality),
    pos     = as.integer(pos),
    depth   = as.integer(depth),
    across(all_of(sample_ids), as.integer),
    across(ends_with("_RefN"), as.integer),
    across(ends_with("_AltN"), as.integer)
  )

cat("Variants read:", nrow(variants), "\n")
```

---

## 3. Vartrix per-cluster read aggregation

Replaces the aggregate VCF-level dosage calls with per-cluster counts derived by
summing raw per-cell ref/alt reads from the vartrix matrices. Haploid genotypes are
called from the cluster-aggregated allele fractions using coverage and purity
thresholds. Variants with no matching position in the vartrix VCF retain their
original GDS calls.

```{r vartrix-refine}
# Load vartrix sparse matrices (variant × cell)
alt_mat <- readMM(file.path(vartrix_dir, "alt.mtx"))
ref_mat <- readMM(file.path(vartrix_dir, "ref.mtx"))

# Column → barcode mapping (same order as barcodes.tsv)
bc_order <- fread(file.path(vartrix_dir, "barcodes.tsv"),
                  header = FALSE, col.names = "barcode")$barcode

# Cluster assignments — singlets, doublets, and unassigned
# Restricted to barcodes present in strn_df (QC-passing cells only).
cluster_assign <- fread(clusters_path) |>
  as_tibble() |>
  mutate(cluster = case_when(
    status == "singlet"    ~ assignment,
    status == "doublet"    ~ "doublet",
    TRUE                   ~ "unassigned"
  )) |>
  select(barcode, cluster) |>
  semi_join(strn_df, by = c("barcode" = label_bcode_col))

cat("Barcodes retained after strn_df filter:", nrow(cluster_assign), "\n")

cluster_ids <- sort(unique(cluster_assign$cluster))

# Sum sparse-matrix columns for each cluster's member cells
agg_counts <- function(mat, cid) {
  idx <- which(bc_order %in% cluster_assign$barcode[cluster_assign$cluster == cid])
  rowSums(mat[, idx, drop = FALSE])
}
agg_alt <- map(cluster_ids, agg_counts, mat = alt_mat)
agg_ref <- map(cluster_ids, agg_counts, mat = ref_mat)

# Haploid caller: 1 = ALT, 0 = REF, NA = below depth or ambiguous
call_gt_vt <- function(a, r) {
  total <- a + r
  frac  <- a / total
  as.integer(case_when(
    total < vt_min_reads   ~ NA_integer_,
    frac  >= vt_alt_thresh ~ 1L,
    frac  <= vt_ref_thresh ~ 0L,
    TRUE                   ~ NA_integer_
  ))
}

dosage_vt <- map_dfc(
  seq_along(cluster_ids),
  \(i) tibble(!!as.character(cluster_ids[i]) := call_gt_vt(agg_alt[[i]], agg_ref[[i]]))
)

# Variant-row alignment: vartrix rows follow order of souporcell_merged_sorted_vcf.vcf.gz
vcf_pos <- fread(
  cmd       = sprintf("zcat %s | grep -v '^#' | cut -f1,2",
                      file.path(vartrix_dir, "souporcell_merged_sorted_vcf.vcf.gz")),
  col.names = c("chrom", "pos")
) |>
  as_tibble() |>
  mutate(pos = as.integer(pos), vt_row = row_number())

row_map <- variants |>
  mutate(v_row = row_number()) |>
  inner_join(vcf_pos, by = c("chrom", "pos"))

# Add columns for extra clusters (doublet, unassigned); initialise as NA, fill from vartrix
extra_ids <- setdiff(cluster_ids, sample_ids)
for (eid in extra_ids) variants[[eid]] <- NA_integer_

# Overwrite/fill dosage columns; singlet unmatched variants retain their GDS call
for (sid in cluster_ids) {
  variants[[sid]][row_map$v_row] <- dosage_vt[[sid]][row_map$vt_row]
}

sample_ids <- c(sample_ids, extra_ids)

cat("Vartrix-refined:", nrow(row_map), "variants |",
    "Unmatched (GDS calls kept):", nrow(variants) - nrow(row_map), "\n")
```

```{r vartrix-call-qc, fig.width=10, fig.height=4}
# For each vartrix-matched position × cluster, classify the call reason.
# "Low support"  : total reads < vt_min_reads (no confident call possible)
# "Low purity"   : reads present but alt fraction is ambiguous (het-like)
# "Called"       : genotype called as REF or ALT
classify_gt_reason <- function(a, r) {
  total <- a + r
  frac  <- ifelse(total > 0, a / total, NA_real_)
  case_when(
    total <  vt_min_reads                          ~ "Low support",
    frac  >= vt_alt_thresh | frac <= vt_ref_thresh ~ "Called",
    TRUE                                           ~ "Low purity"
  )
}

vt_call_long <- map_dfc(
  seq_along(cluster_ids),
  \(i) tibble(!!cluster_ids[i] :=
    classify_gt_reason(agg_alt[[i]][row_map$vt_row],
                       agg_ref[[i]][row_map$vt_row]))
) |>
  pivot_longer(everything(), names_to = "cluster", values_to = "reason") |>
  count(cluster, reason) |>
  group_by(cluster) |>
  mutate(pct = n / sum(n)) |>
  ungroup() |>
  mutate(reason = factor(reason, levels = c("Called", "Low support", "Low purity")))

REASON_COLS <- c(Called = "#4caf50", "Low support" = "#90a4ae", "Low purity" = "#e57373")

p_vt_n <- vt_call_long |>
  ggplot(aes(x = cluster, y = n, fill = reason)) +
    geom_col(width = 0.7) +
    geom_text(aes(label = ifelse(pct > 0.02, scales::comma(n), "")),
              position = position_stack(vjust = 0.5), size = 3, colour = "white") +
    scale_fill_manual(values = REASON_COLS, name = NULL) +
    scale_y_continuous(labels = scales::comma) +
    labs(x = NULL, y = "Variant × cluster calls (n)",
         title = sprintf("Vartrix call outcome per cluster — count  (depth ≥ %d, purity thresholds %.1f / %.1f)",
                         vt_min_reads, vt_ref_thresh, vt_alt_thresh)) +
    theme_classic(base_size = 11) +
    theme(legend.position = "top")

p_vt_pct <- vt_call_long |>
  ggplot(aes(x = cluster, y = pct, fill = reason)) +
    geom_col(width = 0.7) +
    geom_text(aes(label = ifelse(pct > 0.02, scales::percent(pct, accuracy = 0.1), "")),
              position = position_stack(vjust = 0.5), size = 3, colour = "white") +
    scale_fill_manual(values = REASON_COLS, name = NULL) +
    scale_y_continuous(labels = scales::percent) +
    labs(x = NULL, y = "Proportion",
         title = "Vartrix call outcome per cluster — proportion") +
    theme_classic(base_size = 11) +
    theme(legend.position = "top")

p_vt_n
p_vt_pct
```

---

## 4. Annotate with Pf7 reference panel

```{r pf7-annotate}
pf7 <- fread(pf7_path, na.strings = ".") |>
  as_tibble() |>
  select(V1, V2, V3, V4, V11, V12, V13, V14, V15, V16, V17) |>
  rename(
    chrom      = V1,  pos    = V2,  ref    = V3,  alt    = V4,
    pf7_filter = V11, type   = V12, VQSLOD = V13,
    CDS        = V14, AC     = V15, AN     = V16, AF     = V17
  ) |>
  mutate(
    pos    = as.integer(pos),
    AF     = as.numeric(AF),
    in_pf7 = TRUE
  )

variants <- left_join(variants, pf7, by = c("chrom", "pos"),
                      relationship = "many-to-many")
```

---

## 5. Resolve multiallelic sites

Keep one record per position: rank SNP > INDEL > OVERLAP, then by descending Pf7 allele frequency.

```{r resolve-multiallelic}
variants <- variants |>
  group_by(chrom, pos) |>
  mutate(type = factor(type, levels = c("SNP", "INDEL", "OVERLAP"))) |>
  arrange(type, desc(AF), .by_group = TRUE) |>
  slice_head(n = 1) |>
  ungroup()
```

---

## 6. Hard QC filters

```{r qc-filters}
variants_qc <- variants |>
  filter(
    type       == "SNP",       # SNPs only
    chrom      != MT_CHROM,    # exclude mitochondrial
    quality    >= MIN_QUAL,    # site QUAL
    depth      >= MIN_DEPTH,   # site read depth
    !is.na(in_pf7),            # present in Pf7 panel (NA = absent after left_join)
    pf7_filter == "PASS"       # passes Pf7 QC
  )

cat("After QC filters:", nrow(variants_qc), "\n")
```

```{r qc-locus-plots, fig.width=10, fig.height=4}
# Label each variant with its first failing QC criterion
FAIL_LEVELS <- c("PASS", "Not SNP", "Mitochondrial",
                 paste0("QUAL < ", MIN_QUAL), paste0("Depth < ", MIN_DEPTH),
                 "Not in Pf7", "Pf7 non-PASS")

qc_labelled <- variants |>
  mutate(
    fail = case_when(
      type       != "SNP"        ~ "Not SNP",
      chrom      == MT_CHROM     ~ "Mitochondrial",
      quality    <  MIN_QUAL     ~ paste0("QUAL < ", MIN_QUAL),
      depth      <  MIN_DEPTH    ~ paste0("Depth < ", MIN_DEPTH),
      is.na(in_pf7)              ~ "Not in Pf7",
      pf7_filter != "PASS"       ~ "Pf7 non-PASS",
      TRUE                       ~ "PASS"
    ),
    fail = factor(fail, levels = FAIL_LEVELS)
  )

FAIL_COLS <- setNames(
  c("#4caf50", "#e57373", "#ff9800", "#9c27b0", "#2196f3", "#795548", "#607d8b"),
  FAIL_LEVELS
)

# -- QC funnel: variant count remaining after each filter ---------------------
funnel <- tibble(
  stage = factor(
    c("Extracted", "SNP only", "Nuclear",
      paste0("QUAL ≥ ", MIN_QUAL), paste0("Depth ≥ ", MIN_DEPTH),
      "In Pf7", "Pf7 PASS"),
    levels = c("Extracted", "SNP only", "Nuclear",
               paste0("QUAL ≥ ", MIN_QUAL), paste0("Depth ≥ ", MIN_DEPTH),
               "In Pf7", "Pf7 PASS")
  ),
  n = c(
    nrow(variants),
    nrow(filter(variants, type == "SNP")),
    nrow(filter(variants, type == "SNP", chrom != MT_CHROM)),
    nrow(filter(variants, type == "SNP", chrom != MT_CHROM, quality >= MIN_QUAL)),
    nrow(filter(variants, type == "SNP", chrom != MT_CHROM, quality >= MIN_QUAL,
                depth >= MIN_DEPTH)),
    nrow(filter(variants, type == "SNP", chrom != MT_CHROM, quality >= MIN_QUAL,
                depth >= MIN_DEPTH, !is.na(in_pf7))),
    nrow(variants_qc)
  )
)

p_funnel <- ggplot(funnel, aes(x = stage, y = n, group = 1)) +
  geom_line(colour = "#546e7a", linewidth = 0.9) +
  geom_point(size = 3, colour = "#546e7a") +
  geom_text(aes(label = scales::comma(n)), vjust = -0.9, size = 3) +
  scale_y_continuous(labels = scales::comma,
                     expand = expansion(mult = c(0.05, 0.2))) +
  labs(x = NULL, y = "Variants retained", title = "QC filter funnel") +
  theme_classic(base_size = 11) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))

# -- First-failing-criterion bar chart ----------------------------------------
p_fail <- qc_labelled |>
  count(fail) |>
  ggplot(aes(x = reorder(fail, n), y = n, fill = fail)) +
    geom_col(width = 0.7, show.legend = FALSE) +
    geom_text(aes(label = scales::comma(n)), hjust = -0.1, size = 3.2) +
    scale_fill_manual(values = FAIL_COLS) +
    scale_y_continuous(labels = scales::comma,
                       expand = expansion(mult = c(0, 0.18))) +
    coord_flip() +
    labs(x = NULL, y = "Variants (n)", title = "First failing QC criterion") +
    theme_classic(base_size = 11)

p_funnel
p_fail
```

```{r qc-metric-dists, fig.width=10, fig.height=4}
# -- Site QUAL distribution ---------------------------------------------------
p_qual <- variants |>
  filter(type == "SNP") |>
  mutate(pass = quality >= MIN_QUAL) |>
  ggplot(aes(x = pmin(quality, 100), fill = pass)) +
    geom_histogram(binwidth = 2, boundary = 0, colour = NA) +
    geom_vline(xintercept = MIN_QUAL, linetype = "dashed") +
    scale_fill_manual(values = c("FALSE" = "#e57373", "TRUE" = "#4caf50"),
                      labels = c("Fail", "Pass"), name = NULL) +
    labs(x = "QUAL score (capped at 100)", y = "Variants",
         title = paste0("Site quality distribution (threshold = ", MIN_QUAL, ")")) +
    theme_classic(base_size = 11) +
    theme(legend.position = "top")

# -- Site depth distribution --------------------------------------------------
p_depth <- variants |>
  filter(type == "SNP") |>
  mutate(pass = depth >= MIN_DEPTH) |>
  ggplot(aes(x = pmin(depth, 200L), fill = pass)) +
    geom_histogram(binwidth = 5, boundary = 0, colour = NA) +
    geom_vline(xintercept = MIN_DEPTH, linetype = "dashed") +
    scale_fill_manual(values = c("FALSE" = "#e57373", "TRUE" = "#4caf50"),
                      labels = c("Fail", "Pass"), name = NULL) +
    labs(x = "Read depth (capped at 200)", y = "Variants",
         title = paste0("Site depth distribution (threshold = ", MIN_DEPTH, ")")) +
    theme_classic(base_size = 11) +
    theme(legend.position = "top")

p_qual
p_depth
```

```{r vartrix-call-qc-postfilt, fig.width=10, fig.height=4}
# Restrict to vartrix-matched positions that also pass the hard QC filters.
# semi_join on chrom+pos gives exactly the overlap between row_map and variants_qc.
row_map_qc <- semi_join(row_map, variants_qc, by = c("chrom", "pos"))

cat("Post-QC vartrix-matched variants:", nrow(row_map_qc),
    "of", nrow(variants_qc), "post-QC variants\n")

vt_call_qc_long <- map_dfc(
  seq_along(cluster_ids),
  \(i) tibble(!!cluster_ids[i] :=
    classify_gt_reason(agg_alt[[i]][row_map_qc$vt_row],
                       agg_ref[[i]][row_map_qc$vt_row]))
) |>
  pivot_longer(everything(), names_to = "cluster", values_to = "reason") |>
  count(cluster, reason) |>
  group_by(cluster) |>
  mutate(pct = n / sum(n)) |>
  ungroup() |>
  mutate(reason = factor(reason, levels = c("Called", "Low support", "Low purity")))

p_vt_qc_n <- vt_call_qc_long |>
  ggplot(aes(x = cluster, y = n, fill = reason)) +
    geom_col(width = 0.7) +
    geom_text(aes(label = ifelse(pct > 0.02, scales::comma(n), "")),
              position = position_stack(vjust = 0.5), size = 3, colour = "white") +
    scale_fill_manual(values = REASON_COLS, name = NULL) +
    scale_y_continuous(labels = scales::comma) +
    labs(x = NULL, y = "Variant × cluster calls (n)",
         title = sprintf(
           "Vartrix call outcome — post-QC variants (n = %d) — count",
           nrow(row_map_qc))) +
    theme_classic(base_size = 11) +
    theme(legend.position = "top")

p_vt_qc_pct <- vt_call_qc_long |>
  ggplot(aes(x = cluster, y = pct, fill = reason)) +
    geom_col(width = 0.7) +
    geom_text(aes(label = ifelse(pct > 0.02, scales::percent(pct, accuracy = 0.1), "")),
              position = position_stack(vjust = 0.5), size = 3, colour = "white") +
    scale_fill_manual(values = REASON_COLS, name = NULL) +
    scale_y_continuous(labels = scales::percent) +
    labs(x = NULL, y = "Proportion",
         title = "Vartrix call outcome — post-QC variants — proportion") +
    theme_classic(base_size = 11) +
    theme(legend.position = "top")

p_vt_qc_n
p_vt_qc_pct
```

---

## 7. Informative SNPs

Retain only sites polymorphic across strain clusters (≥ 2 distinct non-NA genotype calls).

```{r informative-snps}
dosage_mat     <- as.matrix(select(variants_qc, all_of(sample_ids)))
is_informative <- apply(dosage_mat, 1, function(x) length(unique(na.omit(x))) > 1L)
variants_clean <- variants_qc[is_informative, ]

cat("Informative SNPs retained:", nrow(variants_clean), "\n")
```

```{r cluster-labels}
# Count barcode support for every (cluster, label) pair.
# strn_df was loaded early (load-strn-df chunk) and already filtered cluster_assign.
label_freq <- fread(clusters_path) |>
  as_tibble() |>
  filter(status == "singlet") |>
  select(barcode, assignment) |>
  inner_join(
    select(strn_df, all_of(c(label_bcode_col, label_col))),
    by = c("barcode" = label_bcode_col)
  ) |>
  mutate(label = as.character(.data[[label_col]])) |>
  count(assignment, label, name = "n") |>
  arrange(desc(n))

# Greedy 1-to-1 assignment: process pairs in descending support order;
# accept a pair only when neither its cluster nor its label has been taken.
singlet_labels <- local({
  taken_c <- character(); taken_l <- character(); rows <- list()
  for (i in seq_len(nrow(label_freq))) {
    r <- label_freq[i, ]
    if (!r$assignment %in% taken_c && !r$label %in% taken_l) {
      rows[[length(rows) + 1]] <- r
      taken_c <- c(taken_c, r$assignment)
      taken_l <- c(taken_l, r$label)
    }
  }
  matched <- bind_rows(rows) |> arrange(assignment)
  setNames(matched$label, matched$assignment)
})

extra_labels   <- c("doublet" = "Doublet", "unassigned" = "Unassigned")
cluster_labels <- c(singlet_labels,
                    extra_labels[names(extra_labels) %in% cluster_ids])

rename_cols    <- setNames(names(cluster_labels), cluster_labels)
variants_qc    <- rename(variants_qc,    !!!rename_cols)
variants_clean <- rename(variants_clean, !!!rename_cols)
sample_ids     <- unname(cluster_labels)

cat("Cluster label mapping (k", K, "):\n", sep = "")
print(data.frame(cluster = names(cluster_labels), label = unname(cluster_labels)),
      row.names = FALSE)
```

```{r qc-cluster-plots, fig.width=10, fig.height=4}
CALL_COLS <- c(REF = "#4caf50", ALT = "#2196f3", "NA" = "#bdbdbd")

# -- Dosage composition per cluster (REF / ALT / NA) -------------------------
dosage_long <- variants_clean |>
  select(all_of(sample_ids)) |>
  pivot_longer(everything(), names_to = "cluster", values_to = "dosage") |>
  mutate(
    call    = case_when(is.na(dosage) ~ "NA",
                        dosage == 0L  ~ "REF",
                        dosage == 1L  ~ "ALT"),
    call    = factor(call, levels = c("REF", "ALT", "NA")),
    cluster = paste0("cluster_", cluster)
  )

p_dosage <- dosage_long |>
  count(cluster, call) |>
  group_by(cluster) |>
  mutate(pct = n / sum(n)) |>
  ungroup() |>
  ggplot(aes(x = cluster, y = pct, fill = call)) +
    geom_col(width = 0.7) +
    geom_text(aes(label = ifelse(pct > 0.02, scales::percent(pct, accuracy = 0.1), "")),
              position = position_stack(vjust = 0.5), size = 3, colour = "white") +
    scale_fill_manual(values = CALL_COLS, name = "Call") +
    scale_y_continuous(labels = scales::percent) +
    labs(x = NULL, y = "Proportion of informative SNPs",
         title = "Dosage composition per cluster (informative SNPs)") +
    theme_classic(base_size = 11)

# -- Per-SNP completeness: number of clusters with a call per site ------------
n_called_vec <- rowSums(!is.na(as.matrix(select(variants_clean, all_of(sample_ids)))))

p_completeness <- tibble(n_called = n_called_vec) |>
  count(n_called) |>
  mutate(n_called = factor(n_called)) |>
  ggplot(aes(x = n_called, y = n)) +
    geom_col(fill = "#546e7a", width = 0.6) +
    geom_text(aes(label = scales::comma(n)), vjust = -0.4, size = 3.2) +
    scale_y_continuous(labels = scales::comma,
                       expand = expansion(mult = c(0, 0.15))) +
    labs(x = "Clusters with a call", y = "Informative SNPs",
         title = "Per-SNP completeness across clusters") +
    theme_classic(base_size = 11)

p_dosage
p_completeness
```

```{r qc-na-by-stage, fig.width=7, fig.height=4}
# -- Per-cluster NA rate at two pipeline stages -------------------------------
na_by_stage <- bind_rows(
  variants_qc    |>
    summarise(across(all_of(sample_ids), \(x) mean(is.na(x)))) |>
    pivot_longer(everything(), names_to = "cluster", values_to = "na_rate") |>
    mutate(stage = "After QC filters"),
  variants_clean |>
    summarise(across(all_of(sample_ids), \(x) mean(is.na(x)))) |>
    pivot_longer(everything(), names_to = "cluster", values_to = "na_rate") |>
    mutate(stage = "Informative SNPs")
) |>
  mutate(cluster = paste0("cluster_", cluster),
         stage   = factor(stage, levels = c("After QC filters", "Informative SNPs")))

ggplot(na_by_stage, aes(x = cluster, y = na_rate, fill = stage)) +
  geom_col(position = "dodge", width = 0.6) +
  scale_fill_manual(values = c("After QC filters" = "#90a4ae",
                                "Informative SNPs" = "#37474f"), name = NULL) +
  scale_y_continuous(labels = scales::percent,
                     expand = expansion(mult = c(0, 0.1))) +
  labs(x = NULL, y = "Fraction NA",
       title = "Per-cluster missing-call rate by pipeline stage") +
  theme_classic(base_size = 11) +
  theme(legend.position = "top")
```

---

## 8. Pairwise Identity by State

For haploid *Plasmodium*, IBS at a biallelic SNP is 1 when two clusters share the same
allele and 0 when they differ. Haploid dosage (0 = REF, 1 = ALT) is recoded to homozygous
diploid encoding (0 / 2) before calling `snpgdsIBS()`, which is the standard approach for
haploid organisms.

```{r ibs-compute}
library(SNPRelate)

dosage_haploid <- as.matrix(select(variants_clean, all_of(sample_ids)))
dosage_haploid[dosage_haploid == 1L] <- 2L   # haploid ALT → homozygous encoding

alleles_split <- strsplit(variants_clean$alleles, ",")
gds_ibs_path  <- sub("\\.gds$", "_snprel.gds", gds_path)
if (file.exists(gds_ibs_path)) file.remove(gds_ibs_path)

snpgdsCreateGeno(
  gds.fn         = gds_ibs_path,
  genmat         = t(dosage_haploid),          # samples × SNPs
  sample.id      = sample_ids,
  snp.id         = seq_len(nrow(variants_clean)),
  snp.chromosome = variants_clean$chrom,
  snp.position   = variants_clean$pos,
  snp.allele     = paste(sapply(alleles_split, `[[`, 1),
                         sapply(alleles_split, function(x) x[2]), sep = "/"),
  snpfirstdim    = FALSE
)

gds_ibs <- snpgdsOpen(gds_ibs_path)
ibs     <- snpgdsIBS(gds_ibs, autosome.only = FALSE, verbose = FALSE)
snpgdsClose(gds_ibs)
invisible(file.remove(gds_ibs_path))
```

```{r ibs-heatmap, fig.width=6, fig.height=5}
library(pheatmap)

ibs_mat <- ibs$ibs
rownames(ibs_mat) <- colnames(ibs_mat) <- paste0("cluster_", sample_ids)

pheatmap::pheatmap(
  ibs_mat,
  color           = colorRampPalette(c("white", "#2166ac"))(100),
  breaks          = seq(0, 1, length.out = 101),
  display_numbers = TRUE,
  number_format   = "%.3f",
  fontsize_number = 10,
  main            = "Pairwise IBS — souporcell clusters (haploid Pf)",
  cluster_rows    = TRUE,
  cluster_cols    = TRUE
)
```

---

## 9. Chromosome painting

SNPs are binned into windows across the 14 nuclear chromosomes for every pairwise
cluster comparison. Each window is classified as **Same** (all callable SNPs share
the same allele), **Different** (all differ), **Both** (mixed), or **No SNP**
(shown as the white chromosome body).

Two functions drive this section and can be called interactively:

- `make_chrom_windows(variants, sids, wndw, pairs)` — compute window data
- `chrom_paintr(cw, comps_filter, ncol)` — draw the chromosome painting

```{r chrom-paint-fns}
# Pf3D7_v3 nuclear chromosome sizes in bp (PlasmoDB)
pf_chr_size <- tibble(
  chr  = 1:14,
  size = c(643292, 947102, 1067971, 1200490, 1343557, 1418242,
           1445207, 1472805, 1541735, 1687656, 2038340, 2271494,
           2925236, 3291936)
)

# Compute per-window pairwise similarity across the genome.
#   variants : clean variants tibble (output of section 7)
#   sids     : sample ID vector (sample_ids)
#   wndw     : window width in bp (default 10 kb)
#   pairs    : list of 2-element character vectors; NULL = all C(n,2) combinations
#              e.g. list(c("0","1"), c("0","2"))
make_chrom_windows <- function(variants, sids, wndw = 10000L, pairs = NULL) {
  if (is.null(pairs)) pairs <- combn(sids, 2, simplify = FALSE)

  pw_base <- variants |>
    select(chrom, pos, all_of(sids)) |>
    mutate(chr = as.integer(str_extract(chrom, "[0-9]+(?=_v3)"))) |>
    filter(!is.na(chr))   # drop MIT / API chromosomes

  # Per-SNP same (TRUE) / different (FALSE) / uncallable (NA) for each pair
  pair_sim <- map_dfc(pairs, \(p) {
    x <- pw_base[[p[1]]]; y <- pw_base[[p[2]]]
    tibble(!!paste0(p[1], "_vs_", p[2]) :=
      case_when(!is.na(x) & !is.na(y) & x == y ~ TRUE,
                !is.na(x) & !is.na(y) & x != y ~ FALSE))
  })

  bind_cols(select(pw_base, chr, pos), pair_sim) |>
    mutate(bin_n = ceiling(pos / wndw),
           start = (bin_n - 1L) * wndw / 1e6,   # bp → Mb
           end   = bin_n * wndw / 1e6) |>
    pivot_longer(names(pair_sim), names_to = "comps", values_to = "sim") |>
    group_by(comps, chr, bin_n, start, end) |>
    summarise(n_same = sum( sim, na.rm = TRUE),
              n_diff = sum(!sim, na.rm = TRUE),
              .groups = "drop") |>
    mutate(
      # Continuous IBS ratio: 0 = all SNPs different (black), 1 = all same (yellow).
      # NA when no callable SNPs in a window (white background shows through).
      sim_ratio = if_else(n_same + n_diff > 0,
                          n_same / (n_same + n_diff),
                          NA_real_),
      comps = paste0("Cluster ", str_replace(comps, "_vs_", " vs ")),
      chr   = factor(chr, levels = 1:14)
    )
}

# Draw chromosome painting from a make_chrom_windows() result.
#   cw           : output of make_chrom_windows()
#   comps_filter : character vector of comparison labels to include; NULL = all
#                  e.g. c("Cluster 0 vs 1", "Cluster 0 vs 2")
#   ncol         : number of facet columns
chrom_paintr <- function(cw, comps_filter = NULL, ncol = 2) {
  if (!is.null(comps_filter)) cw <- filter(cw, comps %in% comps_filter)

  # Full chromosome background: all 14 chrs × all comparisons in cw.
  # Built from pf_chr_size so every chromosome is drawn even when it has no
  # SNPs in the window data (e.g. in a dry run covering only chr 1).
  chr_bg <- expand_grid(
    chr   = factor(1:14, levels = 1:14),
    comps = unique(cw$comps)
  ) |>
  left_join(mutate(pf_chr_size, size_mb = size / 1e6,
                                chr     = factor(chr, levels = 1:14)),
            by = "chr")

  wndw_kb <- round(median((cw$end - cw$start) * 1e3))

  ggplot() +
    geom_segment(data = chr_bg,
                 aes(x = -0.002, xend = size_mb + 0.002, y = chr, yend = chr),
                 linewidth = 3.2, lineend = "round", colour = "black") +
    geom_segment(data = chr_bg,
                 aes(x = 0, xend = size_mb, y = chr, yend = chr),
                 linewidth = 2.5, lineend = "round", colour = "white") +
    geom_segment(data = filter(cw, !is.na(sim_ratio)),
                 aes(x = start, xend = end, y = chr, yend = chr, colour = sim_ratio),
                 linewidth = 2.5) +
    # black (0 = all different) → white (0.5) → yellow (1 = all same)
    scale_colour_gradient2(low = "black", mid = "white", high = "#DACB7D",
                           midpoint = 0.5, limits = c(0, 1),
                           name = "IBS ratio\n(same / callable)",
                           labels = scales::label_percent()) +
    scale_y_discrete(limits = rev) +
    facet_wrap(~comps, ncol = ncol) +
    labs(x     = "Position (Mb)", y = "Chromosome",
         title = sprintf("Chromosome painting — pairwise IBS in %d kb windows", wndw_kb)) +
    theme_classic() +
    theme(strip.text      = element_text(size = 9, face = "bold"),
          axis.text       = element_text(size = 8),
          legend.position = "right")
}
```

```{r chrom-paint-run, fig.width=12, fig.height=10}
# All 10 pairwise comparisons with 10 kb windows.
# Change wndw for coarser/finer resolution, or pass pairs= to subset comparisons:
#   make_chrom_windows(variants_clean, sample_ids, wndw = 50000L)
#   make_chrom_windows(variants_clean, sample_ids, pairs = list(c("0","1"), c("0","2")))
# To subset after computing:
#   chrom_paintr(chrom_win, comps_filter = c("Cluster 0 vs 1", "Cluster 0 vs 2"))

paint_sids <- if (is.null(chrom_paint_sids)) sample_ids else chrom_paint_sids
paint_sids <- c("SC1", "SC2", "SC3", "SC4", "SC5")
chrom_win  <- make_chrom_windows(variants_clean, sids = paint_sids, wndw = chrom_paint_wndw)
cat("Comparisons:", n_distinct(chrom_win$comps),
    "| Windows with data:", nrow(chrom_win), "\n")

chrom_paintr(chrom_win, ncol =3)
```

---

## 10. Summary and save

```{r summary}
tibble(
  step  = c("Read", "After QC", "Informative"),
  n     = c(nrow(variants), nrow(variants_qc), nrow(variants_clean))
)
```

```{r preview}
variants_clean |>
  select(chrom, pos, alleles, type, AF,
         all_of(sample_ids[seq_len(min(3, length(sample_ids)))])) |>
  head(10)
```

```{r save}
seqClose(gds)

if (DRY_RUN) {
  message("DRY RUN complete — output not written")
} else {
  dir.create(dirname(out_path), recursive = TRUE, showWarnings = FALSE)
  saveRDS(variants_clean, out_path)
  cat("Saved:", out_path, "\n")
}
```
