---
title: "MSC49 Strain Clusters vs pxPlaFalc52 — Genotype Comparison"
date: '`r Sys.Date()`'
author: 'Jesse Rop'
output:
  html_document:
    toc: true
    theme: united
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/Talleh")
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

## Overview

Compares the genotypes of the five MSC49 souporcell strain clusters (SC1–SC5) against
a single bulk *P. falciparum* isolate (pxPlaFalc52) independently genotyped by bcftools
mpileup. The analysis uses:

1. **Pairwise IBS** — Identity by State across all overlapping, allele-concordant SNPs
2. **Chromosome painting** — windowed IBS ratio across all 14 nuclear chromosomes

The MSC49 cluster genotypes are taken from `clean_variants_k5d_postQClabels.rds`,
which is the output of `vcf_variant_qc.Rmd` after vartrix refinement and Pf7 QC.

Mismatch distribution and mutation bias analysis continues in `mismatch_bias_analysis.Rmd`.

To knit from the command line:
```{bash knit-cmd, eval=FALSE}
module load HGI/softpack/groups/team222/Pf_scRNAseq/34
Rscript -e "rmarkdown::render('Talleh_code/strain_vs_reference_genotype_comparison.Rmd')"
```

---

## Libraries

```{r libs}
.libPaths(c("/opt/view/rlib/R/library", .libPaths()))

suppressPackageStartupMessages({
  library(SeqArray)
  library(SNPRelate)
  library(dplyr)
  library(stringr)
  library(purrr)
  library(tidyr)
  library(ggplot2)
  library(tibble)
})
```

## Parameters

```{r params}
source("Talleh_code/project_common_fns.R")
source("Talleh_code/pf_common_vars.R")

CLEAN_VARIANTS_RDS <- "data/processed/Pf/MSC49/clean_variants_k5_postQClabels.rds"
REF_VCF            <- "/lustre/scratch126/tol/teams/lawniczak/projects/malaria_single_cell/mali_field_runs/2022/genomes/grit/vcfs/lrHMW_um_fasta_onto_Pf3D7_bcftools/pxPlaFalc52_filtered.vcf.gz"
OUT_DIR            <- "data/processed/Pf/MSC49/ref_comparison"
REF_ID             <- "pxPlaFalc52"
WINDOW_SIZE        <- 100000L    # bp window for chromosome painting
MT_CHROM           <- "Pf3D7_MIT_v3"
API_CHROM          <- "Pf3D7_API_v3"
GFF_FILE           <- "/lustre/scratch126/tol/teams/lawniczak/users/jr35/Pf3D7_genomes_gtfs_indexs/PlasmoDB-68_Pfalciparum3D7.gff"

dir.create(OUT_DIR, showWarnings = FALSE, recursive = TRUE)
```

---

## 1. Load Clean Cluster Variants

```{r load-clusters}
variants_clean <- readRDS(CLEAN_VARIANTS_RDS)

# Identify dosage columns: exclude known metadata and read-count columns
meta_cols    <- c("chrom", "pos", "alleles", "type", "quality", "depth",
                  "vcf_filter", "pf7_filter", "VQSLOD", "CDS",
                  "AC", "AN", "AF", "in_pf7")
readcnt_cols <- grep("_RefN$|_AltN$", names(variants_clean), value = TRUE)
non_singlet  <- intersect(names(variants_clean), c("Doublet", "Unassigned", "ref", "alt"))
cluster_ids  <- setdiff(names(variants_clean),
                        c(meta_cols, readcnt_cols, non_singlet))

cat("Cluster columns detected:", paste(cluster_ids, collapse = ", "), "\n")
cat("Informative SNPs in clean variants:", nrow(variants_clean), "\n")
```

---

## 2. Read Reference VCF (pxPlaFalc52)

Convert the bcftools-called VCF to GDS format and extract haploid genotype dosage
(0 = REF, 1 = ALT, NA = missing). Only nuclear chromosomes are retained.

```{r read-ref-vcf}
ref_gds_path <- file.path(OUT_DIR, "pxPlaFalc52_ref.gds")

if (exists("ref_gds")) tryCatch(seqClose(ref_gds), error = function(e) NULL)
seqVCF2GDS(REF_VCF, ref_gds_path, storage.option = "ZIP_RA")
ref_gds <- seqOpen(ref_gds_path)

# Restrict to nuclear chromosomes present in the cluster variants
nuclear_chroms <- setdiff(unique(variants_clean$chrom), c(MT_CHROM, API_CHROM))
seqSetFilter(ref_gds,
             variant.sel = seqGetData(ref_gds, "chromosome") %in% nuclear_chroms)

cat("Reference VCF variants (nuclear):", length(seqGetData(ref_gds, "variant.id")), "\n")

# Extract variant coordinates and alleles
ref_chrom   <- seqGetData(ref_gds, "chromosome")
ref_pos     <- seqGetData(ref_gds, "position")
ref_alleles <- seqGetData(ref_gds, "allele")   # "REF,ALT[,ALT2]" format

# SeqArray $dosage returns 0 = REF, 1 = ALT (standard VCF convention).
# The MSC49 cluster data uses the OPPOSITE encoding (0 = ALT, 1 = REF) because
# vartrix alt.mtx and ref.mtx output files are swapped: alt.mtx contains
# REF-supporting reads and ref.mtx contains ALT-supporting reads. The haploid
# caller in vcf_variant_qc.Rmd therefore assigns 1 to the REF allele and 0 to
# the ALT allele, despite the comment saying the reverse.
# Invert the reference dosage (1 - dosage) so both datasets share the same
# encoding before comparison.
ref_dosage <- seqGetData(ref_gds, "$dosage")   # shape: (1, n_variants)

seqClose(ref_gds)

ref_variants <- tibble(
  chrom   = ref_chrom,
  pos     = ref_pos,
  ref_vcf = str_extract(ref_alleles, "^[^,]+"),
  alt_vcf = str_extract(ref_alleles, "(?<=,)[^,]+")   # first ALT only
) |>
  mutate(!!REF_ID := 1L - as.integer(ref_dosage[1, ]))

cat("Parsed reference variants:", nrow(ref_variants), "\n")
```

---

## 3. Find Overlapping Variants

Join on chromosome + position. Both VCFs were called against the same Pf3D7
reference genome so the REF allele must be identical at any shared position.
Sites where the ALT allele differs (different variant called at the same position,
e.g. one VCF called a SNP and the other an INDEL) are dropped.

```{r find-overlap}
cluster_variants <- variants_clean |>
  mutate(
    ref_sc = str_extract(alleles, "^[^,]+"),
    alt_sc = str_extract(alleles, "(?<=,)[^,]+")
  )

joined <- inner_join(
  cluster_variants,
  ref_variants,
  by = c("chrom", "pos")
)

overlap <- joined |>
  filter(ref_sc == ref_vcf, alt_sc == alt_vcf)

cat("Inner join (chrom + pos):", nrow(joined), "sites\n")
cat("Allele-concordant overlap:", nrow(overlap), "sites\n")
cat("Dropped (different ALT allele):", nrow(joined) - nrow(overlap), "sites\n")
```

---

## 4. Combined Dosage Matrix

```{r combined-matrix}
all_ids <- c(cluster_ids, REF_ID)

overlapping_variants <- overlap |>
  select(chrom, pos, alleles, CDS, ref_sc, alt_sc, all_of(all_ids))

dosage_mat <- as.matrix(select(overlapping_variants, all_of(all_ids)))

cat("Combined dosage matrix:", paste(dim(dosage_mat), collapse = " variants × "), "samples\n")
cat("\nMissing-call rates per sample:\n")
round(colMeans(is.na(dosage_mat)), 3) |> print()
```

---

## 5. Pairwise IBS

Both datasets now use the vartrix convention: 0 = ALT allele, 1 = REF allele.
The value 1 is recoded to 2 for `snpgdsIBS()` (homozygous diploid representation),
matching the pattern in `vcf_variant_qc.Rmd`.

```{r ibs-compute}
dosage_diploid <- dosage_mat
storage.mode(dosage_diploid) <- "double"
dosage_diploid[dosage_diploid == 1] <- 2     # haploid ALT → homozygous encoding

alleles_split <- strsplit(overlapping_variants$alleles, ",")
gds_ibs_path  <- file.path(OUT_DIR, "ibs_tmp.gds")
if (file.exists(gds_ibs_path)) file.remove(gds_ibs_path)

snpgdsCreateGeno(
  gds.fn         = gds_ibs_path,
  genmat         = t(dosage_diploid),              # samples × SNPs
  sample.id      = all_ids,
  snp.id         = seq_len(nrow(overlapping_variants)),
  snp.chromosome = overlapping_variants$chrom,
  snp.position   = overlapping_variants$pos,
  snp.allele     = paste(sapply(alleles_split, `[[`, 1),
                         sapply(alleles_split, function(x) x[2]), sep = "/"),
  snpfirstdim    = FALSE
)

gds_ibs <- snpgdsOpen(gds_ibs_path)
ibs     <- snpgdsIBS(gds_ibs, autosome.only = FALSE, verbose = FALSE)
snpgdsClose(gds_ibs)
invisible(file.remove(gds_ibs_path))

saveRDS(ibs, file.path(OUT_DIR, "ibs_vs_pxPlaFalc52.rds"))
```

```{r ibs-summary}
ibs_mat <- ibs$ibs
rownames(ibs_mat) <- colnames(ibs_mat) <- all_ids

ibs_df <- tibble(
  Cluster   = cluster_ids,
  IBS_score = ibs_mat[cluster_ids, REF_ID]
) |>
  arrange(desc(IBS_score))

cat("Pairwise IBS — MSC49 clusters vs", REF_ID, "\n")
knitr::kable(ibs_df, digits = 4,
             col.names = c("Cluster", "IBS score"),
             caption   = sprintf("Pairwise IBS: MSC49 strain clusters vs %s", REF_ID))
```

```{r ibs-barplot, fig.width = 5, fig.height = 4}
ggplot(ibs_df, aes(x = reorder(Cluster, -IBS_score), y = IBS_score, fill = IBS_score)) +
  geom_col(width = 0.6, colour = "grey30") +
  geom_text(aes(label = sprintf("%.3f", IBS_score)),
            vjust = -0.4, size = 3.5) +
  scale_fill_gradient(low = "white", high = "#2166ac",
                      limits = c(0, 1), guide = "none") +
  scale_y_continuous(limits = c(0, 1),
                     expand = expansion(mult = c(0, 0.1))) +
  labs(
    x     = "MSC49 strain cluster",
    y     = "Pairwise IBS",
    title = sprintf("Genotype similarity — MSC49 clusters vs %s", REF_ID)
  ) +
  theme_classic(base_size = 12)

ggsave(file.path(OUT_DIR, "cluster_ref_ibs_barplot.pdf"), width = 5, height = 4)
```

---

## 6. Chromosome Painting

Each cluster is painted against pxPlaFalc52 using `make_chrom_windows()` and
`chrom_paintr()` from `project_common_fns.R`. Windows coloured yellow indicate
high allele sharing (IBS ≈ 1); black indicates all SNPs differ; white indicates
no callable SNPs in that window.

```{r chrom-paint, fig.width = 10, fig.height = 4.5}
# pairs_vs_ref <- map(cluster_ids, ~c(.x, REF_ID))
pairs_vs_ref <- map(cluster_ids, ~c(.x, "HMW DNA"))
all_ids_hmw <- all_ids
all_ids_hmw[6] <- "HMW DNA"

cw <- make_chrom_windows(
  variants = overlapping_variants %>% rename("HMW DNA" = pxPlaFalc52),
  sids     = all_ids_hmw,
  wndw     = WINDOW_SIZE,
  pairs    = pairs_vs_ref
)

p_paint <- chrom_paintr(cw, ncol = 3)
print(p_paint)

ggsave(file.path(OUT_DIR, "cluster_ref_chrom_painting.pdf"),
       plot = p_paint, width = 10, height = 8)
```

---

## 7. Summary

```{r summary}
best_match <- ibs_df |> slice_max(IBS_score, n = 1)

cat(sprintf(
  "Best match to %s: %s (IBS = %.4f)\n",
  REF_ID, best_match$Cluster, best_match$IBS_score
))
cat(sprintf("SNPs used for comparison: %d\n", nrow(overlapping_variants)))
cat(sprintf("Chromosome painting window size: %d kb\n", WINDOW_SIZE / 1000L))
cat(sprintf("Outputs written to: %s/\n", OUT_DIR))
```

---

## 8. Mismatch Distribution — Best Cluster vs Reference

Identifies sites where **`r best_match$Cluster`** disagrees with **`r REF_ID`** and tests
whether those mismatches are randomly distributed across chromosomes, genome windows,
CDS / intergenic regions, and (optionally) annotated gene families.

### 8a. Identify mismatches

```{r mismatch-identify}
best_cluster <- best_match$Cluster

mismatch_df <- overlapping_variants |>
  mutate(
    cluster_gt  = .data[[best_cluster]],
    ref_gt      = .data[[REF_ID]],
    is_callable = !is.na(cluster_gt) & !is.na(ref_gt),
    is_mismatch = is_callable & (cluster_gt != ref_gt)
  )

cat(sprintf("Callable sites (both genotyped): %d\n",   sum(mismatch_df$is_callable)))
cat(sprintf("Mismatch sites:                  %d (%.2f%% of callable)\n",
            sum(mismatch_df$is_mismatch),
            100 * sum(mismatch_df$is_mismatch) / sum(mismatch_df$is_callable)))
```

### 8b. Per-chromosome mismatch rates

```{r mismatch-per-chrom}
chrom_stats <- mismatch_df |>
  filter(is_callable) |>
  group_by(chrom) |>
  summarise(
    n_callable    = n(),
    n_mismatch    = sum(is_mismatch),
    mismatch_rate = n_mismatch / n_callable,
    .groups = "drop"
  ) |>
  arrange(chrom)

knitr::kable(chrom_stats, digits = 4,
             caption = sprintf("Per-chromosome mismatch rates: %s vs %s",
                               best_cluster, REF_ID))
```

```{r save-mismatch-inputs}
# Save objects needed by mismatch_bias_analysis.Rmd
saveRDS(mismatch_df,  file.path(OUT_DIR, "mismatch_df.rds"))
saveRDS(chrom_stats,  file.path(OUT_DIR, "chrom_stats.rds"))
saveRDS(
  list(best_cluster = best_cluster,
       REF_ID       = REF_ID,
       OUT_DIR      = OUT_DIR,
       WINDOW_SIZE  = WINDOW_SIZE,
       GFF_FILE     = GFF_FILE),
  file.path(OUT_DIR, "mismatch_params.rds")
)
cat("Inputs for mismatch_bias_analysis.Rmd saved to", OUT_DIR, "\n")
```

---

## Session info

```{r session-info}
sessionInfo()
```
