---
title: "MSC QC and preprocessing"
date: '2022-07-11'
author: 'Jesse Rop'
output:
  html_document:
    toc: true
    theme: united
    code_folding: hide
    df_print: paged
---

```{r}
suppressPackageStartupMessages({
  library(Seurat)
  library(patchwork)
  library(ggrepel)
  library(ggpubr)
  library(dplyr) 
  library(stringr)
  library(purrr)
  library(readr)
  library(tibble)
  library(ggplot2)
  library(tidyr)
  library(ggvenn)
  library(conflicted)
  library(ggsankey)
  
  conflict_prefer("select", "dplyr")
  conflict_prefer("filter", "dplyr")
  conflict_prefer("rename", "dplyr")
  conflict_prefer("map", "purrr")
  conflict_prefer("count", "dplyr")
  conflicts_prefer(base::saveRDS)
  conflicts_prefer(SingleCellExperiment::colData)
  conflicts_prefer(base::as.data.frame)
  conflicts_prefer(base::intersect)
  conflicts_prefer(base::setdiff)
  conflicts_prefer(dplyr::desc)
  conflicts_prefer(dplyr::count)
  conflicts_prefer(dplyr::vars)
})
```



# MSC Preprocessing

## setup, read input files and exploratory visualizations and general functions

```{r, setup}
knitr::opts_knit$set(root.dir = '/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/Talleh')

```

### Variable declarations

### Common functions

#### Source commong functions, varriables and tables
```{r}
##Boxplot labeling function
source("~/Talleh_code/pf_common_vars.R")
source("~/Talleh_code/project_common_fns.R")

```

Visualize strain assignment 
Read sankey plots input

```{r, fig.height=6, fig.width=24, eval=T}
## Save for visualization of sankey plots for the manuscript
strn_harmnzd_tbls <- readRDS("data/processed/Pf/m2_all/strain_calling/sstrn_harmnzd_tbls.RDS")
don_k_names_aln <- readRDS("data/processed/Pf/m2_all/strain_calling/don_k_names_aln.RDS")
opt_clusters_nms_aln <- readRDS("data/processed/Pf/m2_all/strain_calling/opt_clusters_nms_aln.RDS")
don_hmonizd_k_names_aln <- readRDS("data/processed/Pf/m2_all/strain_calling/don_hmonizd_k_names_aln.RDS")
```

### Sankey plots: within-aligner strain label transitions

```{r, fig.width=6, fig.height=4, eval=T}
snk_plts <- list(strn_harmnzd_tbls, don_k_names_aln, opt_clusters_nms_aln, don_hmonizd_k_names_aln) %>%
  pmap(~{  
    
    snk_pt_fn(..1, strns_2_plt = c(..2[1], ..4[2:length(..4)]), labls = paste0("K=", ..3[1:5]), bp_face = c("plain"))
    
  })

```


```{r, eval=T}
snk_plts[str_detect(names(snk_plts), "hsat")] %>% imap(., ~.x + labs(title = .y))
```

Visualizations and tabulations for Tallehs dataset for assesment

```{r, eval=T}
## Save asexual cells for Talleh
msc49_sct_asx_domStrainSC2_ta19 <- readRDS("data/raw/Pf/MSC49/alt_qcd_objects/msc49_sct_asx_domStrainSC3_ta19.RDS")

msc49_sct_ta19 <- readRDS("data/raw/Pf/MSC49/alt_qcd_objects/msc49_sct_ta19.RDS")

```


```{r, eval=T}
msc49_sct_ta19@assays$RNA
```


```{r, eval=T}
msc49_sct_ta19@meta.data %>% count(Stage)
```

```{r, eval=T}
msc49_sct_ta19@meta.data %>% count(Stage, Strain)
```

###### Filters

```{r, eval=F}
filt_flow_diff_t <- readRDS("data/processed/Pf/m2_all/filt_flow_diff_t.RDS") 
filt_flow_dbt_diff_t <- readRDS("data/processed/Pf/m2_all/filt_flow_dbt_diff_t.RDS") 
```

```{r, fig.width=4.5, fig.height=3, eval=F}

filt_flow %>% 
  select(MSC49, filt_metric) %>%  
  ggplot(aes(x = filt_metric, y = MSC49)) + geom_col()

```

```{r, eval=T}
cb_mdata_orig <- read_csv("data/raw/Pf/MSC49/cbender/cb_mdata.csv")
```

```{r, eval=T}
cb_mdata_orig_set_m49ta <-  cb_mdata_orig %>% filter(barcode %in% colnames(msc49_sct_ta19))
```


```{r, eval=T}
summary(cb_mdata_orig_set_m49ta$background_fraction)
summary(cb_mdata_orig_set_m49ta$cell_probability)

```


```{r, message=F, warning=F}
##Read qcd dataset qcd by me
## Combined JC + non-JC
all_msc_qcd_anotd_m49C <- readRDS("data/raw/Pf/MSC49/alt_qcd_objects/cln_anotd3_jc_combind_norm.RDS")

# ## JC only
# all_msc_qcd_anotd_m49 <- readRDS("data/raw/Pf/MSC49/alt_qcd_objects/cln_anotd3_norm.RDS")

## New JC only - use singleR stage
msc49_cb_no_dbt_sct <- readRDS("data/processed/Pf/MSC49/seu_obj/all_cb_no_dbt_sct.RDS")

```


```{r, eval=T}
## Assign the stages used in respective datasets into one common variable and assign gametocyte developing to rings
# ifelse(msc49_sct_ta19@Stage == "", "", "")
all_msc_qcd_anotd_m49C$Stage <- all_msc_qcd_anotd_m49C$stage_fld 
msc49_cb_no_dbt_sct$Stage <- msc49_cb_no_dbt_sct$sR_fine_stg

## Fill a missing stage variable
msc49_sct_ta19$stage_fld <- msc49_sct_ta19$sR_fine_stg
```

## cells in Talleh's dataset that are in Jesse dataset
```{r, eval=T}
table(colnames(msc49_sct_ta19) %in% colnames(all_msc_qcd_anotd_m49C))
# table(colnames(msc49_sct_ta19) %in% colnames(all_msc_qcd_anotd_m49))
```


```{r, eval=T}
## Cells in Tallehs dataset
msc49_sct_ta19$qc_dset <- ifelse(colnames(msc49_sct_ta19) %in% colnames(all_msc_qcd_anotd_m49C) , "talleh_n_jesse", "talleh_only" )
all_msc_qcd_anotd_m49C$qc_dset <- ifelse(colnames(all_msc_qcd_anotd_m49C) %in% colnames(msc49_sct_ta19) , "talleh_n_jesse", "jesse_only" )
# all_msc_qcd_anotd_m49$qc_dset <- ifelse(colnames(all_msc_qcd_anotd_m49) %in% colnames(msc49_sct_ta19) , "talleh_n_jesse", "jesse_only" )
msc49_cb_no_dbt_sct$qc_dset <- ifelse(colnames(msc49_cb_no_dbt_sct) %in% colnames(msc49_sct_ta19) , "talleh_n_jesse", "jesse_only" )

```

```{r, eval=T}
## Put objects into list for harmonised manipulation
# msc49_ta_new_n_jr35 <- list(all_msc_qcd_anotd_m49C, msc49_cb_no_dbt_sct)
msc49_ta_n_jr35 <- list(all_msc_qcd_anotd_m49C, msc49_sct_ta19)

```

```{r, eval=T}
## Recode strain
### Get counts to see how they match
# strain_map_counts <- map(msc49_ta_new_n_jr35, ~.x@meta.data %>% select(Strain_DN) %>% rownames_to_column('bcode') %>%
strain_map_counts <- map(msc49_ta_n_jr35, ~.x@meta.data %>% select(Strain_DN) %>% rownames_to_column('bcode') %>%
  left_join(., msc49_cb_no_dbt_sct@meta.data %>% select(Strain_DN) %>% rownames_to_column('bcode'), by = 'bcode') %>% 
    mutate(across(contains("Strain"), as.character)) %>%
  count(Strain_DN.x, Strain_DN.y) %>%
  filter(!is.na(Strain_DN.y) & Strain_DN.x != "Negative" & Strain_DN.y != "Negative"))

strain_map_counts

### Get mapping key
strain_key <- map(strain_map_counts, ~.x %>%
  group_by(Strain_DN.x) %>% 
  slice_max(n, n = 1, with_ties = FALSE) %>% 
  ungroup() %>% 
  select(Strain_DN.x, Strain_DN.y) %>% 
  deframe())

strain_key

```

```{r, eval=T}
## Add recoded strain
msc49_ta_n_jr35 <- map2(msc49_ta_n_jr35, strain_key, ~{
  .x@meta.data %>% 
    mutate(Strain = case_when(Strain_DN == "Negative" ~ Strain_DN,
                              T ~ .y[Strain_DN])) %>%
    select(Strain) %>%
  AddMetaData(.x, .)
  })

```

```{r, eval=T}
# msc49_dsets <- list(all_msc_qcd_anotd_m49C, all_msc_qcd_anotd_m49, msc49_sct_ta19, msc49_cb_no_dbt_sct) %>% set_names(c("m49_JRcombnd", "m49_JR", "m49_TA", "m49_TA_new"))
# msc49_dsets <- c(msc49_ta_new_n_jr35, list(msc49_sct_ta19)) %>% set_names(c("m49_JRcombnd", "m49_TA_new", "m49_TA"))
msc49_dsets <- c(msc49_ta_n_jr35, list(msc49_cb_no_dbt_sct)) %>% set_names(c("m49_JRcombnd", "m49_TA", "m49_TA_new"))

```

```{r, eval=T}
# Save Qcd recoded strain for chromosome paint plotting
# msc49_dsets_ta <- msc49_dsets["m49_TA"]
# names(msc49_dsets_ta) <- "MSC49_hsat"

saveRDS(msc49_dsets[["m49_TA"]]@meta.data %>% rownames_to_column("bcode") %>% select(bcode, Strain), "data/processed/Pf/pbulk_gtypes_postQC_resc_hspf/qcd_strain_barcodes.RDS")

```

```{r, warning=FALSE, message=FALSE}
imap(msc49_dsets, ~DimPlot(.x, dims = c(1,2), group.by = "stage_fld", pt.size = 0.1)+ scale_color_manual(values = sp_fld_colors2) + labs(title = .y))
imap(msc49_dsets, ~DimPlot(.x, dims = c(1,2), group.by = "Stage", pt.size = 0.1)+ scale_color_manual(values = sp_fld_colors2) + labs(title = .y))
```

```{r, warning=FALSE, message=FALSE, fig.width= 5, fig.height=4}
DimPlot(msc49_dsets[["m49_TA"]], dims = c(1,2), group.by = "stage_fld", pt.size = 0.1)+ scale_color_manual(values = sp_fld_colors2) + 
  theme(legend.position = "none",
        text = element_text(size = 18))
```


```{r, warning=FALSE, message=FALSE, fig.width= 9, fig.height=4}
DimPlot(msc49_dsets[["m49_TA"]], dims = c(1,2), group.by = "stage_fld", pt.size = 0.1)+ scale_color_manual(values = sp_fld_colors2) + 
  theme(legend.position = "bottom",
        text = element_text(size = 18))
```


```{r, warning=FALSE, message=FALSE, fig.width= 7, fig.height=4}
msc49_dsets_ploting <- msc49_dsets[["m49_TA"]]
# msc49_dsets_ploting$Stage <- ifelse(msc49_dsets_ploting$stage_fld == "Gametocyte (developing)", "Developing \ngametocyte", msc49_dsets_ploting$stage_fld)
msc49_dsets_ploting$Stage <- ifelse(msc49_dsets_ploting$stage_fld == "Gametocyte (developing)", "Early ring", msc49_dsets_ploting$stage_fld)

msc49_dsets_ploting$Stage <- factor(msc49_dsets_ploting$Stage, levels = c("Early ring", "Late ring", "Early trophozoite", "Developing \ngametocyte", "Female", "Female (LE)", "Male"))
# msc49_dsets_ploting$Stage <- factor(msc49_dsets_ploting$Stage, levels = c("Early ring", "Late ring", "Early trophozoite", "Female", "Female (LE)", "Male"))

DimPlot(msc49_dsets_ploting, dims = c(1,2), group.by = "Stage", pt.size = 0.4)+ scale_color_manual(values = sp_fld_colors2) + 
  theme(legend.position = "right",
        text = element_text(size = 20),
        plot.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank()) +
  guides(color = guide_legend(title.position = 'top', ncol = 1, name = 'Stage', override.aes = list(size = 4)))
```

```{r, eval=T}
map(msc49_dsets, ~.x@meta.data %>% count(mthds_union))
```

```{r, eval=T}
map(msc49_dsets, ~.x@meta.data %>% count(Strain))

```

```{r, eval=T}
map(msc49_dsets, ~.x@meta.data %>% count(Stage))

```

```{r, eval=T}
msc49_dsets <- map(msc49_dsets, ~subset(.x, Stage %in% c("discrepantFM", "weak_score", "Schizont"), invert = T))

```

```{r, warning=FALSE, message=FALSE, fig.width=6, fig.height=3.5}

msc49_dsets <- map(msc49_dsets, ~.x@meta.data %>% 
  mutate(stage_ag = ifelse(str_detect(Stage,"ale"), "Sexual", "Asexual"),
         stage_afm = case_when(str_detect(Stage,"Female") ~ "Female",
                               str_detect(Stage,"Male") ~ "Male",
                               T ~ "Asexual"),
         # Strain = Strain_DN
         ) %>% 
  select(stage_ag, stage_afm, Strain) %>%
  AddMetaData(.x, .)
)

```

```{r, eval=T}
msc49_dsets[["m49_TA"]]@meta.data %>% count( Strain, stage_afm)
msc49_dsets[["m49_TA"]]@meta.data %>% count(stage_afm, Strain)
msc49_dsets[["m49_TA"]]@meta.data %>% count(Stage, Strain)
msc49_dsets[["m49_TA"]]@meta.data %>% count(Strain, Stage)
```

```{r, warning=FALSE, message=FALSE, fig.width= 5, fig.height=4}
DimPlot(msc49_dsets[["m49_TA"]], dims = c(1,2), cells.highlight = colnames(msc49_dsets[["m49_TA"]][,msc49_dsets[["m49_TA"]]$stage_ag == "Asexual"]) , pt.size = 0.1)+ 
  theme(legend.position = "none",
        text = element_text(size = 18))
```

## asexual strains in Talleh's dataset that are in Jesse dataset and vice versa
```{r, eval=T}
map(msc49_dsets, ~.x@meta.data %>% filter(stage_afm == "Asexual") %>% count(Strain, qc_dset)) 
```

```{r, eval=T}
map(msc49_dsets, ~.x@meta.data %>% filter(stage_afm == "Asexual") %>% count(qc_dset)) 
```

```{r, eval=T}
map(msc49_dsets, ~.x@meta.data %>% count(qc_dset)) 
```

```{r, warning=FALSE, message=FALSE}
map(msc49_dsets, ~.x@meta.data %>% count(stage_ag, Strain) %>% group_by(stage_ag) %>% mutate(prop_n = round(n/sum(n), 2)))

```


```{r, message=FALSE, warning=FALSE}

stage_df_snky <- map2(msc49_dsets, names(msc49_dsets), function(obj, nm) obj@meta.data %>% rename_with(~paste0(., "_", nm), everything()) %>% rownames_to_column("bcode") ) %>%
  purrr::reduce(., full_join, by = "bcode") %>%  
  mutate(across(contains("Stage_"), \(x) replace_na(as.character(x), "Missing")),
         across(contains("Stage_"), ~str_replace_all(., c("discrepantFM" = "dFM", "weak_score" = "WS", "Early " = "E", "Late " = "L", "ring" = "R", "trophozoite" = "T", "Schizont" = "S", "Female" = "F", "Male" = "M", " \\(LE\\)" = "LE", "Gametocyte \\(developing\\)" = "Gd", "Missing" = "Mis"))),
         across(contains("Strain_"), \(x) replace_na(as.character(x), "Mis"))) 



```

```{r, warning=FALSE, message=FALSE}
stage_df_snky %>% count(Stage_m49_JRcombnd, Stage_m49_TA, Stage_m49_TA_new) %>% arrange(desc(n))

```


```{r, message=FALSE, warning=FALSE}
# Generate sankey plots comparing SingleR fine stage to scmap field stage
snk_pt_fn_stg(stage_df_snky, strns_2_plt = paste0("Stage_", names(msc49_dsets)), labls = names(msc49_dsets), bp_face = rep("bold", length(msc49_dsets)), ttl = "obj", strain_colours = sp_col_new_abrv, remove_misssing_vals = F)

```

```{r, message=FALSE, warning=FALSE}
# Generate sankey plots comparing SingleR fine stage to scmap field stage
stage_df_snky %>% filter(Stage_m49_TA != "Mis" | Stage_m49_TA_new != "Mis") %>%
snk_pt_fn_stg(., strns_2_plt = c("Stage_m49_TA", "Stage_m49_TA_new"), labls = c("m49_TA", "m49_TA_new"), bp_face = c("bold", "bold"), ttl = "obj", strain_colours = sp_col_new_abrv, remove_misssing_vals = F)

```


```{r, message=FALSE, warning=FALSE}
# Generate sankey plots comparing strains between datasets
snk_pt_fn_stg(stage_df_snky, strns_2_plt = paste0("Strain_", names(msc49_dsets)), labls = names(msc49_dsets), bp_face = rep("bold", length(msc49_dsets)), ttl = "obj",  remove_misssing_vals = F)

```



```{r, message=FALSE, warning=FALSE}
# Generate sankey plots comparing SingleR fine stage to scmap field stage
stage_df_snky %>% filter(Strain_m49_TA != "Mis" | Strain_m49_TA_new != "Mis") %>%
snk_pt_fn_stg(., strns_2_plt = c("Strain_m49_TA", "Strain_m49_TA_new"), labls = c("m49_TA", "m49_TA_new"), bp_face = c("bold", "bold"), ttl = "obj", remove_misssing_vals = F)

```



```{r, message=FALSE, warning=FALSE, fig.width=3, fig.height=2.5}
stage_df_snky %>% filter(Strain_m49_TA == "SC2" & stage_ag_m49_TA == "Asexual") %>%
snk_pt_fn_stg(., strns_2_plt = paste0("Strain_", names(msc49_dsets)), labls =  str_remove_all(names(msc49_dsets), "m49_|combnd"), bp_face = rep("bold", length(msc49_dsets)), ttl = "obj", remove_misssing_vals = F)

```

```{r, message=FALSE, warning=FALSE, fig.width=3, fig.height=2.5}
msc49_dsets_2 <- msc49_dsets[c("m49_JRcombnd", "m49_TA")]
```

```{r, message=FALSE, warning=FALSE, fig.width=2.5, fig.height=3}
stage_df_snky %>% 
  # filter(Strain_m49_TA == "SC3" & stage_ag_m49_TA == "Asexual") %>%
  filter((Strain_m49_TA == "SC2" & stage_ag_m49_TA == "Asexual")  | (Strain_m49_JRcombnd == "SC2" & stage_ag_m49_JRcombnd == "Asexual")) %>%
  snk_pt_fn_stg(., strns_2_plt = paste0("Strain_", names(msc49_dsets_2)), labls =  str_remove_all(names(msc49_dsets_2), "m49_|combnd"), bp_face = rep("bold", length(msc49_dsets_2)), ttl = "obj", remove_misssing_vals = F)

```

```{r, message=FALSE, warning=FALSE, fig.width=2.5, fig.height=3}
msc49_dsets_asex_2_bcodes <- map(msc49_dsets_2, ~colnames(.x[,.x$stage_ag =="Asexual" & .x$Strain == "SC2"]))
names(msc49_dsets_asex_2_bcodes) <- str_remove_all(names(msc49_dsets_asex_2_bcodes), "m49_|combnd")
```

```{r, message=FALSE, warning=FALSE, fig.width=5.5, fig.height=5.4}
ggvenn(data = msc49_dsets_asex_2_bcodes, text_size = 6)
```

```{r, message=FALSE, warning=FALSE, fig.width=3, fig.height=2.5}
stage_df_snky %>% 
  # filter(Strain_m49_TA == "SC3" & stage_ag_m49_TA == "Asexual") %>%
  filter((Strain_m49_TA == "SC2" & stage_ag_m49_TA == "Asexual")  | (Strain_m49_JRcombnd == "SC2" & stage_ag_m49_JRcombnd == "Asexual")) %>%
  snk_pt_fn_stg(., strns_2_plt = paste0("Stage_", names(msc49_dsets)), labls =  str_remove_all(names(msc49_dsets), "m49_|combnd"), bp_face = rep("bold", length(msc49_dsets)), ttl = "obj", strain_colours = sp_col_new_abrv, remove_misssing_vals = F)

```

```{r, message=FALSE, warning=FALSE, fig.width=2.5, fig.height=3}
stage_df_snky %>% 
  # filter(Strain_m49_TA == "SC3" & stage_ag_m49_TA == "Asexual") %>%
  filter((Strain_m49_TA == "SC2" & stage_ag_m49_TA == "Asexual")  | (Strain_m49_JRcombnd == "SC2" & stage_ag_m49_JRcombnd == "Asexual")) %>%
  snk_pt_fn_stg(., strns_2_plt = paste0("Stage_", names(msc49_dsets_2)), labls =  str_remove_all(names(msc49_dsets_2), "m49_|combnd"), bp_face = rep("bold", length(msc49_dsets_2)), ttl = "obj", strain_colours = sp_col_new_abrv, remove_misssing_vals = F)

```

```{r, eval=T}
map(msc49_dsets, ~.x@meta.data %>% count(Stage))

```

```{r, warning=FALSE, message=FALSE}
ta_asexuals <- msc49_dsets[["m49_TA"]]@meta.data %>% filter(stage_ag == "Asexual" & Strain == "SC2") %>% rownames()
jr_asexuals <- msc49_dsets[["m49_JRcombnd"]]@meta.data %>% filter(stage_ag == "Asexual" & Strain == "SC2") %>% rownames()

ta_jr_comp <- list(overlap = intersect(ta_asexuals, jr_asexuals),
     only_TA = setdiff(ta_asexuals, jr_asexuals),
     only_JR = setdiff(jr_asexuals, ta_asexuals))

map(ta_jr_comp, length)
```
```{r, warning=FALSE, message=FALSE}
ta_asexuals <- msc49_dsets[["m49_TA"]]@meta.data %>% filter(stage_ag == "Asexual" & Strain == "SC2") %>% rownames()
ta_replicate_asexuals <- msc49_dsets[["m49_TA_new"]]@meta.data %>% filter(stage_ag == "Asexual" & Strain == "SC2") %>% rownames()

ta_taRep_comp <- list(overlap = intersect(ta_asexuals, ta_replicate_asexuals),
     only_TA = setdiff(ta_asexuals, ta_replicate_asexuals),
     only_JR = setdiff(ta_replicate_asexuals, ta_asexuals))

map(ta_taRep_comp, length)
```


```{r, warning=FALSE, message=FALSE}
msc49_dsets_mrg_mdata <- map2(msc49_dsets, str_remove(names(msc49_dsets), "m49"), function(obj, nm) obj@meta.data %>% rename_with(~paste0(., nm), everything()) %>% rownames_to_column("bcode")) %>% purrr::reduce(., full_join, by = "bcode") 

```


```{r, warning=FALSE, message=FALSE}
## Are the asexual SC3 strains used by Talleh similar to the strain assignments in Jesse's dataset
msc49_dsets_mrg_mdata %>% filter(stage_ag_TA == "Asexual" & Strain_TA == "SC2") %>% count(Strain_TA, Strain_JRcombnd, stage_ag_JRcombnd)

```

```{r, warning=FALSE, message=FALSE}
## Are the asexual SC3 strains used by Talleh called in a similar manner, in terms of cellranger assignments, in both datasets 
msc49_dsets_mrg_mdata %>% filter(stage_ag_TA == "Asexual" & Strain_TA == "SC2") %>% count(mthds_union_TA, mthds_union_JRcombnd)

```


```{r, warning=FALSE, message=FALSE}
## Remove the negatives which were retained in Jesse's dataset
msc49_dsets_no_neg <- map(msc49_dsets, ~subset(.x, Strain != "Negative"))
```

```{r, warning=FALSE, message=FALSE}
map(msc49_dsets_no_neg, ~DimPlot(.x, dims = c(1,3), group.by = "Stage", pt.size = 0.4)+ scale_color_manual(values = sp_fld_colors2))
map(msc49_dsets_no_neg, ~DimPlot(.x, dims = c(1,2), group.by = "Stage", pt.size = 0.4)+ scale_color_manual(values = sp_fld_colors2))
```


```{r, warning=FALSE, message=FALSE}
map(msc49_dsets_no_neg, ~DimPlot(.x, dims = c(1,2), group.by = "qc_dset", pt.size = 0.4))
map(msc49_dsets_no_neg, ~DimPlot(.x, dims = c(1,2), split.by = "qc_dset", pt.size = 0.4))

```

```{r, warning=FALSE, message=FALSE}
map(msc49_dsets_no_neg, ~DimPlot(.x, dims = c(1,2), cells.highlight = colnames(msc49_sct_ta19[, msc49_sct_ta19$Stage== "Gametocyte (developing)"])) )

```

```{r, warning=FALSE, message=FALSE}
map(msc49_dsets_no_neg, ~FeaturePlot(.x, dims = c(1,2), features = "PF3D7-1222600", order = T, pt.size = 0.4))
```

```{r, warning=FALSE, message=FALSE}
map(msc49_dsets_no_neg, ~VlnPlot(.x, features = "PF3D7-1222600", group.by = "Stage", pt.size = 0.01, slot = "data", layer = "RNA"))
map(msc49_dsets_no_neg, ~VlnPlot(.x, features = "PF3D7-1102500", group.by = "Stage", pt.size = 0.01, slot = "data", layer = "RNA"))

```


```{r, warning=FALSE, message=FALSE, fig.width=15, fig.height=4}

don_clst_strn_barplot <- function(dset, count_var, facets, label_var, y_lab = 'Counts' ) {
  ggplot(dset, aes(x = Strain, y = !!rlang::sym(count_var), fill = Strain)) +
  geom_col(position = position_dodge2()) + #width=0.4, position = position_dodge(width=0.5) 
  geom_text(aes(label=!!rlang::sym(label_var)), hjust=-0.05, vjust= 0.5, position = position_dodge(width = .9), size = 5, show.legend = F, fontface ='bold', angle = 90) +
  scale_fill_manual(values = strain_cols_n) +
  ggh4x::facet_nested_wrap(vars(!!!syms(facets)),  nrow = 1, scales = "free_x")+
  scale_y_continuous(expand = expansion(mult = c(0.01, 0.55))) +
  scale_x_discrete(expand = expansion(add = c(1, 1))) +
  theme_classic() +
  labs( y = y_lab)+
  theme(
    axis.text.y  = element_text(size = 14),
    axis.title.y  = element_text(size = 17, face = "bold"),
    axis.text.x  = element_text(size = 14, face = "bold", angle = 90),
    # axis.title.x = element_blank(),
    axis.title.x = element_text(size = 17, face = "bold"),
    legend.text = element_text(size = 17),
    legend.title = element_text(size = 17, face = "bold"),
    legend.position="bottom",
    strip.text =element_text(size=16,face="bold", colour = 'black'),
    panel.border=element_rect(colour="black",fill='transparent')
  ) + 
  guides(fill = guide_legend(title.position = 'left', nrow = 1, name = 'Strain'))
}
```


```{r, warning=FALSE, message=FALSE, fig.width=4, fig.height=3}

map(msc49_dsets_no_neg, ~.x@meta.data %>% 
  count(stage_ag, Strain) %>% 
  group_by(stage_ag) %>% 
  mutate(prop_n = round(n/sum(n), 2)) %>% 
  don_clst_strn_barplot(., count_var = "prop_n", label_var= "n", facets =c("stage_ag"), y_lab = "Proportion") +
    theme(legend.position = "none",
          axis.text.x = element_text(vjust = 0.5))
)
```


```{r, warning=FALSE, message=FALSE, fig.width=4, fig.height=3}

map(msc49_dsets_no_neg, ~.x@meta.data %>% 
  count(stage_afm, Strain) %>% 
  group_by(stage_afm) %>% 
  mutate(prop_n = round(n/sum(n), 2)) %>% 
  don_clst_strn_barplot(., count_var = "prop_n", label_var= "n", facets =c("stage_afm"), y_lab = "Proportion") +
    theme(legend.position = "none",
          axis.text.x = element_text(vjust = 0.5))
)
```


```{r, warning=FALSE, message=FALSE, fig.width=6, fig.height=3.5}

map(msc49_dsets_no_neg, ~.x@meta.data %>% 
  count(stage_ag, Strain) %>% 
  group_by(stage_ag) %>% 
  mutate(prop_n = round(n/sum(n), 2)) %>% 
  don_clst_strn_barplot(., count_var = "n", label_var= "prop_n", facets =c("stage_ag"))
)
```

```{r, warning=FALSE, message=FALSE, fig.width=6, fig.height=3.5}

map(msc49_dsets_no_neg, ~.x@meta.data %>% 
  count(stage_afm, Strain) %>% 
  group_by(stage_afm) %>% 
  mutate(prop_n = round(n/sum(n), 2)) %>% 
  don_clst_strn_barplot(., count_var = "n", label_var= "prop_n", facets =c("stage_afm"))
)
```


```{r, warning=FALSE, message=FALSE, fig.width=6, fig.height=3.5}

map(msc49_dsets_no_neg, ~.x@meta.data %>% 
  count(stage_afm, Strain) %>% 
  group_by(stage_afm) %>% 
  mutate(prop_n = round(n/sum(n), 2)) %>% 
  don_clst_strn_barplot(., count_var = "n", label_var= "prop_n", facets =c("stage_afm"))
)
```

Visualize chromosome paint plots
```{r, warning=FALSE, message=FALSE, fig.width=6, fig.height=3.5}
clean_variants_k5 <- readRDS("data/processed/Pf/MSC49/clean_variants_k5.rds")
```

```{r, warning=FALSE, message=FALSE, fig.width=6, fig.height=3.5}
clean_variants_k5
```


```{r chrom-paint-run, fig.width=12, fig.height=10}
# Chromosome paints of pairwise comparisons with windows.
# Change wndw for coarser/finer resolution, or pass pairs= to subset comparisons:

chrom_win <- make_chrom_windows(clean_variants_k5, sids = c("SC1", "SC2", "SC3", "SC4", "SC5") , wndw = 50000L)
cat("Comparisons:", n_distinct(chrom_win$comps),
    "| Windows with data:", nrow(chrom_win), "\n")

chrom_paintr(chrom_win)
```




