---
title: "msc_strain_decon_new"
date: '2023-08-10'
author: 'Jesse Rop'
output:
  html_document:
    toc: true
    theme: united
    code_folding: hide
    df_print: paged
---


```{r, message=FALSE}
suppressPackageStartupMessages({
  library(Seurat)
  library(dplyr) 
  library(stringr)
  library(purrr)
  library(readr)
  library(tibble)
  library(ggplot2)
  library(tidyr)
  library(cowplot)
  library(pheatmap)
  library(conflicted)
  library(ComplexHeatmap)
  library(ggvenn)
  library(ggsankey)
  
  conflict_prefer("select", "dplyr")
  conflict_prefer("count", "dplyr")
  conflict_prefer("filter", "dplyr")
  conflict_prefer("rename", "dplyr")
  conflict_prefer("saveRDS", "base")
  conflicts_prefer(base::as.data.frame)
})

```


# MSC parasites strain deconvolution and VCF visualization 

## Variable declarations


```{r setup}
knitr::opts_knit$set(root.dir = '/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/Talleh/')
```

### Common functions

#### Source commong functions, varriables and tables
```{r}
##Boxplot labeling function
source("~/Talleh_code/pf_common_vars.R")
source("~/Talleh_code/project_common_fns.R")

```

## Specify variables and sample name details

```{r}
# Select sample names
sample_name_nms <- c("MSC49", "MSC68")

```

```{r}
# Variables
sample_set = 'Mali2'
algn_nm = 'minmap'
species = 'Pf'

soup_dir = "soupc_resc_hspf"
sv_dir = "pbulk_gtypes_preQC_resc_hspf"


strain_levels_univrs <- c(paste0("SC", 1:15), "Doublet", "Negative")

## Alignment methods
algn_nms = c('minmap', 'hsat')

## Combine donor ids with alignment methods since each donor is ran once per alignment method
sample_name_nms_al <- rep(sample_name_nms, length(algn_nms))
algn_nms_al <- rep(algn_nms, each = length(sample_name_nms))

sm_aln_nms_al <- paste(rep(sample_name_nms, length(algn_nms)),
                       rep(algn_nms, each = length(sample_name_nms)),
                       sep = '_')

## number of optimum clusters to investigate per sample
opt_clusters_nms <- list(
  "MSC49" = c(5:9),
  "MSC68" = c(2:6)
) %>% set_names(sample_name_nms)


opt_narrow_clusters_nms <-  list(
  "MSC49" = c(5:7), #c(4:7) Should be c(4:7) for better visualisation of k5
  "MSC68" = c(3:4)
) %>% set_names(sample_name_nms)

opt_narrow_clusters_nms_aln <- rep(opt_narrow_clusters_nms, length(algn_nms)) %>% set_names(sm_aln_nms_al)

optimum_k = c("MSC49" =5, 
              "MSC68" =3) %>% set_names(sample_name_nms)

optimum_k_aln = rep(optimum_k, length(algn_nms)) %>% set_names(sm_aln_nms_al)


## number of optimum clusters (K) iterations for each donor
opt_clusters_len <- map(opt_clusters_nms, length)

## number of optimum clusters to investigate per sample factoring the number of alignments
opt_clusters_nms_aln <- rep(opt_clusters_nms, length(algn_nms)) %>% set_names(sm_aln_nms_al)
opt_clusters_len_aln <- map(opt_clusters_nms_aln, length)


## Donors factoring the number of alignments
sample_name <- map2(sample_name_nms, opt_clusters_len[sample_name_nms], ~
                      rep(.x, each = .y))

sample_name <- unlist(flatten(sample_name))

opt_clusters <- unlist(flatten(opt_clusters_nms[sample_name_nms]))

## sorted donor_opt_cluster names
sorted_names <- paste0(sample_name, "_", opt_clusters)

## sorted factoring alignment
sorted_names_al <- paste0(sample_name,
                          "_",
                          rep(algn_nms, each = length(sample_name)),
                          "_",
                          opt_clusters)

iter = opt_clusters_len

downsample_cells = "yes"
```



## Log likelihood Elbow plots visualisations for all MSCs

```{r, message=FALSE, warning=FALSE, eval=T}
## Read in log likelihoods for estimation of number of clusters from elbow plots
clusters_log_likelihoods_lst = map2(sample_name_nms_al, rep(algn_nms, each= length(sample_name_nms)), ~read_table(paste0('data/processed/',species,'/', .x,'/', soup_dir, '/',.y,'/clusters_log_likelihoods.txt'), col_names = c('No_of_clusters', 'td1','td2','td3', 'td4','log_likelihood')) %>% select(-(starts_with("td"))) %>% mutate(across(No_of_clusters, ~str_remove_all(., ".*out_dir_k|/clusters.err:best")))) %>% set_names(paste(sample_name_nms_al, rep(algn_nms, each= length(sample_name_nms)), sep ='_'))
```


```{r, eval=T} 
map(clusters_log_likelihoods_lst, ~.x %>%
  mutate(across(c('No_of_clusters'), ~factor(as.numeric(.)))) %>%
  ggplot(., aes(x = No_of_clusters, y = log_likelihood)) +
  geom_point() +
  geom_line(aes(group = 1)) +
  # facet_grid(rows = .~id)+
  theme(axis.text=element_text(size=10),
        axis.title=element_text(size=16,face="bold"),
        legend.text = element_text(size=15),
        title = element_text(size=16,face="bold"),
        strip.text.x = element_text(size = 18, face = "bold")))
```


```{r, fig.width= 20, fig.height=4, eval=T} 
clusters_log_likelihoods_lst %>% 
  bind_rows(., .id = "sample_algnr") %>% 
  # separate_wider_regex(cols = "sample_algnr", c(sample = "MSC[0-9]*|MSC[0-9]*_[A-Z]*|MSC[0-9]*[A-Za-z]*|MSC[0-9]*[A-Za-z]*_[A-Z]*", "_",algnr = "[a-z]*$")) %>%
  separate_wider_regex(cols = "sample_algnr", patterns = c(sample = ".*", "_", algnr  = "minmap$|hsat$")) %>%
  mutate(across(c('sample'), ~factor(., levels = sample_name_nms))) %>%
  mutate(across(c('No_of_clusters'), ~factor(as.numeric(.)))) %>%
  ggplot(., aes(x = No_of_clusters, y = log_likelihood)) +
    geom_point() +
    geom_line(aes(group = 1)) + 
  facet_grid(algnr ~ sample, scales = "free_y")
```

Read all the souporcell files for all the iterations. May be excessive but since we have the memory and cpu and we may need to use the files downstream then why not.

combi
```{r, eval=T}
soupc_all_al <- list.files(paste0("data/processed/", species), pattern = "clusters.tsv", full.names = T, recursive = T, include.dirs = T)%>% set_names(nm = (str_remove_all(.,paste0('data/processed/', species, '/|', soup_dir,'/|k|/clusters.tsv'))) %>% str_replace_all(., '/', '_'))  %>% .[sorted_names_al]%>%
  map_df(read_delim, delim = "\t", col_types = 'cccddd', .id = "sample_k") 


```

Select only information from clusters.tsv for the best k as determined from the log likelihood plot for each MSC sample

```{r, eval=T}

soupo_kselect_al <- soupc_all_al %>%
  dplyr::select(sample_k, barcode, status, assignment) %>% 
  dplyr::mutate("Strain" = case_when(status == "doublet" ~ "Doublet",
                                     status == "unassigned" ~ "Negative",
                                     assignment == "0" ~ "SC1",
                                     assignment == "1" ~ "SC2",
                                     assignment == "2" ~ "SC3",
                                     assignment == "3" ~ "SC4",
                                     assignment == "4" ~ "SC5",
                                     assignment == "5" ~ "SC6",
                                     assignment == "6" ~ "SC7",
                                     assignment == "7" ~ "SC8",
                                     assignment == "8" ~ "SC9",
                                     assignment == "9" ~ "SC10",
                                     assignment == "10" ~ "SC11",
                                     assignment == "11" ~ "SC12",
                                     assignment == "12" ~ "SC13",
                                     assignment == "13" ~ "SC14",
                                     assignment == "14" ~ "SC15",
                                     assignment == "15" ~ "SC16")
  ) 
```




```{r, eval=T}
# Compare HISAT and MINIMAP
soupo_kselect_al_cor <- soupo_kselect_al %>%
  separate_wider_regex(cols = "sample_k", patterns = c(sample = "M.*", "_", aligner  = "minmap|hsat", "_", k = "\\d+")) %>%
  unite(col = "sample_k", c('sample','k'), remove = FALSE)  %>%
  select("sample_k","aligner","barcode","Strain") %>% 
  dplyr::group_split(sample_k) %>%
  purrr::set_names(purrr::map_chr(., ~.x$sample_k[1])) %>%
  .[sorted_names]

hsat_mnmp_comp = map(soupo_kselect_al_cor, ~pivot_wider(.x, names_from = aligner, values_from = Strain) %>%
  dplyr::select(barcode, all_of(algn_nms)))
  
```

pheatmap
```{r , fig.width = 10, fig.height = 5, warning=FALSE}
## Heatmap function for plotting pheatmap for number of cells assigned labels from Clara's and Roser's datasets
phmap_lab_transfr <- function(dset_lab, col_labs, row_labs, row_labs_c, col_labs_c, strn_order, ttl){
  anno_row = dset_lab %>% dplyr::count({{ col_labs }}, {{ row_labs }}) %>%
    pivot_wider(values_from = n, names_from = {{ col_labs }}) %>%
    column_to_rownames(names(.)[1]) %>%
    mutate({{ row_labs }} := row_labs_c) %>%
    dplyr::select({{ row_labs }})

  anno_col = dset_lab %>% dplyr::count({{ col_labs }}, {{ row_labs }}) %>%
    pivot_wider(values_from = n, names_from = {{ row_labs }}) %>%
    column_to_rownames(names(.)[1]) %>%
    mutate({{ col_labs }} := col_labs_c) %>%
    dplyr::select({{ col_labs }})

  hm_mtx <- dset_lab %>% dplyr::count({{ col_labs }}, {{ row_labs }}) %>%
    arrange({{ col_labs }}, {{ row_labs }}) %>%
    mutate(across(c({{ col_labs }}, {{ row_labs }}), ~droplevels(factor(., levels = strain_levels_univrs)))) %>%
    spread({{ col_labs }},n,fill = 0) %>% ## Using spread instead of pivot_longer since it orders the columns based on factors
    # pivot_wider(values_from = n, names_from = {{ col_labs }}) %>%
    # replace(is.na(.), 0) %>%
    column_to_rownames(names(.)[1]) #%>%
    # .[,strn_order] %>%

    # print(head(hm_mtx))

    hm_mtx %>%
    ComplexHeatmap::pheatmap(display_numbers = T,treeheight_row=0, treeheight_col=0, annotation_row = anno_row, annotation_col = anno_col, legend = F, annotation_legend = T, cluster_rows =F, cluster_cols = F, fontsize = 20, annotation_names_row=T, annotation_names_col = T,
                       # color = 'white',
                       number_format = "%.0f", main = ttl#,
                       # annotation_colors = list('minmap' = "green", 'hsat' = "lightblue")
                       )

}

```




```{r , fig.width = 10, fig.height = 5, warning=FALSE}

imap(hsat_mnmp_comp, possibly(~phmap_lab_transfr(.x, col_labs= `minmap`, row_labs=`hsat`, row_labs_c ='hsat', col_labs_c= 'minmap', ttl = .y)), 'k=1 so no groups to compare which is necessary for the heatmap')

```

```{r , fig.width = 10, fig.height = 5, warning=FALSE}
imap(hsat_mnmp_comp, possibly(~phmap_lab_transfr(.x, col_labs= `minmap`, row_labs=`hsat`, row_labs_c ='minmap', col_labs_c= 'hsat', ttl = .y)), 'k=1 so no groups to compare which is necessary for the heatmap')

```

```{r , fig.width = 10, fig.height = 5, warning=FALSE}
imap(hsat_mnmp_comp, possibly(~phmap_lab_transfr(.x, col_labs= `minmap`, row_labs=`hsat`, row_labs_c ='minmap', col_labs_c= 'hsat', ttl = .y)), 'k=1 so no groups to compare which is necessary for the heatmap')

```



```{r , fig.width = 10, fig.height = 5, eval = F}
## Heatmap function for plotting pheatmap for number of cells assigned labels
imap(hsat_lr_ref_comp, possibly(~phmap_lab_transfr(.x, col_labs= `lr_ref`, row_labs=`hsat`, row_labs_c ='lr_ref', col_labs_c= 'hsat', ttl = .y)), 'k=1 so no groups to compare which is necessary for the heatmap')

```


```{r , fig.width = 10, fig.height = 5, eval = F}
## Heatmap function for plotting pheatmap for number of cells assigned labels from Clara's and Roser's datasets
imap(mnmp_lr_ref_comp, possibly(~phmap_lab_transfr(.x, col_labs= `lr_ref`, row_labs=`minmap`, row_labs_c ='lr_ref', col_labs_c= 'minmap', ttl = .y)), 'k=1 so no groups to compare which is necessary for the heatmap')

```



Split the information for each sample into different labels and put into a list named appropriately
```{r}
soupo_kselect_list <- soupo_kselect_al %>%
# filter(str_detect(sample_k, algn_nm)) %>%
# soupo_kselect_list <- soupo_kselect %>%
  # mutate(across(sample_k, ~str_remove(.,paste0("_", algn_nm)))) %>%
  dplyr::group_split(sample_k) %>%
  purrr::set_names(purrr::map_chr(., ~.x$sample_k[1])) %>%
  map(. %>% column_to_rownames(., "barcode") %>% dplyr::select(Strain))

soupo_kselect_list <- soupo_kselect_list[as.character(sorted_names_al)]


##Get strain levels for samples with more than one strain and mutate Strain column to reflect this
# strain_levels <- map(soupo_kselect_list[2:8], ~.x %>% filter(str_detect(Strain, 'Strain')) %>% pull(Strain) %>% unique() %>% sort() %>% c(.,"Doublet","Negative"))
strain_levels <- map(soupo_kselect_list, ~.x %>% dplyr::filter(str_detect(Strain, '^SC')) %>% pull(Strain) %>% unique() %>% sort() %>% c(.,"Doublet","Negative"))

soupo_kselect_list <- map2(soupo_kselect_list,strain_levels, ~.x %>% mutate(across(Strain, ~factor(., levels = .y)))) 
```



```{r, message=FALSE}
##Copy souporcell files for the optimum clusters investigation on igv - all except negative

soupo_kmerge <- map(sm_aln_nms_al, ~{
soupo_kselect_list[str_detect(names(soupo_kselect_list), paste0(.x, "_"))] %>% 
  imap(., ~{
    nms = .y
    .x %>% rename_with(~paste0(.,'_', nms), everything()) %>% rownames_to_column('bcode')
        }) %>% 
    purrr::reduce(., left_join, by = 'bcode')
  }) %>% 
  set_names(sm_aln_nms_al)

```



```{r, fig.width=4, fig.height=4}

strain_harmonize_fn <- function(strn_tbl = tbl1, strn1 = "Strain_2nd", strn2 = "Strain_final", out_strn = "n_strains") {
  
  strn_tbl %>% group_by(!!rlang::sym(strn1)) %>% 
            add_count(!!rlang::sym(strn2), name = "stf_n") %>%
            mutate(across(stf_n, ~case_when(str_detect(!!rlang::sym(strn2), "SC") & str_detect(!!rlang::sym(strn1), "Doublet|Negative") ~ 0,
                                            str_detect(!!rlang::sym(strn1), "SC") & str_detect(!!rlang::sym(strn2), "Doublet|Negative") ~ 0,
                                            str_detect(!!rlang::sym(strn1), "Doublet|Negative") & str_detect(!!rlang::sym(strn2), "Doublet|Negative") ~ 0,
                                            T ~ .)),
                   top_st = max(stf_n),
                   nw_st = case_when(!!rlang::sym(strn2) == "Doublet" | !!rlang::sym(strn2) == "Negative" ~ !!rlang::sym(strn2),
                                     top_st == stf_n & top_st != 0 ~ !!rlang::sym(strn1))) %>% 
            group_by(top_st, nw_st) %>%
            mutate(lab_uniq = row_number(),
                   across(lab_uniq, ~case_when(top_st == 0 | nw_st == "Doublet" | nw_st == "Negative" ~ NA, T ~ .))) %>%
            group_by(!!rlang::sym(strn2)) %>%
            mutate(top_st2 = case_when(top_st == stf_n & top_st != 0 & lab_uniq == 1 ~ top_st, T ~ 0), 
                   top_top_st = max(top_st2),
                   ty_brakr = case_when(!!rlang::sym(strn2) == "Doublet" | !!rlang::sym(strn2) == "Negative" ~ !!rlang::sym(strn2),
                                        top_top_st == stf_n & top_top_st != 0 ~ !!rlang::sym(strn1))) %>% 
        arrange(!!rlang::sym(strn2))  %>%
        fill(ty_brakr, .direction = "downup") %>%
        group_by(is.na(ty_brakr))%>%
        mutate(nro_count = data.table::rleid(!!rlang::sym(strn2))) %>% ## Gets the count of distinct items in !!rlang::sym(strn2) into row numbers within the 'is.na(ty_brakr)' groups
        ungroup()%>%
        mutate(sc_numbr = as.numeric(str_remove_all(!!rlang::sym(strn1), "Doublet|Negative|SC")))%>%
        mutate(!! out_strn:= case_when(is.na(ty_brakr) ~ paste0("SC",(max(sc_numbr, na.rm = T)+nro_count)), T ~ ty_brakr)) %>%
        # select(c("bcode", strn1, out_strn)) %>% 
        select(-c('stf_n', 'top_st',   'ty_brakr', 'is.na(ty_brakr)', 'nro_count')) #%>%
        #column_to_rownames('bcode')
  
}
```


```{r, fig.width=4, fig.height=4}

snk_pt_fn <- function(decod_tbl, strns_2_plt = c('Strain_2nd', 'n_strains'), labls = paste0("K=", opt_clusters_nms[[3]][1:2]), bp_face = c("bold", "plain"), ttl = "plot", strain_colours = strain_cols_n) {
  decod_tbl %>% 
        mutate(across(contains("train"), ~str_replace_all(., c("Doublet" = "Dbt", "Negative" = "Neg")))) %>%
        make_long(all_of(strns_2_plt)) %>% 
        add_count(x,node, "n") %>%
        ggplot(., aes(x = x, next_x = next_x,node = node,next_node = next_node,fill = factor(node),label = paste0(node,": ",n))) +
            geom_sankey() +
        theme_void()+
        scale_fill_manual(values = strain_colours) +
        scale_x_discrete(labels=labls)+
        geom_sankey_label(size = 3.6, color = 1, fill = "white") +
        labs(title = ttl) +
        theme_sankey(base_size = 9) +
        theme(legend.position = "none",
              axis.title.x = element_blank(),
              axis.text.x = element_text(size = 15, face = bp_face))
      
}

```


```{r, fig.width=4, fig.height=4}
soupo_seu_kmerge <- soupo_kmerge
```


```{r, fig.width=4, fig.height=4}
# don_k_names <- imap(opt_clusters_nms, ~paste0('Strain_',.y, '_', .x))
# don_hmonizd_k_names <- imap(opt_clusters_nms, ~paste0('Strain_hmo_',str_remove(.y, "SC"), '_', .x))
don_k_names_aln <- imap(opt_clusters_nms_aln, ~paste0('Strain_',.y, '_', .x))
don_hmonizd_k_names_aln <- imap(opt_clusters_nms_aln, ~paste0('Strain_hmo_',str_remove(.y, "SC"), '_', .x))

# strn_harmnzd_tbls <- list(soupo_seu_kmerge, don_k_names, don_hmonizd_k_names) %>%
strn_harmnzd_tbls <- list(soupo_seu_kmerge, don_k_names_aln, don_hmonizd_k_names_aln) %>%
  pmap(~{
    
    strain_harmonize_fn(strn_tbl = ..1, strn1 = ..2[1], strn2 = ..2[2], out_strn = ..3[2]) %>%
      strain_harmonize_fn(strn_tbl = ., strn1 = ..3[2], strn2 = ..2[3], out_strn = ..3[3])
   
  })
    
# list(strn_harmnzd_tbls, don_k_names, opt_clusters_nms, don_hmonizd_k_names) %>%
list(strn_harmnzd_tbls, don_k_names_aln, opt_clusters_nms_aln, don_hmonizd_k_names_aln) %>%
  pmap(~{  
    
snk_pt_fn(..1, strns_2_plt = c(..2[1], ..4[2:3]), labls = paste0("K=", ..3[1:3]), bp_face = c("bold", "plain", "plain"))
    
 })

```



```{r, fig.width=6, fig.height=4}

strn_hm_tbl_fn <- function(soupo_seu_kmerge_f, don_k_names_aln_f, don_hmonizd_k_names_aln_f){
  list(soupo_seu_kmerge_f, don_k_names_aln_f, don_hmonizd_k_names_aln_f) %>%
    pmap(~{
      
      if (length(..2) == 3) { 
          
            strain_harmonize_fn(strn_tbl = ..1, strn1 = ..2[1], strn2 = ..2[2], out_strn = ..3[2]) %>%
            strain_harmonize_fn(strn_tbl = ., strn1 = ..3[2], strn2 = ..2[3], out_strn = ..3[3])
         
          } else if (length(..2) == 4) {
          
            strain_harmonize_fn(strn_tbl = ..1, strn1 = ..2[1], strn2 = ..2[2], out_strn = ..3[2]) %>%
            strain_harmonize_fn(strn_tbl = ., strn1 = ..3[2], strn2 = ..2[3], out_strn = ..3[3])%>%
            strain_harmonize_fn(strn_tbl = ., strn1 = ..3[3], strn2 = ..2[4], out_strn = ..3[4])
       
          } else if  (length(..2) == 5) {
          
            strain_harmonize_fn(strn_tbl = ..1, strn1 = ..2[1], strn2 = ..2[2], out_strn = ..3[2]) %>%
            strain_harmonize_fn(strn_tbl = ., strn1 = ..3[2], strn2 = ..2[3], out_strn = ..3[3])%>%
            strain_harmonize_fn(strn_tbl = ., strn1 = ..3[3], strn2 = ..2[4], out_strn = ..3[4])%>%
            strain_harmonize_fn(strn_tbl = ., strn1 = ..3[4], strn2 = ..2[5], out_strn = ..3[5])
         
          } else {
          print("Not accounted for")
      }
    })
}


```



```{r, fig.width=6, fig.height=4}

strn_harmnzd_tbls <- strn_hm_tbl_fn(soupo_seu_kmerge_f = soupo_seu_kmerge, don_k_names_aln_f =don_k_names_aln, don_hmonizd_k_names_aln_f = don_hmonizd_k_names_aln)
```

```{r, fig.width=6, fig.height=4, eval=T}
snk_plts <- list(strn_harmnzd_tbls, don_k_names_aln, opt_clusters_nms_aln, don_hmonizd_k_names_aln) %>%
  pmap(~{  
    
snk_pt_fn(..1, strns_2_plt = c(..2[1], ..4[2:length(..4)]), labls = paste0("K=", ..3[1:5]), bp_face = c("plain"))
    
 })

```

```{r, eval=T}
snk_plts[str_detect(names(snk_plts), "minmap")] %>% imap(., ~.x + labs(title = .y))
```

```{r, eval=T}
snk_plts[str_detect(names(snk_plts), "hsat")] %>% imap(., ~.x + labs(title = .y))
```


```{r, fig.height=6, fig.width=24, eval=T}
## Save for visualization of sankey plots for the manuscript
saveRDS(snk_plts, "data/processed/Pf/m2_all/strain_calling/snk_plts.RDS")

```


```{r}
#######!!!!!!! NOTE - Defined in header file
## Select most likely optimum k
# optimum_k = c("MSC33" =8, "MSC48" =5, "MSC49" =6, "MSC55" =7, "MSC60" =8, "MSC66" =9, "MSC38" =4, "MSC40" =9, "MSC50_MACS" =2, "MSC53" =1, "MSC57" =1, "MSC24" =1, "MSC37" =5, "MSC39" =4, "MSC45" =1, "MSC50_SLO" =2, "MSC54" =1, "MSC67" =1, "MSC68" =3, "MSC70" =1, "MSC41"= 3, "MSC51"= 1, "MSC25"= 2, "MSC1" = 5, "MSC3" = 2, "MSC13" = 2, "MSC14" = 8, "MSC12" = 2) %>% .[sample_name_nms]
# 
# optimum_k_aln = rep(optimum_k, length(algn_nms)) %>% set_names(sm_aln_nms_al)

```


```{r, fig.width=6, fig.height=4}
## Rename the initial k colnames which is unchanged when harmonizing strain annotation for Sankey plot and factorized newly generated strain harmonized vectors

strn_harmnzd_tbls <- list(strn_harmnzd_tbls, map(don_k_names_aln, ~.x[1]), map(don_hmonizd_k_names_aln, ~.x[1])) %>%
  pmap(~{
    ..1 %>% 
      mutate(!! ..3:= !!rlang::sym(..2)) %>%
      mutate(across(contains("Strain_hmo"), ~droplevels(factor(.,levels = strain_levels_univrs))))
    
  })

```


```{r, fig.height=6, fig.width=24, eval=T}
## Save for visualization of sankey plots for the manuscript
saveRDS(strn_harmnzd_tbls, "data/processed/Pf/m2_all/strain_calling/sstrn_harmnzd_tbls.RDS")
saveRDS(don_k_names_aln, "data/processed/Pf/m2_all/strain_calling/don_k_names_aln.RDS")
saveRDS(opt_clusters_nms_aln, "data/processed/Pf/m2_all/strain_calling/opt_clusters_nms_aln.RDS")
saveRDS(don_hmonizd_k_names_aln, "data/processed/Pf/m2_all/strain_calling/don_hmonizd_k_names_aln.RDS")
```


```{r, fig.width=6, fig.height=4}
## !! NOTE - VERY IMPORTANT to verify that the strain renaming to harmonize names for sankey plot function has been done correctly

strn_harmnzd_assesmnt <- list(strn_harmnzd_tbls, don_k_names_aln, don_hmonizd_k_names_aln) %>%
  pmap(~{
    tbl <- ..1
    nms <- ..2
    nms_hmo <- ..3
    
    map2(nms, nms_hmo,  ~{
      tbl %>% 
        mutate("strn_o" =!!rlang::sym(.x) , "strn_n" = !!rlang::sym(.y)) %>% 
        dplyr::count(strn_o, strn_n) %>% 
        arrange(strn_o)
      }) %>% 
      set_names(nms) %>% 
      bind_rows(., .id = "strn_k") 
    
  })

strn_harmnzd_assesmnt
```

```{r, fig.width=6, fig.height=4}
## !! NOTE - VERY IMPORTANT to verify that the strain renaming to harmonize names for sankey plot function has been done correctly - Detailed assesment - may be time consuming for many donor samples
bind_rows(strn_harmnzd_assesmnt) #%>% view
```

```{r, fig.width=6, fig.height=4}
## !! NOTE - VERY IMPORTANT to verify that the strain renaming to harmonize names for sankey plot function has been done correctly - Summary of table above
hmonizd_k_test_summary <- bind_rows(strn_harmnzd_assesmnt) %>% group_by(strn_k) %>% mutate(dup = duplicated(strn_o) | duplicated(strn_n)) %>% count(dup) %>% mutate(k=n-2) %>% mutate(k_test = as.numeric(str_remove(strn_k, "Str.*_")) == k)
hmonizd_k_test_summary

```

```{r, fig.width=6, fig.height=4}
## !! NOTE - VERY IMPORTANT to verify that the strain renaming to harmonize names for sankey plot function has been done correctly - Verify those that are FALSE
table(hmonizd_k_test_summary$k_test)
```


Harmonize minimap and hisat labels for each k - change hisat2 labels to match those of minimap2

merge minimap and hisat harmonized strain labels for further harmonization between the 2 aligners
```{r , fig.width = 10, fig.height = 5, warning=FALSE}
mnmp_hsat_comp_tbl <- list(strn_harmnzd_tbls[str_detect(names(strn_harmnzd_tbls), "minmap")], strn_harmnzd_tbls[str_detect(names(strn_harmnzd_tbls), "hsat")]) %>% 
  pmap(., ~..1 %>% 
         left_join(., ..2, by = "bcode") %>%
         select(bcode, starts_with("Strain_hmo")) 
       )
mnmp_hsat_comp_tbl[[1]]
```

```{r, fig.width=4, fig.height=4}
# don_k_names <- imap(opt_clusters_nms, ~paste0('Strain_',.y, '_', .x))
# don_hmonizd_k_names <- imap(opt_clusters_nms, ~paste0('Strain_hmo_',str_remove(.y, "SC"), '_', .x))
don_k_names_minmap <- imap(opt_clusters_nms_aln[str_detect(names(opt_clusters_nms_aln), "minmap")], ~paste0('Strain_hmo_', str_remove(.y, "SC"), '_', .x)) %>% unlist(., recursive = F)
don_k_names_hsat <- imap(opt_clusters_nms_aln[str_detect(names(opt_clusters_nms_aln), "hsat")], ~paste0('Strain_hmo_', str_remove(.y, "SC"), '_', .x)) %>% unlist(., recursive = F)


don_hmonizd_minmap_base <- imap(opt_clusters_nms_aln[str_detect(names(opt_clusters_nms_aln), "hsat")], ~paste0('Strain_hmo_', str_replace(str_remove(.y, "SC"), "hsat","hsatHMO"), '_', .x)) %>% unlist(., recursive = F)

mnmp_hsat_comp_tbl_repd <-  map2(mnmp_hsat_comp_tbl, opt_clusters_nms_aln[str_detect(names(strn_harmnzd_tbls), "minmap")],  ~rep(list(.x),  length(.y))) %>% unlist(., recursive=F)
mnmp_hsat_comp_tbl_repd <- mnmp_hsat_comp_tbl_repd %>% set_names(map2(names(mnmp_hsat_comp_tbl_repd), unlist(opt_clusters_nms_aln[str_detect(names(strn_harmnzd_tbls), "minmap")]), ~str_replace_all(.x, "minmap.*", paste0("K", .y))))


# strn_harmnzd_tbls <- list(soupo_seu_kmerge, don_k_names, don_hmonizd_k_names) %>%
strn_harmnzd_tbls_mn_hs <- list(mnmp_hsat_comp_tbl_repd, don_k_names_minmap, don_k_names_hsat, don_hmonizd_minmap_base) %>%
  pmap(~{
    
    strain_harmonize_fn(strn_tbl = ..1, strn1 = ..2, strn2 = ..3, out_strn = ..4) 
    
  })
    

strn_harmnzd_tbls_mn_hs[[1]]
```


```{r , fig.width = 10, fig.height = 5, warning=FALSE}

## Select relevant variables
strn_harmnzd_tbls_mn_hs2 <- pmap(list(strn_harmnzd_tbls_mn_hs, don_k_names_minmap, don_k_names_hsat, don_hmonizd_minmap_base), ~..1 %>% select(bcode, ..2, ..3, ..4))

```


```{r , fig.width = 10, fig.height = 5, warning=FALSE}


# strn_harmnzd_tbls_mn_hs2[["MSC60_K9"]] %>%
#          count(Strain_hmo_M60_minmap_9, Strain_hmo_M60_hsatHMO_9)

```


Rechecking the overlap of strain assignments between different mappers after harmonization of hisat strain names to match those of minimap2
```{r , fig.width = 10, fig.height = 5, warning=FALSE}
aln_comp_tbl <- map(strn_harmnzd_tbls_mn_hs2, ~.x %>% 
                      select(bcode, contains("_minmap_"), contains("hsatHMO")) %>%
        pivot_longer(cols = starts_with("Strain_hmo"),
         names_to = c(".value", "K"),
         names_pattern = ".*[0-9].*_(.*)_([0-9]+$)")
        ) 

imap(aln_comp_tbl, possibly(~phmap_lab_transfr(.x, col_labs= `minmap`, row_labs=`hsatHMO`, row_labs_c ='hsatHMO', col_labs_c= 'minmap', ttl = .y)), 'k=1 so no groups to compare which is necessary for the heatmap')

```



```{r, fig.width=6, fig.height=4, eval=T}
# list(strn_harmnzd_tbls_5, don_k_names, opt_clusters_nms, don_hmonizd_k_names) %>%
snk_plts_hmo <- list(strn_harmnzd_tbls_mn_hs2, don_k_names_minmap,  don_k_names_hsat, str_remove_all(don_k_names_hsat, "S.*_"), don_hmonizd_minmap_base) %>%
  pmap(~{  
    
  snk_pt_fn(..1, strns_2_plt = c(..2, ..5, ..3), labls = c("minimap", "hisatHMO", "hisat"), bp_face = c("plain"))
    
 })

```


```{r, eval=T}
# snk_plts[str_detect(names(snk_plts), "minmap")]
snk_plts_hmo <- snk_plts_hmo %>% imap(., ~.x + labs(title = .y))
```


```{r, fig.width=6, fig.height=4, eval=T}
# list(strn_harmnzd_tbls_5, don_k_names, opt_clusters_nms, don_hmonizd_k_names) %>%
snk_plts_hmo[str_detect(names(snk_plts_hmo), "MSC14|MSC49|MSC55|MSC60")]
```

#### Adding stage assignments to the sankey plots to see how they transition across Ks and alignment method

Rechecking the overlap of strain assignments between different mappers after harmonization of strain names
hisat and hisat harmonized  should be 100% identical to confirm no errors in harmonizing labels 
```{r , fig.width = 10, fig.height = 5, warning=FALSE}
aln_comp_tbl <- map(strn_harmnzd_tbls_mn_hs2, ~.x %>% 
                      select(bcode, contains("_hsat_"), contains("hsatHMO")) %>%
        pivot_longer(cols = starts_with("Strain_hmo"),
         names_to = c(".value", "K"),
         names_pattern = ".*[0-9].*_(.*)_([0-9]+$)")
        ) 

imap(aln_comp_tbl, possibly(~phmap_lab_transfr(.x, col_labs= `hsat`, row_labs=`hsatHMO`, row_labs_c ='hsatHMO', col_labs_c= 'hsat', ttl = .y)), 'k=1 so no groups to compare which is necessary for the heatmap')

```


```{r, message=FALSE}
##Coalesce the genotypes and get a consensus genotype taking labels for highest resolution K
##VARIABLE
## !!!NB!!!!!When trialing out different souporcell Ks use Strain_qc_cutoff > 20 but once decided use Strain_qc_cutoff > 0 the optimum K 

Strain_qc_cutoff=0
# optimum_k = c(8,5,6,7,7) %>% set_names(sample_name_nms)

# opt_narrow_clusters_nms <-  list("MSC33" = c(7:9),"MSC48" = c(4:6),"MSC49" = c(3:5),"MSC55" = c(4:6),"MSC60" = c(3:5))
# opt_narrow_clusters_nms_aln <- rep(opt_narrow_clusters_nms, length(algn_nms)) %>% set_names(sm_aln_nms_al)


strn_c_mdata <- list(strn_harmnzd_tbls, 
                     names(strn_harmnzd_tbls), 
                     opt_clusters_nms_aln[names(strn_harmnzd_tbls)] %>% map(.,max), 
                     optimum_k_aln, 
                     imap(opt_narrow_clusters_nms_aln, ~paste0(.y, '_', .x))) %>%
  pmap(~{
    ..1 %>%
      unite(col = 'Strn_c', starts_with("Strain_hmo_M"), sep='_', remove = F) %>% 
      add_count(Strn_c) %>% 
      mutate(
        'Strain_max_qc' = case_when( 
          !str_detect(Strn_c, 'Negative|Doublet') & n>Strain_qc_cutoff ~ !!rlang::sym(paste0('Strain_hmo_', str_remove(..2, 'SC'), '_',..3)), 
          T ~ 'PoorQC'),
        'Strain_qc' = case_when(
          !str_detect(!!rlang::sym(paste0('Strain_hmo_', str_remove(..2, 'SC'), '_',..4)), 'Negative|Doublet') & n>Strain_qc_cutoff ~ !!rlang::sym(paste0('Strain_hmo_', str_remove(..2, 'SC'), '_',..4)), 
          T ~ 'PoorQC'),
        'Strain_DN' = !!rlang::sym(paste0('Strain_hmo_', str_remove(..2, 'SC'), '_',..4)),
        'dbt_neg_all' = case_when(
          if_all(c(paste0('Strain_', ..5)), ~ .== "Doublet") ~ "DoubletAll",
          if_all(c(paste0('Strain_', ..5)), ~ .== "Negative") ~ "NegativeAll",
          if_all(c(paste0('Strain_', ..5)), ~ .== "Doublet" | .== "Negative") ~ "DbtNegMix", 
          T ~ 'LikelySinglet')
        )
  })



```


```{r}
## visualize top clusters
# list(strn_c_mdata, sample_name_nms, optimum_k)  %>% 
list(strn_c_mdata, sm_aln_nms_al, optimum_k_aln)  %>% 
  purrr::pmap(.,~{
  ..1 %>% dplyr::count(Strn_c, Strain_qc, !!rlang::sym(paste0("Strain_hmo_", str_remove(..2, "SC"), "_", ..3))) %>% arrange(desc(n))
})
```



```{r}
## Doublets confirmed removal round 2
## visualize top clusters
list(strn_c_mdata, sm_aln_nms_al, optimum_k_aln)  %>% 
  purrr::pmap(.,~{
  ..1 %>% filter(if_all(starts_with('Strain_hmo_M'), ~ .== "Doublet" | .== "Negative")) %>% dplyr::count(Strn_c, dbt_neg_all, !!rlang::sym(paste0("Strain_hmo_", str_remove(..2, "SC"), "_", ..3))) 
})
```


```{r}
## visualize top clusters
list(strn_c_mdata, sm_aln_nms_al, optimum_k_aln)  %>% 
  purrr::pmap(.,~{
  ..1 %>% dplyr::count(Strn_c, Strain_qc, !!rlang::sym(paste0("Strain_hmo_", str_remove(..2, "SC"), "_", ..3))) %>% arrange(desc(n))
})
```




### PCA calculation
```{r}
# soupo_PC_mtx <- soupc_all %>% 
soupo_PC_mtx <- soupc_all_al %>% 
  dplyr::select(-c(status, assignment, log_prob_singleton, log_prob_doublet)) %>% 
  dplyr::group_split(sample_k) %>%
  purrr::set_names(purrr::map_chr(., ~.x$sample_k[1])) %>%
  map(. %>% column_to_rownames(., "barcode") %>%
        dplyr::select(where(~!all(is.na(.)))) %>%
        dplyr::select(-sample_k) %>%
        mutate(across(where(is.numeric), ~ (. / rowSums(across(where(is.numeric)))))) %>%
        as.matrix() %>%
        t())

soupo_PC_mtx <- soupo_PC_mtx[as.character(sorted_names_al)]
```



```{r}
##Calculate PCA
soupo_PCA <- soupo_PC_mtx %>% 
  map(. %>% prcomp(.))
```


Calculate proportion of variance explained by PC1 and PC2
```{r}
# Assuming your data is stored in a matrix called 'data'
# Extract the standard deviations of the principal components
# pc_standard_deviations <- soupo_PCA[[9]]$sdev
# 
# #PC1 variance explained
# (pc_standard_deviations[1] ^ 2) / sum(pc_standard_deviations^2)
# 
# #PC2 variance explained
# (pc_standard_deviations[2] ^ 2) / sum(pc_standard_deviations^2)
```

```{r}
## make directory for strain summary output
system(paste0("mkdir -p data/processed/",species, "/" , sv_dir))

```

```{r}
## Save PCA
saveRDS(soupo_PCA, paste0("data/processed/", species, "/" , sv_dir,"/soupo_PCA_m22.RDS"))
```



```{r}
## visualize top clusters
list(strn_c_mdata)  %>% 
  purrr::pmap(.,~{
  ..1 %>% dplyr::count(Strain_qc) %>% arrange(desc(n))
})
```

```{r}
## visualize top clusters
# list(strn_c_mdata, sample_name_nms, optimum_k)  %>% 
list(strn_c_mdata, sm_aln_nms_al, optimum_k_aln)  %>% 
  purrr::pmap(.,~{
  ..1 %>% dplyr::count(Strn_c, Strain_qc, !!rlang::sym(paste0("Strain_hmo_", str_remove(..2, "SC"), "_", ..3))) %>% arrange(desc(n))
})
```


###msc3 = SC1+SC3
###ms13 = SC2+SC3
```{r}
##Add the number per xluster as a column in metadata

map(strn_c_mdata, . %>% dplyr::count(Strain_DN))

```

```{r}

# ##Save strain + new metadata as RDS
strn_c_mdata <- map(strn_c_mdata,~mutate(., 'strain' = Strain_DN))


```


```{r}
##Add the number per xluster as a column in metadata

map(strn_c_mdata, . %>% 
      dplyr::count(strain) 
)

map(strn_c_mdata, dim)

```


```{r}
##Save strain + new metadata as RDS
map(strn_c_mdata, 
    . %>% 
      dplyr::select(-c(starts_with("Strain_MSC"))) %>%
      dplyr::select("bcode", "Strn_c", starts_with("Strain"), "strain", "dbt_neg_all")) %>%
  saveRDS(paste0("data/processed/", species, "/" , sv_dir, "/strn_c_new_mdata_m22_no_stage.RDS"))


```
