---
title: "MSC QC and preprocessing"
date: '2022-07-11'
author: 'Jesse Rop'
output:
  html_document:
    toc: true
    theme: united
    code_folding: hide
    df_print: paged
---

```{r}
suppressPackageStartupMessages({
  library(Seurat)
  library(sctransform)
  library(glmGamPoi)
  # library(rgl)
  library(patchwork)
  # library(DoubletFinder)
  library(ggrepel)
  library(ggpubr)
  library(dplyr) 
  library(stringr)
  library(purrr)
  library(readr)
  library(tibble)
  library(ggplot2)
  library(tidyr)
  library(cowplot)
  # library(scDblFinder)
  library(SeuratData)
  library(SeuratDisk)
  library(plotly)
  library(Hmisc)
  library(rmarkdown)
  library(RColorBrewer)
  # library(Platypus)
  library(ggvenn)
  library(conflicted)
  # library(SoupX)
  library(MetBrewer)
  library(ggh4x)
  library(SingleCellExperiment)
  # library(scmap)
  # library(ggsankey)
  # library(findPC)
  library(BPCells)
  # library(furrr)
  # library(future.batchtools)
  library(parallel)
  library(EnhancedVolcano)
  library(DropletUtils)
  library(ggforce)
  library(scran)
  library(SingleR)
  library(scDblFinder)
  library(ggsankey)
  library(findPC)
  
  conflict_prefer("select", "dplyr")
  conflict_prefer("filter", "dplyr")
  conflict_prefer("rename", "dplyr")
  conflict_prefer("map", "purrr")
  conflict_prefer("count", "dplyr")
  conflicts_prefer(base::saveRDS)
  conflicts_prefer(SingleCellExperiment::colData)
  conflicts_prefer(base::as.data.frame)
  conflicts_prefer(base::intersect)
  conflicts_prefer(dplyr::desc)
  conflicts_prefer(dplyr::count)
})
```



# MSC stage annotation and gene expression QC 

## setup, read input files and exploratory visualizations and general functions

```{r, setup}
knitr::opts_knit$set(root.dir = '/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/Talleh')

```

### Variable declarations


#### Source common functions, variables and tables
```{r}
##Boxplot labeling function
source("~/Talleh_code/pf_common_vars.R")
source("~/Talleh_code/project_common_fns.R")
```

## Read in sample name details

```{r}
# Select sample names
sample_name <- c("MSC49", "MSC68")

```


```{r, eval=T}
##Read annotated seurat object
bp_prelim_stg <- list.files("data/processed/Pf", pattern = "bp_prelim_stg.RDS$", full.names = T, recursive = T, include.dirs = T) %>% 
  set_names(nm = (str_remove_all(.,'data/processed/Pf/|/seu_obj/|bp_prelim_stg.*.RDS'))) %>% 
  .[sample_name] %>%
  map(readRDS)
```


```{r, eval=T}
all_seu_prelim_stg_sct <- map(bp_prelim_stg, ~sct_run_fn(Seurat_object = .x))
all_seu_prelim_stg_sct
```



```{r, warning=FALSE, message=FALSE}
imap(all_seu_prelim_stg_sct, ~DimPlot(.x, dims = c(1,2), group.by = "stage_fld", pt.size = 0.1, reduction = "pca")+ scale_color_manual(values = sp_col_new) + labs(title = .y))
imap(all_seu_prelim_stg_sct, ~DimPlot(.x, dims = c(1,2), group.by = "stage_fld", pt.size = 0.1)+ scale_color_manual(values = sp_col_new) + labs(title = .y))
```

```{r, warning=FALSE, message=FALSE}

all_seu_prelim_stg_sct_umap_mdata <- map(all_seu_prelim_stg_sct, ~{
  .x@meta.data %>% 
    rownames_to_column("bcode") %>%
    left_join(., .x@reductions$umap@cell.embeddings %>% data.frame() %>% rownames_to_column("bcode"), by = "bcode")
})

```


```{r, message=FALSE, warning=FALSE, eval=T}
# Save the poorQC cells barcodes for plotting
saveRDS(all_seu_prelim_stg_sct_umap_mdata, "data/processed/Pf/m2_all/all_seu_prelim_stg_sct_umap_mdata.RDS")

```

```{r, message=F, warning=F, eval=T}
## Get mdata file list
all_seu_prelim_stg_norm_list_files <- Sys.glob("data/processed/Pf/*/seu_obj/bp_prelim_stg_norm.RDS")

all_seu_prelim_stg_norm_list_files
```

```{r, message=F, warning=F, eval=T}
## read filtered barcodes
all_seu_prelim_stg_norm <- map(all_seu_prelim_stg_norm_list_files, ~.x %>% 
                                 set_names(nm = (str_remove_all(.,'data/processed/Pf/|/seu_obj.*')))) %>% 
  unlist(., recursive = F) %>% 
  .[sample_name] %>%
  .[!is.na(.)] %>%
  map(readRDS) 

```


```{r, warning=FALSE, message=FALSE}
imap(all_seu_prelim_stg_norm, ~DimPlot(.x, dims = c(2,3), group.by = "stage_fld", pt.size = 0.1, reduction = "pca")+ scale_color_manual(values = sp_col_new) + labs(title = .y))
imap(all_seu_prelim_stg_norm, ~DimPlot(.x, dims = c(1,2), group.by = "stage_fld", pt.size = 0.1)+ scale_color_manual(values = sp_col_new) + labs(title = .y))
```

```{r, warning=FALSE, message=FALSE}

all_seu_prelim_stg_norm_umap_mdata <- map(all_seu_prelim_stg_norm, ~{
  .x@meta.data %>% 
    rownames_to_column("bcode") %>%
    left_join(., .x@reductions$umap@cell.embeddings %>% data.frame() %>% rownames_to_column("bcode"), by = "bcode")
})
```


```{r, message=FALSE, warning=FALSE, eval=T}
# Save the poorQC cells barcodes for plotting
saveRDS(all_seu_prelim_stg_norm_umap_mdata, "data/processed/Pf/m2_all/all_seu_prelim_stg_norm_umap_mdata.RDS")

```


```{r, warning=FALSE, message=FALSE, fig.width=12, fig.height=3.5}
## swap sct for norm standardisation
all_seu_prelim_stg_norm <- all_seu_prelim_stg_sct

for (i in 1:length(all_seu_prelim_stg_norm)) {
  DefaultAssay(all_seu_prelim_stg_norm[[i]]) <- "RNA"
  all_seu_prelim_stg_norm[[i]] <- NormalizeData(all_seu_prelim_stg_norm[[i]])
}
```

```{r, warning=FALSE, message=FALSE, fig.width=12, fig.height=3.5}
#SURF1.2 gexp02
imap(all_seu_prelim_stg_norm, ~FeaturePlot(.x, reduction = "umap",  label = T, features = c("PF3D7-0113600", "PF3D7-1102500", "PF3D7-1222600"), order = T, ncol = 3) + labs(title = .y))
```

```{r}
all_seu_prelim_stg_norm[1] %>% map(., ~.x@meta.data %>% 
                                     count(stage_final2, combined_labs, stage_fld) %>% 
                                     filter(str_detect(stage_fld, "LE"))
)

```


```{r, warning=FALSE, message=FALSE, fig.width=4, fig.height=4.5}
## in cellbender but not in edrops
plts_stg <- imap(all_seu_prelim_stg_norm, possibly(~{
  cell_stg_plot <- DimPlot(.x, group.by = "stage_fld", cols = sp_col_new, pt.size = 0.2) + 
    theme_void() + 
    theme(legend.position = "none", panel.border = element_rect(color = "black", fill = NA, size = 1))  +
    labs(title = .y)
  
  cell_stg_plot$data <- cell_stg_plot$data %>% rownames_to_column('bcode') %>% left_join(., .x@meta.data[,'seurat_clusters', drop = F] %>% rownames_to_column('bcode'), by='bcode') %>% column_to_rownames('bcode')
  
  cell_stg_plot<- LabelClusters(cell_stg_plot, id = "seurat_clusters", size = 6, repel = T)
  return(cell_stg_plot)
  
}, "no cells"))

```

```{r, warning=FALSE, message=FALSE, fig.height= 3.5, fig.width= 10, eval=T}
plot_grid(plotlist = plts_stg, ncol = 4)
```



## 1. MITOCHONDRIAL PERCENT VISUALIZATION AND CELL REMOVAL

## !!!!!!NB - RECONSIDER FILTRATION PROPERLY LATER
```{r, eval=FALSE}
## !!!!!!NB -CONSIDER FORE REMOVAL LATER
#filter out cells with high ncountRNA 
# all_seu_prelim_stg_norm <- all_seu_prelim_stg_norm %>%
#   map2(., ncount_filt, ~.x  %>%
#          subset(., pf_umi_count < .y )
#   )

```

Get mt_filt based on automatic removal of quantiles
```{r}
all_seu_prelim_stg_norm <- all_seu_prelim_stg_norm%>% 
  map(., ~{
    
    mit_fts <- rownames(.x)[str_detect(rownames(.x), "MIT")]
    api_fts <- rownames(.x)[str_detect(rownames(.x), "API")]
    
    .x %>% PercentageFeatureSet(., features = mit_fts, col.name = 'percent.mt', assay = "RNA") %>%
      PercentageFeatureSet(., features = api_fts, col.name = 'percent.api', assay = "RNA") 
  })

```

```{r}
# mt_filt <- unlist(map(all_seu_prelim_stg_norm, ~quantile(.x@meta.data[, "percent.mt"], 0.975, na.rm = T)))
mt_filt <- unlist(map(all_seu_prelim_stg_norm, ~quantile(.x@meta.data[, "percent.mt"], 0.995, na.rm = T)))
# api_filt <- unlist(map(all_seu_prelim_stg_norm, ~quantile(.x@meta.data[, "percent.api"], 0.995, na.rm = T)))

```


```{r, results='asis'}
std_theme <- theme(axis.text=element_text(size=18), 
                   axis.title=element_text(size=18,face="bold"), 
                   legend.text = element_text(size=22), 
                   legend.title = element_text(size=22,face="bold"))


```

```{r, results='asis'}
pmap(list(all_seu_prelim_stg_norm, mt_filt, names(all_seu_prelim_stg_norm)), ~{
  FeatureScatter(..1, feature1 = "pf_umi_count", feature2 = "percent.mt", group.by = "stage_fld", pt.size = 0.1)+ 
    scale_color_manual(values = sp_col_new) +
    geom_hline(yintercept = ..2)  +
    labs(title = ..3) +
    std_theme
  
})

```


```{r, eval=T}

after_umi100_filt <- map(all_seu_prelim_stg_norm, dim)%>% as.data.frame() %>% .[2,]
after_umi100_filt
```



```{r}
map2(all_seu_prelim_stg_norm, mt_filt, ~.@meta.data %>% dplyr::count(percent.mt>.y ))
```

######## !!!!!!NB - RECONSIDER FILTRATION PROPERLY LATER

```{r}
## Barcodes removed
mt_filt_rm_bcodes <- map2(all_seu_prelim_stg_norm, mt_filt, ~colnames(subset(.x, percent.mt < .y, invert = T ) ))

```

## Compare the original sct object that was shared with talleh to the new seurat object that was created with the same counts and meta.data but with the new stage annotations to check that the stage annotations are the same between the two objects


```{r, eval=T}
## Generated in ~/mali22_code_lnk/for_talleh_mapping_analysis/m2_prelim_stg_annotn_ta19.Rmd
msc49_ta19 <- readRDS("data/processed/Pf/MSC49/seu_obj/msc49_ta19.RDS")
```

```{r, eval=T}
all_seu_prelim_stg_norm[["MSC49"]]@meta.data %>% rownames_to_column("bcode") %>% right_join(msc49_ta19@meta.data %>% rownames_to_column("bcode"), by = "bcode", suffix = c("_nw", "_old")) %>% ggplot(aes(x = percent.mt_nw, percent.mt_old, color = stage_fld)) + geom_point() + geom_abline()
```

```{r, eval=T}
# DefaultAssay(msc49_ta19) <- "RNA"
msc49_ta19 <- PercentageFeatureSet(msc49_ta19, features = rownames(msc49_ta19)[str_detect(rownames(msc49_ta19), "MIT")], col.name = 'percent.mt_rna', assay = "RNA")
```

```{r, eval=T}
msc49_ta19@meta.data %>% ggplot(aes(x = nCount_RNA, pf_umi_count)) + geom_point() + geom_abline()
msc49_ta19@meta.data %>% ggplot(aes(x = percent.mt, percent.mt_rna)) + geom_point() + geom_abline()
```

```{r, eval=T}
all_seu_prelim_stg_norm[["MSC49"]]@meta.data %>% ggplot(aes(x = nCount_RNA, pf_umi_count)) + geom_point() + geom_abline()

```

```{r, eval=T}
all_seu_prelim_stg_norm[["MSC49"]]@meta.data %>% rownames_to_column("bcode") %>% right_join(msc49_ta19@meta.data %>% rownames_to_column("bcode"), by = "bcode", suffix = c("_nw", "_old")) %>% ggplot(aes(x = percent.mt_nw, percent.mt_rna, color = stage_fld)) + geom_point(size = 0.01) + geom_abline()
```

```{r, eval=T}
all_seu_prelim_stg_norm[["MSC49"]]@meta.data %>% rownames_to_column("bcode") %>% right_join(msc49_ta19@meta.data %>% rownames_to_column("bcode"), by = "bcode", suffix = c("_nw", "_old")) %>% ggplot(aes(x = nFeature_RNA_nw, nFeature_RNA_old, color = stage_fld)) + geom_point() + geom_abline()
```

```{r, eval=T}
FetchData(all_seu_prelim_stg_norm[["MSC49"]], vars = "PF3D7-MIT02100", layer = "counts") %>% rownames_to_column("bcode") %>% right_join(., FetchData(msc49_ta19, vars = "PF3D7-MIT02100", layer = "counts") %>% rownames_to_column("bcode"), by = "bcode", suffix = c("_nw", "_old")) %>% ggplot(aes(x = `PF3D7-MIT02100_nw`, `PF3D7-MIT02100_old`)) + geom_point() + geom_abline()

```

```{r, eval=T}
all_seu_prelim_stg_norm[["MSC49"]]@meta.data %>% rownames_to_column("bcode") %>% right_join(msc49_ta19@meta.data %>% rownames_to_column("bcode"), by = "bcode", suffix = c("_nw", "_old")) %>% ggplot(aes(x = pf_umi_count_nw, pf_umi_count_old, color = stage_fld)) + geom_point() + geom_abline()
```

```{r, eval=T}
## !!!!!!NB -CONSIDER FORE REMOVAL LATER
#filter out cells with high mitochondrial count
all_seu_prelim_stg_norm <- all_seu_prelim_stg_norm %>%
  map2(., mt_filt, ~.x  %>%
         subset(., percent.mt < .y )
  )

```

```{r, eval=T}

after_mt_filt <- map(all_seu_prelim_stg_norm, dim)%>% as.data.frame() %>% .[2,]
after_mt_filt
```


```{r, eval=T}
all_seu_prelim_stg_norm[["MSC49"]]@meta.data %>% rownames_to_column("bcode") %>% right_join(msc49_ta19@meta.data %>% rownames_to_column("bcode"), by = "bcode", suffix = c("_nw", "_old")) %>% count(sR_fine_stg_nw, sR_fine_stg_old)
```


```{r, eval=T}
all_seu_ta_mt_filt_missing_subset <- msc49_ta19[,!(colnames(msc49_ta19) %in% colnames(all_seu_prelim_stg_norm[["MSC49"]]))]
all_seu_ta_mt_filt_missing_subset
```



```{r, warning=FALSE, message=FALSE}
imap(all_seu_prelim_stg_norm, ~DimPlot(.x, dims = c(1,3), group.by = "stage_fld", pt.size = 0.1)+ scale_color_manual(values = sp_col_new) + labs(title = .y))
imap(all_seu_prelim_stg_norm, ~DimPlot(.x, dims = c(1,2), group.by = "stage_fld", pt.size = 0.1)+ scale_color_manual(values = sp_col_new) + labs(title = .y))
```


```{r, warning=FALSE, message=FALSE}
imap(all_seu_prelim_stg_norm, ~VlnPlot(.x, features = "pf_gn_count", group.by = "stage_fld", pt.size = 0.1)+ scale_color_manual(values = sp_col_new) + labs(title = .y))
```


```{r, warning=FALSE, message=FALSE}
imap(all_seu_prelim_stg_norm, ~VlnPlot(.x, features = "nFeature_RNA", group.by = "stage_fld", pt.size = 0.1) + scale_color_manual(values = sp_col_new) + scale_y_log10() +  labs(title = .y))
```

```{r, warning=FALSE, message=FALSE}
# imap(all_seu_prelim_stg_norm, ~VlnPlot(.x, features = "sim_lab1_mca", group.by = "stage_fld", pt.size = 0.1)+ scale_color_manual(values = sp_col_new) + labs(title = .y))
```
### 2.1 GENE UMI COUNT VISUALIZATIONS

```{r, fig.height = 3.5, fig.width = 12}

all_seu_prelim_stg_norm_mdata_mrg <- map(all_seu_prelim_stg_norm, ~.@meta.data) %>% 
  bind_rows(., .id = 'donor') 

```

```{r, fig.height = 3.5, fig.width = 12}
all_seu_prelim_stg_norm_mdata_mrg %>%
  ggplot(aes(x = log10(pf_gn_count), y = log10(pf_umi_count), colour = mthds_union)) +
  geom_point(size = 0.01) +
  geom_abline() +
  facet_wrap(donor ~., nrow = 1) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  guides(color = guide_legend(title = 'Call methods', override.aes = list(size = 4), ncol = 1))

```



```{r, warning=FALSE, message=FALSE, fig.width=4, fig.height=4.5}
## umi
plts_nc <- imap(all_seu_prelim_stg_norm, possibly(~FeaturePlot(.x, label = T, features = "pf_umi_count") +  theme_void() + theme(legend.position = "none") + labs(title = .y), "no cells"))

```

```{r, warning=FALSE, message=FALSE, fig.height= 3.5, fig.width= 10, eval=T}
plot_grid(plotlist = plts_nc, ncol = 4)
```


```{r, warning=FALSE, message=FALSE, fig.width=16, fig.height=4}
imap(all_seu_prelim_stg_norm, ~DimPlot(.x, label = T, group.by = c("stage_fld", "seurat_clusters"), ncol = 3) + 
       labs(title = .y) )

```


```{r, warning=FALSE, message=FALSE, fig.width=4, fig.height=4.5}
## in cellbender but not in edrops
plts_l100 <- imap(all_seu_prelim_stg_norm, possibly(~DimPlot(.x, label = T, cells.highlight =  colnames(.x[,.x@meta.data$pf_gn_count <100]), sizes.highlight = 0.05) + 
                                                      theme_void() + theme(legend.position = "none", panel.border = element_rect(color = "black", fill = NA, size = 1)) +
                                                      labs(title = .y), "no cells"))

```

```{r, warning=FALSE, message=FALSE, fig.height= 3.5, fig.width= 10, eval=T}
plot_grid(plotlist = plts_l100, ncol = 4)
```




## 2. CELLBENDER METRICS CELL VISUALIZATIONS



### 2.1 CELL PROBABILITY VISUALIZATIONS


```{r, warning=FALSE, message=FALSE, fig.width=4, fig.height=4.5}
## in cellbender but not in edrops
plts_seu_cp9 <- imap(all_seu_prelim_stg_norm, possibly(~DimPlot(.x, label = T, cells.highlight =  colnames(.x[ , .x@meta.data$cell_probability < 0.9]), sizes.highlight = 0.05) + theme_void() + theme(legend.position = "none", panel.border = element_rect(color = "black", fill = NA, size = 1))  + labs(title = .y), "no cells"))

```

```{r, warning=FALSE, message=FALSE, fig.height= 3.5, fig.width= 10, eval=T}
plot_grid(plotlist = plts_seu_cp9, ncol = 4)

```


### 2.1 BACKGROUND FRACTION VISUALIZATION


## Assess cellbender estimated metrics cell probability and background fraction
```{r, fig.height = 3.5, fig.width = 12}
all_seu_prelim_stg_norm_mdata_mrg %>%
  ggplot(aes(x = mthds_union, y = background_fraction)) +
  geom_boxplot() +
  facet_wrap(donor ~., ncol = 4) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


```{r, fig.height = 3.5, fig.width = 15}
all_seu_prelim_stg_norm_mdata_mrg %>%
  ggplot(aes(x = pf_gn_count, y = background_fraction, colour = mthds_union)) +
  geom_point(size = 0.01) +
  geom_hline(yintercept = 0.5) +
  geom_vline(xintercept = 0.5) +
  geom_abline() +
  facet_wrap(donor ~., ncol = 4, scales = "free") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  guides(color = guide_legend(title = 'Call methods', override.aes = list(size = 4), ncol = 1))

```

```{r, warning=FALSE, message=FALSE, fig.width=4, fig.height=4.5}
## background fraction cellbender
all_seu_prelim_stg_norm_mdata_mrg %>% count(donor, background_fraction> 0.5 )

```


```{r, warning=FALSE, message=FALSE, fig.width=4, fig.height=4.5}
## 
plts_seu_bg5 <- imap(all_seu_prelim_stg_norm, possibly(~DimPlot(.x, label = T, cells.highlight =  colnames(.x[, .x@meta.data$background_fraction > 0.5 ]), sizes.highlight = 0.05) + theme_void() + theme(legend.position = "none", panel.border = element_rect(color = "black", fill = NA, size = 1))  +
                                                         labs(title = .y), "no cells"))

```

```{r, warning=FALSE, message=FALSE, fig.height= 3.5, fig.width= 15, eval=T}
plot_grid(plotlist = plts_seu_bg5, ncol = 4)

```



```{r, warning=FALSE, message=FALSE}
#ap2g
imap(all_seu_prelim_stg_norm, ~FeaturePlot(.x, dims = c(1,2), features = "PF3D7-1222600", order = T, slot = "data")+ labs(title = .y))
```


```{r, warning=FALSE, message=FALSE, fig.width=4, fig.height=4.5}
## background fraction cellbender
plts_cbg <- imap(all_seu_prelim_stg_norm, possibly(~FeaturePlot(.x, label = T, features = "background_fraction") +  theme_void() + theme(legend.position = "none") + labs(title = .y), "no cells"))

```

```{r, warning=FALSE, message=FALSE, fig.height= 3.5, fig.width= 10, eval=T}
plot_grid(plotlist = plts_cbg, ncol = 4)
```


### 4 UMI AND GENE COUNTS OUTLIERS VISUALIZATIONS AND REMOVAL


########################### START Remove outliers ##############################

```{r}
# gn_uc <- c(1900, 2600, 3300, 3800, 3200)
# gn_lc <- c(250, 250, 250, 300, 250)
# all_seu_prelim_stg_norm <- pmap(list(all_seu_prelim_stg_norm, rep('pf_gn_count', length(sample_name)), rep('pf_umi_count', length(sample_name))), ~{
all_seu_prelim_stg_norm <- pmap(list(all_seu_prelim_stg_norm, rep('nFeature_RNA', length(sample_name)), rep('nCount_RNA', length(sample_name))), ~{
  ..1@meta.data %>%
    rownames_to_column('bcode') %>%
    group_by(stage_fld) %>%
    mutate("gn_lci_cl_stg" = quantile(!!rlang::sym(..2), 0.01, na.rm = T),
           "gn_uci_cl_stg" = quantile(!!rlang::sym(..2), 0.99, na.rm = T),
           "rna_lci_cl_stg" = quantile(!!rlang::sym(..3),  0.01, na.rm = T),
           "rna_uci_cl_stg" = quantile(!!rlang::sym(..3), 0.99, na.rm = T)
    ) %>%
    ungroup()%>%
    group_by(seurat_clusters) %>%
    mutate("gn_lci_cl_seu" = quantile(!!rlang::sym(..2), 0.01, na.rm = T),
           "gn_uci_cl_seu" = quantile(!!rlang::sym(..2), 0.99, na.rm = T),
           "rna_lci_cl_seu" = quantile(!!rlang::sym(..3),  0.01, na.rm = T),
           "rna_uci_cl_seu" = quantile(!!rlang::sym(..3), 0.99, na.rm = T)
    ) %>%
    ungroup()%>%
    select(bcode,contains("ci_cl")) %>%
    column_to_rownames('bcode')%>%
    AddMetaData(..1, .)
})
```


```{r, warning=FALSE, message=FALSE, fig.width=4, fig.height=4.5}
## in cellbender but not in edrops
plts_bg <- imap(all_seu_prelim_stg_norm, possibly(~{
  DimPlot(.x, label = T, 
          cells.highlight =  .x@meta.data %>% rownames_to_column('bcode') %>% filter(pf_gn_count<gn_lci_cl_seu | nCount_RNA < rna_lci_cl_seu | nCount_RNA > rna_uci_cl_seu | pf_gn_count > gn_uci_cl_seu)  %>% pull(bcode), 
          sizes.highlight = 0.05) + 
    theme_void() + theme(legend.position = "none", panel.border = element_rect(color = "black", fill = NA, size = 1))  +
    labs(title = .y)}, "no cells"))

```

```{r, warning=FALSE, message=FALSE, fig.height= 3.5, fig.width= 10, eval=T}
plot_grid(plotlist = plts_bg, ncol = 4)
```
#### Violin plots


```{r, fig.height = 8, fig.width = 12}

fct_theme <- theme(text=element_text(size=14),
                   axis.text=element_text(size=20), 
                   axis.title=element_text(size=20,face="bold"), 
                   legend.text = element_text(size=20), 
                   legend.title = element_text(size=20,face="bold"),
                   strip.text =element_text(size=22,face="plain", colour = 'black'))


```


```{r, fig.height = 4.5, fig.width = 12}

md_don_cl <- map(all_seu_prelim_stg_norm, ~.@meta.data #%>% 
                 # mutate(across(c(sim_lab1, sim_lab2), as.numeric))
) %>% 
  bind_rows(., .id = 'donor') %>%
  mutate(across(donor, ~factor(., levels = as.character(sample_name))))

```

```{r, fig.height = 8, fig.width = 12}

my21_qc_vln_fn <- function(clust_var = "StagePrelimC", clust_cols, ci_sufx = "_stg", row_numbers = 2) {
  pmap(list(c('pf_umi_count', 'pf_gn_count'), c('Transcript', 'Gene'), paste0(c("rna_lci_cl","gn_lci_cl"), ci_sufx), paste0(c("rna_uci_cl","gn_uci_cl"), ci_sufx)), ~{
    
    md_don_cl %>% 
      ggplot(aes(x = !!rlang::sym(clust_var), y = !!rlang::sym(..1)/1000, color = !!rlang::sym(clust_var))) +
      scale_color_manual(values = clust_cols) +
      theme_classic() +
      geom_violin(drop = F) +
      ggforce::geom_sina(size = 0.1,  alpha = 0.5)+
      facet_wrap(donor ~., nrow = row_numbers, scales = "free_x")+
      labs(y = "Gene count (X 1000)")+
      geom_segment(aes(x=as.numeric(!!rlang::sym(clust_var))-0.3, xend=as.numeric(!!rlang::sym(clust_var))+0.3,
                       y = !!rlang::sym(..3)/1000, yend= !!rlang::sym(..3)/1000 ), inherit.aes=F, color="orange", size=0.5)+
      geom_segment(aes(x=as.numeric(!!rlang::sym(clust_var))-0.3, xend=as.numeric(!!rlang::sym(clust_var))+0.3,
                       y = !!rlang::sym(..4)/1000, yend= !!rlang::sym(..4)/1000 ), inherit.aes=F, color="red", size=0.5) +
      labs(x= 'Cluster', y= paste0(..2, ' count (X 1000)')) +
      fct_theme+
      theme(legend.position = "bottom",
            axis.text.x = element_text(angle =40, hjust = 1, size = 10)) +
      guides(color = guide_legend(title = 'Cluster',override.aes = list(size = 4), nrow = 1))
  })
}


```


```{r, fig.height = 5, fig.width = 15}
my21_qc_vln_fn(clust_var = "stage_fld", clust_cols = sp_col_new, ci_sufx = "_stg", row_numbers = 1 ) 

```



```{r}

## filter based on gene/transcript count
all_seu_filtOlMtCb_rm <- map(all_seu_prelim_stg_norm, ~subset(.x, nFeature_RNA < gn_lci_cl_seu | nCount_RNA < rna_lci_cl_seu | nCount_RNA > rna_uci_cl_seu | nFeature_RNA > gn_uci_cl_seu))

map(all_seu_filtOlMtCb_rm, ~.x@meta.data %>% count(stage_fld))

```


```{r, fig.height = 4, fig.width = 15}
my21_qc_vln_fn(clust_var = "seurat_clusters", clust_cols = cluster_cols, ci_sufx = "_seu", row_numbers = 1 ) 

```


```{r, warning=FALSE, message=FALSE, fig.width=4, fig.height=4.5}
## 
plts_outlrs <- imap(all_seu_prelim_stg_norm, possibly(~{
  DimPlot(.x, label = T, 
          cells.highlight =  .x@meta.data %>% rownames_to_column('bcode') %>% filter(nFeature_RNA<gn_lci_cl_seu | nCount_RNA < rna_lci_cl_seu | nCount_RNA > rna_uci_cl_seu | nFeature_RNA > gn_uci_cl_seu)  %>% pull(bcode), 
          sizes.highlight = 0.05) + 
    theme_void() + theme(legend.position = "none", panel.border = element_rect(color = "black", fill = NA, size = 1))  +
    labs(title = .y)}, "no cells"))

```

```{r, warning=FALSE, message=FALSE, fig.height= 3.5, fig.width= 15, eval=T}
plot_grid(plotlist = plts_outlrs, ncol = 4)
```



Number of cells per strain_gt to remove based on REVISED CUT-OFFS to assist souporcell to perform clustering correctly
```{r}

pmap(list(all_seu_prelim_stg_norm), ~..1@meta.data %>% dplyr::count(nFeature_RNA<gn_lci_cl_seu ,nCount_RNA < rna_lci_cl_seu, nCount_RNA > rna_uci_cl_seu, nFeature_RNA > gn_uci_cl_seu ))
```


```{r, warning=FALSE, message=FALSE}
#REX1
imap(all_seu_prelim_stg_norm, ~FeaturePlot(.x, dims = c(1,2), features = "PF3D7-0935900", reduction = "umap")+ labs(title = .y))

```


```{r, warning=FALSE, message=FALSE}
#ap2g
imap(all_seu_prelim_stg_norm, ~FeaturePlot(.x, dims = c(1,2), features = "PF3D7-1222600", order = T, slot = "data")+ labs(title = .y))
```
```{r, warning=FALSE, message=FALSE}
#ap2g
imap(all_seu_prelim_stg_norm, ~FeaturePlot(.x, dims = c(2,3), features = "PF3D7-1222600", order = T, slot = "data")+ labs(title = .y))
imap(all_seu_prelim_stg_norm, ~FeaturePlot(.x, dims = c(1,3), features = "PF3D7-1222600", order = T, slot = "data")+ labs(title = .y))
```

Remove those with too few genes transcripts

## 3.  REMOVALS

###########################START Remove cells######################################
Remove those with too few genes transcripts

```{r}

after_mt_filt <- map(all_seu_prelim_stg_norm, dim)%>% as.data.frame() %>% .[2,]
after_mt_filt
```
## 3.1  OUTLIER REMOVALS
Check the number of lost cells from original talleh dataset before outlier removal
```{r, eval=T}
all_seu_ta_mt_filt_missing_subset_b4OutlierRm <- msc49_ta19[,!(colnames(msc49_ta19) %in% colnames(all_seu_prelim_stg_norm[["MSC49"]]))]
all_seu_ta_mt_filt_missing_subset_b4OutlierRm
```

```{r}
## Barcodes removed
low_count_rm_bcodes <- map(all_seu_prelim_stg_norm, ~colnames(subset(.x, nFeature_RNA > gn_lci_cl_seu & nCount_RNA > rna_lci_cl_seu & nCount_RNA < rna_uci_cl_seu & nFeature_RNA < gn_uci_cl_seu, invert = T)))

```

```{r}

## filter based on gene/transcript count
all_seu_filtOlMtCb <- map(all_seu_prelim_stg_norm, ~subset(.x, nFeature_RNA > gn_lci_cl_seu & nCount_RNA > rna_lci_cl_seu & nCount_RNA < rna_uci_cl_seu & nFeature_RNA < gn_uci_cl_seu))

aftr_low_count_rm <- map(all_seu_filtOlMtCb, dim)%>% as.data.frame() %>% .[2,]
aftr_low_count_rm

```
Check the number of lost cells from original talleh dataset after outlier removal
```{r, eval=T}
all_seu_ta_mt_filt_missing_subset_afterOutlierRm <- msc49_ta19[,!(colnames(msc49_ta19) %in% colnames(all_seu_filtOlMtCb[["MSC49"]]))]
all_seu_ta_mt_filt_missing_subset_afterOutlierRm
```

## 3.1  CELL PROBABILITY REMOVALS

```{r}
## 
map(all_seu_filtOlMtCb, ~.x@meta.data %>% count(cell_probability >= 0.9))
```


```{r}
## 
imap(all_seu_filtOlMtCb, ~.x@meta.data %>% count((cell_probability >= 0.9), stage_fld))

```

```{r}
## Barcodes removed
cell_prob9_rm_bcodes <- map(all_seu_filtOlMtCb, possibly(~colnames(subset(.x, cell_probability >= 0.9, invert = T))))

```

```{r}
## 
all_seu_filtOlMtCb <- map(all_seu_filtOlMtCb, ~subset(.x, cell_probability >= 0.9))


aftr_cell_prob9_rm <- map(all_seu_filtOlMtCb, dim)%>% as.data.frame() %>% .[2,]
aftr_cell_prob9_rm
```


## 3.1  BACKGROUND FRACTION REMOVALS


```{r, warning=FALSE, message=FALSE, fig.width=4, fig.height=4.5}
## 
map(all_seu_filtOlMtCb, ~.x@meta.data %>% count(background_fraction > 0.5  ))

```

```{r}
## 
map(all_seu_filtOlMtCb, ~.x@meta.data %>% count(background_fraction <= 0.5))

```


```{r, fig.height = 3.5, fig.width = 12}
all_seu_filtOlMtCb_mdata_mrg <- map(all_seu_filtOlMtCb, ~.@meta.data) %>% 
  bind_rows(., .id = 'donor') 


```

```{r, warning=FALSE, message=FALSE, fig.width=4, fig.height=4.5}
#
plts_seu_bg5 <- imap(all_seu_filtOlMtCb, possibly(~{
  DimPlot(.x, label = T, 
          cells.highlight =  colnames(
            .x[, !(.x@meta.data$background_fraction <= 0.5 )]
          ), 
          sizes.highlight = 0.05) +   
    theme_void() + 
    theme(legend.position = "none", panel.border = element_rect(color = "black", fill = NA, size = 1))  +
    labs(title = .y)
}, "no cells"))

```

```{r, warning=FALSE, message=FALSE, fig.height= 3.5, fig.width= 10, eval=T}
plot_grid(plotlist = plts_seu_bg5[unlist(map(plts_seu_bg5, ~!is.character(.x)))], ncol = 4)

```


```{r}
## Barcodes removed
cb_bg_frac_rm_bcodes <- map(all_seu_filtOlMtCb, 
                            possibly(~colnames(.x[, !(.x@meta.data$background_fraction <= 0.5 )]
                            )))

```


```{r}
# 
all_seu_filtOlMtCb <- map(all_seu_filtOlMtCb, 
                          ~.x[, (.x@meta.data$background_fraction <= 0.5 )]
)

aftr_cb_bg_frac_rm <- map(all_seu_filtOlMtCb, dim) %>% as.data.frame() %>% .[2,]
aftr_cb_bg_frac_rm
```

###########################END Remove cells######################################


```{r, warning=FALSE, message=FALSE}
#ap2g
imap(all_seu_filtOlMtCb, ~FeaturePlot(.x, dims = c(1,2), features = "PF3D7-1222600", order = T, slot = "counts")+ labs(title = .y))

```


```{r, warning=FALSE, message=FALSE}

imap(all_seu_filtOlMtCb, ~DimPlot(.x, dims = c(1,2), group.by = "stage_fld", pt.size = 0.1)+ scale_color_manual(values = sp_col_new) + labs(title = .y))

```

```{r, warning=FALSE, message=FALSE}

imap(all_seu_filtOlMtCb, ~DimPlot(.x, dims = c(1,2), group.by = "seurat_clusters", pt.size = 0.1, label = T)+ labs(title = .y))

```

```{r, eval=T}
## write out seurat bpcells object files for normalization processing in farm
imap(all_seu_filtOlMtCb, ~saveRDS(.x, paste0("data/processed/Pf/", .y,"/seu_obj/all_seu_filtOlMtCb.RDS")))

```


```{r}

filt_flow <- list(after_umi100_filt, after_mt_filt, aftr_low_count_rm,  aftr_cell_prob9_rm, aftr_cb_bg_frac_rm) %>% set_names("100umi_filt", "mt_filt", "low_count",  "cell_prob9", "cb_bg_frac") %>% bind_rows(., .id = "filt_metric") %>% mutate(across(filt_metric, ~factor(., levels = c("100umi_filt", "mt_filt", "low_count",  "cell_prob9", "cb_bg_frac"))))

```

```{r}
## Number of cell population removed by respective filter
filt_diff <- (filt_flow[-nrow(filt_flow), ] - filt_flow[-1, ]) 

filt_diff$filt_metric <- paste0(filt_flow[-1, ]$filt_metric, "_rm")
filt_diff_t <- filt_diff  %>% column_to_rownames("filt_metric") %>% t() %>% as.data.frame() 
filt_diff_t

## Proportion of cell population removed by respective filter
filt_diff_prop <- round(filt_diff/filt_flow[-nrow(filt_flow), ] * 100, 2)
filt_diff_prop$filt_metric <- paste0(filt_flow[-1, ]$filt_metric, "_rm_prop")
filt_diff_prop_t <- filt_diff_prop %>% column_to_rownames("filt_metric") %>% t() %>% as.data.frame() 

## Merge number of cells and those that are removed
filt_flow_diff <- bind_rows(
  filt_flow %>% 
    mutate(row_id = row_number(), source = 2),
  filt_diff %>% 
    mutate(row_id = row_number(), source = 2)
) #%>%
# arrange(row_number() + (row_number() > nrow(df)))

filt_flow_diff_t <- filt_flow_diff %>% arrange(row_id, source) %>% 
  select(-row_id, -source) %>% column_to_rownames("filt_metric") %>% t() %>% as.data.frame() 

filt_flow_diff_t
```


```{r}
# saveRDS(filt_flow_diff_t, "data/processed/Pf/m2_all/filt_flow_diff_t.RDS")
saveRDS(filt_flow_diff_t, "data/processed/Pf/m2_all/filt_flow_diff_t.RDS") 

```

```{r}

filt_flow %>% 
  pivot_longer(cols = contains("MSC"), names_to = "donor", values_to = "cell_counts") %>%  
  ggplot(aes(x = filt_metric, y = cell_counts, colour = donor, group = donor)) + geom_point() + geom_line()

```


```{r}
poorcellQC_bcodes_rm <- list(mt_filt_rm_bcodes, low_count_rm_bcodes, cell_prob9_rm_bcodes, cb_bg_frac_rm_bcodes) %>% 
  set_names(c("mt_filt_rm_bcodes", "low_count_rm_bcodes", "cell_prob9_rm_bcodes", "cb_bg_frac_rm_bcodes"))

```


```{r, message=FALSE, warning=FALSE, eval=T}
# Save the poorQC cells barcodes for plotting
saveRDS(poorcellQC_bcodes_rm, "data/processed/Pf/m2_all/lowQC_rm_plotting/poorcellQC_bcodes_rm.RDS")
```

!!!NOTE - If the below breaks the code then just remove and rerun. I added it but didn't rerun
```{r, warning=FALSE, message=FALSE}

all_seu_filtOlMtCb_umap_mdata <- map(all_seu_filtOlMtCb, ~{
  .x@meta.data %>% 
    rownames_to_column("bcode") %>%
    left_join(., .x@reductions$umap@cell.embeddings %>% data.frame() %>% rownames_to_column("bcode"), by = "bcode")
})
```


```{r, message=FALSE, warning=FALSE, eval=T}
# Save the poorQC cells barcodes for plotting
saveRDS(all_seu_filtOlMtCb_umap_mdata, "data/processed/Pf/m2_all/lowQC_rm_plotting/all_seu_filtOlMtCb_umap_mdata.RDS")

```

```{r, eval=T}
all_seu_filtOlMtCb[["MSC49"]]@meta.data %>% rownames_to_column("bcode") %>% left_join(msc49_ta19@meta.data %>% rownames_to_column("bcode"), by = "bcode", suffix = c("_nw", "_old")) %>% count(sR_fine_stg_nw, sR_fine_stg_old)

all_seu_filtOlMtCb[["MSC49"]]@meta.data %>% rownames_to_column("bcode") %>% right_join(msc49_ta19@meta.data %>% rownames_to_column("bcode"), by = "bcode", suffix = c("_nw", "_old")) %>% count(sR_fine_stg_nw, sR_fine_stg_old)
```


```{r, eval=T}
all_seu_ta_missing_subset <- msc49_ta19[,!(colnames(msc49_ta19) %in% colnames(all_seu_filtOlMtCb[["MSC49"]]))]

```


```{r, eval=T}
all_seu_ta_missing_subset 
```

```{r}
msc49_ta19@meta.data %>% count(scrub_doublet, scf_class_norm, scf_class_rand)
```
